### 2020-11-19 documents/ClassReportMethods.Rmd
some stuff

---
title: "ClassReportMethods"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-11-19 scripts/combogenerator.R

@@ -1,0 +1,20 @@
+abundinc <- c("Y", "N")
+MRinc <- c("Y", "N")
+nestinc <- c("Y", "N")
+
+param1 <- c("L", "H")
+param2 <- c("L", "H")
+param3 <- c("L", "H")
+param4 <- c("L", "H")
+param5 <- c("L", "H")
+
+paramcombs <- expand.grid(param1, param2, param3, param4, param5)
+paramcombs <- cbind(paramcombs, paramcombs[, 5])
+
+combs <- expand.grid(abundinc, MRinc, nestinc, param1, param2, param3, param4, param5)
+
+combs <- cbind(combs, combs[, 8])
+
+
+
+write.csv(combs, )

### 2020-11-19 scripts/dataSims/IPMSimData_nonest.R
abby edits

@@ -8,49 +8,48 @@
 
 library(nimble)
 
-n.years=10; n.data=c(50,50,50);init.age = c(100,100); 
-phi.1=0.7; phi.ad=0.76;p.1=0.98; p.ad=0.65;
+n.years=10; n.data=c(50,50,50);init.age = c(10,10); 
+phi.1=0.5; phi.ad=0.76;p.1=0.98; p.ad=0.65;
 f.1=0.8;f.ad=0.8; p.sur=0.8; p.prod=0.77
 
-# TODO
-# abby put nest params
+n.initiation.dates <- 31
+
+max.nest.age <- 30
+first.initiation.date <- 1
+last.fledge.date <- n.initiation.dates + max.nest.age
+season.length <- last.fledge.date - first.initiation.date + 1 + 2
+
+# mean clutch size
+mean.clutch.size <- 3
+
+# daily nest survival
+phi.nest <- 0.975
+
+prop.nests.found <- 0.8
+
 simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
  f.1, f.ad, p.sur, p.prod){
+  
   ti<-n.years
   ni<-n.years-1
   nd<-n.data
   Ni<-init.age
   phi1<-phi.1
   phi2<-phi.ad
-  #fl1<-f.1
-  #fl2<-f.ad
   pjuv<-p.1
   padults<-p.ad
   psur<-p.sur
-  #pprod<-p.prod
-  
-  # Prod parameters
-  
-  n.initiation.dates <- 120
-  max.nest.age <- 30
-  first.initiation.date <- 1
+  n.initiation.dates <- n.initiation.dates
+  max.nest.age <- max.nest.age
+  first.initiation.date <- first.initiation.date
   last.fledge.date <- n.initiation.dates + max.nest.age
   season.length <- last.fledge.date - first.initiation.date + 1 + 2
+  mean.clutch.size <- mean.clutch.size
+  phi.nest <- phi.nest
+  prop.nests.found <- prop.nests.found
   
-  # nests for population
-  N.nests.total <- 100
-  prop.nests.found <- 0.8
-  
-  # mean clutch size
-  mean.clutch.size <- 5
-  
-  # daily nest survival
-  phi.nest <- 0.975
-  
+  # DERIVE TRUE FECUNDITY
   true.fec <- 1/2 * mean.clutch.size * phi.nest^max.nest.age
-  
-  # observation params
[…]

### 2020-11-19 scripts/dataSims/Non_m_arrayCJSIPMNimble.R
hannah updates

@@ -9,22 +9,18 @@ source(here("scripts", "dataSims","IPMsimData_nonest.R")) #changed the MR data
 # 11.3. Example of a simple IPM (counts, capture-recapture, reproduction)
 # 11.3.1. Load data
 # Population counts (from years 1 to n.years)
-y <- df$SUR
+y <- SUR#df$SUR
 
 # Capture-recapture data (in m-array format, from years 1 to n.years)
-m <- df$ch
+m <- ch#df$ch
 
 
 #Productivity/nest success not in the data sim right now
 #these will need to be changed once we have the nest success data
 # Productivity data (from years 1 to n.years-1)
-J<-df$nestlings
-R<-df$R
+J<-nestlings#df$nestlings
+R<-R#df$R
 
-
-#Use the IPM from WinBugs, but put in nimble
-# 11.3.2. Analysis of the model
-# Specify model in BUGS language
 library("nimble")
 IPMmod<-nimbleCode({
   #-------------------------------------------------
@@ -36,85 +32,57 @@ IPMmod<-nimbleCode({
   #  - All vital rates assumed to be constant
   #-------------------------------------------------
   
-  #-------------------------------------------------
-  # 1. Define the priors for the parameters
-  #-------------------------------------------------
-  # Observation error
-  tauy <- pow(sigma.y, -2)
-  sigma.y ~ dunif(0, 50)
-  sigma2.y <- pow(sigma.y, 2)
-  
+  #population projection part
   # Initial population sizes
-  # Note that this part is different than in BUGS:
-  # 1. JAGS seems to be very sensitive to the choice of the prior distribution for the initial population sizes and of the choice of the observation model (priors)
-  # 2. Since the initial population sizes are used in binomial distributions, the numbers must be integers, otherwise JAGS does not run.
-  # The following specification seems to work very well and produces similar results as BUGS:
-  
-  n1 ~ T(dnorm(25, tauy),0,)     # 1-year
-  nad ~ T(dnorm(25, tauy),0,)    # Adults
-  N1[1] <- round(n1)
-  Nad[1] <- round(nad)
[…]

### 2020-11-19 documents/simulate data.R
add files

@@ -1,0 +1,24 @@
+# load libraries ####
+
+library(here)
+library(dplyr)
+library(readxl)
+library(lubridate)
+library(hms)
+library(stringr)
+library(tidyr)
+
+# load data ####
+
+# site codes
+sites <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx", range = "E2:K5"))
+
+# create a field for file names
+
+# source
+
+source(here("scripts", "dataSims", ""))
+
+# this should load the function we need
+
+# save data objects 

### 2020-11-19 documents/simulate data.R
scaffolding code for running simulations

@@ -1,24 +1,0 @@
-# load libraries ####
-
-library(here)
-library(dplyr)
-library(readxl)
-library(lubridate)
-library(hms)
-library(stringr)
-library(tidyr)
-
-# load data ####
-
-# site codes
-sites <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx", range = "E2:K5"))
-
-# create a field for file names
-
-# source
-
-source(here("scripts", "dataSims", ""))
-
-# this should load the function we need
-
-# save data objects 

### 2020-11-19 scripts/dataSims/simulate data.R

@@ -1,0 +1,78 @@
+# load libraries ####
+
+library(here)
+library(dplyr)
+library(readxl)
+library(lubridate)
+library(hms)
+library(stringr)
+library(tidyr)
+
+# source files ######
+
+# source data simulation function
+# source model function
+
+# load scenarios ####
+
+# site
+paramlevels <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx"), range = "E1:K4")
+paramlevels[, 1] <- c("L", "M", "H")
+colnames(paramlevels)[1] <- "Level"
+scenarios <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx"), range = "A5:K22")
+scenarios <- scenarios[-c(6, 12), ]
+
+n.scenarios <- max(scenarios$`Scenario Number`)
+sims.per <- max(scenarios$`Sims per`)
+
+scenarios <- scenarios %>% 
+  mutate(`MR detection` = case_when(`MR detection` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "MR detection"]),
+                                    `MR detection` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "MR detection"]),
+                                    `MR detection` == "H" ~ as.numeric(paramlevels[paramlevels$Level == "H", "MR detection"]))
+         ) %>% 
+  mutate(`Abund detection` = case_when(`Abund detection` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "Abund detection"]),
+                                    `Abund detection` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "Abund detection"]),
+                                    `Abund detection` == "H" ~ as.numeric(paramlevels[paramlevels$Level == "H", "Abund detection"]))
+  ) %>% 
+  mutate(`Adult Surv` = case_when(`Adult Surv` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "Adult Surv"]),
+                                    `Adult Surv` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "Adult Surv"]),
+                                    `Adult Surv` == "H" ~ as.numeric(paramlevels[paramlevels$Level == "H", "Adult Surv"]))
+  ) %>% 
[…]

### 2020-11-19 scripts/dataSims/simulate data.R
some updates

@@ -59,8 +59,13 @@ scenarios <- scenarios %>%
 
 for (s in 1:n.scenarios) {
   for (i in 1:sims.per) {
-    # run model with appropriate simulation
-    
+    if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 1) {
+      # run full model
+    } else if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 0) {
+      # run no nest model
+    } else if (scenarios[s, `MR Included`] == 0 & scenarios[s, `Nests Includedd`] == 1) {
+      # run no resight model
+    }
     
     # save results to file and to environment
     outcopy <- out

### 2020-11-20 scripts/dataSims/IPMSimData_nonest.R
abby updates

@@ -9,7 +9,7 @@
 library(nimble)
 
 n.years=10; n.data=c(50,50,50);init.age = c(10,10); 
-phi.1=0.5; phi.ad=0.76;p.1=0.98; p.ad=0.65;
+phi.1=0.85; phi.ad=0.76;p.1=0.98; p.ad=0.65;
 f.1=0.8;f.ad=0.8; p.sur=0.8; p.prod=0.77
 
 n.initiation.dates <- 31
@@ -29,8 +29,12 @@ prop.nests.found <- 0.8
 
 n.sam<-3
 
+visit.interval <- 3
+
 simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
- f.1, f.ad, p.sur, p.prod){
+ f.1, f.ad, p.sur, 
+ n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 
+ season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval){
   
   ti<-n.years
   ni<-n.years-1
@@ -49,6 +53,7 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
   mean.clutch.size <- mean.clutch.size
   phi.nest <- phi.nest
   prop.nests.found <- prop.nests.found
+  visit.interval <- visit.interval
   
   # DERIVE TRUE FECUNDITY
   true.fec <- 1/2 * mean.clutch.size * phi.nest^max.nest.age
@@ -169,8 +174,10 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
         # assume everybody attempts a nest
         N.nests.total <- sum(IND[age, year, ], na.rm = TRUE)
         
-        total.nests.age <- matrix(NA, nrow = N.nests.total, ncol = season.length)
-        total.nests.status <- matrix(NA, nrow = N.nests.total, ncol = season.length)
+        inds.nesting <- which(!is.na(IND[age, year, ]))
+        
+        total.nests.age <- matrix(NA, nrow = Ntotal, ncol = season.length)
+        total.nests.status <- matrix(NA, nrow = Ntotal, ncol = season.length)
         
         # initiation date
         # uniform across season length
@@ -179,17 +186,17 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
         
         for (i in 1:N.nests.total) {
[…]

### 2020-11-20 scripts/dataSims/IPMSimData_nonest.R
added nest stuff but model doesnt work :(

@@ -34,7 +34,7 @@ visit.interval <- 3
 simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
  f.1, f.ad, p.sur, 
  n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 
- season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval){
+ season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval, n.sam){
   
   ti<-n.years
   ni<-n.years-1
@@ -54,6 +54,7 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
   phi.nest <- phi.nest
   prop.nests.found <- prop.nests.found
   visit.interval <- visit.interval
+  n.sam <- n.sam
   
   # DERIVE TRUE FECUNDITY
   true.fec <- 1/2 * mean.clutch.size * phi.nest^max.nest.age
@@ -441,9 +442,10 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
               H=Htidy, Fledged=Fledgedtidy, 
               first.nest = first.nesttidy, last.nest = last.nesttidy, 
               max.nest.age = max.nest.age, 
-              n.nests = N.nests.found, n.succ.nests = N.nests.successful))
+              n.nests = N.nests.found, n.succ.nests = N.nests.successful, 
+              n.sam = n.sam))
 }
 df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
                f.1, f.ad, p.sur, 
                n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 
-               season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval)  
+               season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval, n.sam)  

### 2020-11-20 scripts/dataSims/Non_m_arrayCJSIPMNimble.R

@@ -9,17 +9,27 @@ source(here("scripts", "dataSims","IPMsimData_nonest.R")) #changed the MR data
 # 11.3. Example of a simple IPM (counts, capture-recapture, reproduction)
 # 11.3.1. Load data
 # Population counts (from years 1 to n.years)
-y <- SUR#df$SUR
+y <- df$SUR
+n.sam <- df$n.sam
 
 # Capture-recapture data (in m-array format, from years 1 to n.years)
-m <- ch#df$ch
-
+m <- df$ch
+first <- df$first
+age_ch <- df$age_ch
 
 #Productivity/nest success not in the data sim right now
 #these will need to be changed once we have the nest success data
 # Productivity data (from years 1 to n.years-1)
-J<-nestlings#df$nestlings
-R<-R#df$R
+#J<-nestlings#df$nestlings
+#R<-R#df$R
+ 
+H <- df$H
+Fledged <- df$Fledged
+first.nest <- df$first.nest
+last.nest <- df$last.nest
+max.nest.age <- df$max.nest.age
+n.nests <- df$n.nests
+n.succ.nests <- df$n.succ.nests
 
 library("nimble")
 IPMmod<-nimbleCode({
@@ -44,14 +54,14 @@ IPMmod<-nimbleCode({
   p.surv~dunif(0,1)
   #  productivity
   for (t in 1:(nyears-1)){
-    f[t] <- mean.fec
+    f[t] <- fec
   }
-  mean.fec ~ dunif(0, 20)
+  #mean.fec ~ dunif(0, 20)
   
   # 3.1. Likelihood for population population count data (state-space model)
   # 3.1.1 System process
   for (t in 2:nyears){
-    mean1[t] <- f[t-1] / 2 * mean.phi[1] * Ntot[t-1]
+    mean1[t] <- f[t-1] * mean.phi[1] * Ntot[t-1]
     N1[t] ~ dpois(mean1[t])
     Nad[t] ~ dbin(mean.phi[2], Ntot[t-1])
   }
@@ -84,11 +94,29 @@ IPMmod<-nimbleCode({
   #################################
   #nest success model goes here
   
-  # 3.3. Likelihood for productivity data: Poisson regression
-  for (t in 1:(nyears-1)){
-    J[t] ~ dpois(rho[t])
-    rho[t] <- R[t]*f[t]
+  # Priors and constraints #####
+  lambdaf ~ dunif(0, 10)
[…]

### 2020-11-21 documents/DraftReport.Rmd
Added draft markdown report.

---
title: 'Title'
output:
  word_document:
    reference_docx: mytemplate.docx
  pdf_document: default
---

```{r include = FALSE}

#set default performance for chunks and then modify individually if necessary
#for reference: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
# echo == show code
# warning/message == show warnings/messages
# results == type of output
# eval == whether to evaluate code or not, defaults to TRUE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)
```

```{r load packages, results = 'hide', message = FALSE}

library(tidyr)
library(dplyr)
library(cowplot)
library(ggplot2)
library(captioner)
library(knitr)
library(reshape2)
library(here)

#functions to use for calling figures/tables
figs <- captioner(prefix = "Figure")
tbls <- captioner(prefix = "Table")

#source a plot theme for figures
source(here::here('scripts', 'PlotTheme.R'))

```

```{r load data or model output, echo = FALSE}

# #load data
# rain <- data.frame(rain_std = rnorm(100, 10, 10), 
#                        site = factor(rep(1:5, each = 20)))
# 
# #handy objects for summary stats?
# avg_rain <- rain %>%
#   group_by(site) %>%
#   dplyr::summarize(avg_rain = mean(rain_std))

```

```{r create all figures and tables, echo = FALSE, results = 'hide'}

# #Figures - there might be a more eloquent way, but in the past I've found you have to create objects above (figures and tables) so you can reference them in-text and then call them below where you actually want them to appear in the document
# 
# plotA <- ggplot(rain, aes(x = site, y = rain_std, col = site, group = site)) +
#   geom_point() +
#   xlab('') + ylab('') +
#   plot_theme()
# 
# plotB <- ggplot(rain, aes(x = site, y = rain_std, group = site)) +
#   geom_boxplot() +
#   xlab('') + ylab('') +
#   plot_theme()
# 
# plots <- plot_grid(plotA, plotB, labels = "auto", align = 'v', ncol = 1, hjust = 0, label_size = 10)
# 
# #use the 'figs' function created above
# figs("plots", caption = "Scatter and boxplots of rain per site.")
# 
# 
# #create table
# rain_table <- avg_rain

```

**Title**

Authors

*Affiliations*
Text

**Abstract**

Don't forget that it's easy to call summaries from the data. For example, the sample size for this project is `r dim(rain)[1]`, and the average rainfall at site `r unique(as.numeric(rain[rain$site == 1,'site']))` over the study period is `r round(mean(rain[rain$site == 1,'rain_std']),1)`.
<!-- AJW: make comments like this -->

**Introduction**

**Methods**

*Data simulation and life history*  
We will simulate data for a hypothetical passerine (or seabird) species that is best described with two age-classes: young-of-year (YOY) and adults. We will assume an even sex ratio between males and females, but we will use a female-only model because we assume that the population dynamics of our population are driven by female abundance and demography. We assume that the population is surveyed annually just prior to chicks hatching. Young and adult survival could be assumed to be the same. Adults produce new chicks at rate f, and those chicks survive to be YOY at rate $\phi_1$. Current YOY will also breed the following year, and also produce surviving chicks at rate f*phi1$. A life cycle diagram for this population is given below:

Our hypothetical bird species nests once per year and we assume all adult females are breeders. The population is quite small, so we are able to monitor it over its entire range. Therefore we assume no immigration or emigration.

*Model framework and assumptions*  
Productivity: Nests are monitored according to the procedures described in Martin and Geupel (1993). In summary, adults are carefully monitored for behavioral clues that indicate nesting behavior, and followed until a nest is found. We assume that the nests found are independent and representative of the population. After locating a nest, we revisit the nest every 2-3 days to check its status, until it either fledges or fails. We assume our observations of the nests are perfect. It takes 30 days from nest initiation for chicks to fledge, regardless of clutch size. We assume that nests can be correctly aged when first found. We also assume the daily survival probability is the same regardless of nest stage (i.e. laying, incubating, nestlings). We use a daily nest survival model to obtain unbiased estimates for the probability of nest success (i.e. at least 1 chick fledges), and a Poisson model to estimate the average number of chicks fledged conditional on a nest being successful.

Equations

MR: Chicks are marked in the nest just prior to fledging with a unique combination of colored leg bands such that they can be individually identified over the course of their life. Every chick in every nest that fledges is banded. We perform 1 exhaustive mark-resight survey per breeding season. We will use a Cormack-Jolly-Seber (CJS) model to estimate juvenile and adult survival based on the resulting individual encounter histories. In accordance with the CJS model, we assume that each marked individual is independent with the same probability of detection and same probability of survival in and between sampling occasions. We also assume that marks are not lost and they are perfectly observed.

Equations

Counts: Count surveys are conducted 3 times per year just prior to breeding. Since we assume no immigration and emigration, and these surveys are done in quick succession, we assume the population is closed during this period of time. We assume no individual is double counted on a survey, and that individual detections are independent and occur with equal probability. Therefore we will use an N-mixture model to estimate annual population abundance from this repeated count data.

Equations

Integration: leslie matrix/pop model; joint likelihood equations

*Simulation scenarios*

Scenarios - see sheet

Metrics to monitor
CV (precision)
RMSE (accuracy)
Relative bias

Vary data quantity/combinations and demographic rates
Low MR detection vs high
Low daily nest survival vs high
Low abund detection vs high

Best scenario, using all three datasets
Potential add-on: vary life history traits:
Survival: 0.3 vs. 0.8
Fecundity: 0.3 vs. 0.5
Remove nest monitoring entirely - use other two datasets (MR and counts)
Model adjustments: informed priors? 
Sensitivity to priors?
Remove MR data entirely - use other two datasets (nest monitoring and counts)
Model adjustments: informed priors?
Sensitivity to priors?
Poorer quality MR
Quantity: decrease N (20 individuals vs. 100 resights per year)
Quality: decrease p (0.1, 0.5, 0.8) or annual surveys (5 vs. 10)

*Estimation and diagnostics*  
Nimble; iterations, burn-in, etc
Examined relative bias, RMSE, etc

**Results**

**Discussion**

\newpage

**Literature cited**

\newpage

**Figures and Tables**

```{r, fig.height = 8.5, fig.width = 7, fig.align = 'center'}

plots

```

`r figs("plots", caption = "Scatter and boxplots of rain per site.")`

\newpage

`r tbls("rain_table", caption = "Table of rain by site.")`
```{r}

kable(rain_table, align = 'c', format.args = list(big.mark = ","))

```

### 2020-11-21 scripts/PlotTheme.R

@@ -1,0 +1,42 @@
+
+
+plot_theme <- function(...) {
+  theme(
+    #text = element_text(size = 11),
+    # axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, color = "black", size = 10), 
+    axis.text.x = element_text(vjust = 0.5, color = "black", size = 10), 
+    axis.text = element_text(vjust = 0.5, color = "black", size = 10), 
+    axis.title = element_text(size = 11),
+    axis.line.y = element_line(colour = "black"), 
+    axis.line.x = element_line(colour = "black"), 
+    plot.background = element_rect(), 
+    panel.background = element_rect(fill = 'white'), 
+    #panel.border = element_rect(fill = NA), #for square around plot
+    panel.grid = element_blank(), 
+    legend.key = element_blank(),
+    strip.background = element_blank(), 
+    strip.text = element_text(size = 10),
+    legend.text = element_text(size = 9),
+    ...)
+}
+
+plot_theme_x90 <- function(...) {
+  theme(
+    #text = element_text(size = 11),
+    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, color = "black", size = 10),
+    # axis.text.x = element_text(vjust = 0.5, color = "black", size = 10), 
+    axis.text = element_text(vjust = 0.5, color = "black", size = 10), 
+    axis.title = element_text(size = 11),
+    axis.line.y = element_line(colour = "black"), 
+    axis.line.x = element_line(colour = "black"), 
+    plot.background = element_rect(), 
+    panel.background = element_rect(fill = 'white'), 
+    #panel.border = element_rect(fill = NA), #for square around plot
+    panel.grid = element_blank(), 
+    legend.key = element_blank(),
+    strip.background = element_blank(), 
+    strip.text = element_text(size = 10),
+    legend.text = element_text(size = 9),
+    ...)
+}
+rainbow <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")

### 2020-11-21 scripts/dataSims/Non_m_arrayCJSIPMNimble.R
Found the reason for the slice sampler issue, it was how we were projecting the population. Needs a double check to make sure it is right!

@@ -61,9 +61,15 @@ IPMmod<-nimbleCode({
   # 3.1. Likelihood for population population count data (state-space model)
   # 3.1.1 System process
   for (t in 2:nyears){
-    mean1[t] <- f[t-1] * mean.phi[1] * Ntot[t-1]
+    #I think this was causing the problem, it was a little mis-specified
+    #Double check that this is right, but it doesnt cause an error in 
+    #Slice samplers this way
+    mean1[t]<-((f[t-1]*N1[t-1])+(f[t-1]*Nad[t-1]))*mean.phi[1]
+    #for reference, this is how it was:
+    #mean1[t] <- f[t-1] * mean.phi[1] * Ntot[t-1]
     N1[t] ~ dpois(mean1[t])
-    Nad[t] ~ dbin(mean.phi[2], Ntot[t-1])
+    #Also changed that the total is the sum of N1 and Nad, it was just Ntot before
+    Nad[t] ~ dbin(mean.phi[2], (N1[t-1]+Nad[t-1]))
   }
   for (t in 1:nyears){
     Ntot[t] <- Nad[t] + N1[t] 
@@ -185,8 +191,8 @@ inits <- list(mean.phi=runif(2,0,1),
               #mean.fec = runif(1, 0, 10), 
               z=z.state,
               p.surv=runif(1,0,1),
-              n1.start=sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
-              nad.start=sample(1:30,1),
+              n1.start=100,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
+              nad.start=100,#sample(1:30,1),
               phi.nest = runif(1, 0, 1), 
               lambdaf = runif(1, 0, 10),
               H = Hinits
@@ -200,8 +206,11 @@ conf<-configureMCMC(mod)
 conf$addMonitors(parameters)
 Rmcmc<-buildMCMC(conf)
 Cmodel<-compileNimble(mod)
+#for 
+#Cmodel$setInits(inits)
+#Cmodel$setData(newdata)
 Cmcmc<-compileNimble(Rmcmc, project=Cmodel)
-Cmcmc$run(thin=10, reset=T, niter=10000, nburnin=5000)
+Cmcmc$run(thin=10, reset=T, niter=100000, nburnin=50000)
 out<-as.data.frame(as.matrix(Cmcmc$mvSamples))
 
 # TODO

### 2020-11-21 scripts/dataSims/Non_m_arrayCJSIPMNimble.R
Changed the prior for phi.nest to unif(0.925,1) The model doesnt throw the same errors with that prior

@@ -64,12 +64,13 @@ IPMmod<-nimbleCode({
     #I think this was causing the problem, it was a little mis-specified
     #Double check that this is right, but it doesnt cause an error in 
     #Slice samplers this way
-    mean1[t]<-((f[t-1]*N1[t-1])+(f[t-1]*Nad[t-1]))*mean.phi[1]
+    #mean1[t]<-((f[t-1]*N1[t-1])+(f[t-1]*Nad[t-1]))*mean.phi[1]
     #for reference, this is how it was:
     #mean1[t] <- f[t-1] * mean.phi[1] * Ntot[t-1]
-    N1[t] ~ dpois(mean1[t])
+    N1[t] ~ dpois(((f[t-1]*N1[t-1])+(f[t-1]*Nad[t-1]))*mean.phi[1])
+    #N1[t] ~ dpois(mean1[t])
     #Also changed that the total is the sum of N1 and Nad, it was just Ntot before
-    Nad[t] ~ dbin(mean.phi[2], (N1[t-1]+Nad[t-1]))
+    Nad[t] ~ dbin(mean.phi[2],(Ntot[t-1]))
   }
   for (t in 1:nyears){
     Ntot[t] <- Nad[t] + N1[t] 
@@ -103,7 +104,9 @@ IPMmod<-nimbleCode({
   # Priors and constraints #####
   lambdaf ~ dunif(0, 10)
   
-  phi.nest ~ dunif(0, 1)
+  #changed this prior, since if it is too low the population goes negative
+  #and I gather from looking that this may be reasonable?
+  phi.nest ~ dunif(0.925, 1)
    
   for (t in 1:nyears) {
     for (n in 1:n.nests[t]) {
@@ -118,8 +121,7 @@ IPMmod<-nimbleCode({
   }
 
   # Derived quantities #####
-  
-  fec <- 1/2 * lambdaf * phi.nest^max.nest.age
+  fec <- 1/2 * lambdaf * (phi.nest^max.nest.age)
   
   # END derived quantities
   
@@ -168,7 +170,7 @@ z.state <- state.data(m)
 # data
 #HAS- can probably clean all of this up when the data sim is ready to go, 
 #so it automatically inputs the right data and constants 
-datipm <- list(ch.y = m, y = y, #J = J, R = R, 
+datipm <- list(ch.y = m, y = y, 
                H = H, 
                Fledged = Fledged)
 constants<-list(nyears = ncol(m), 
@@ -189,12 +191,13 @@ Hinits[is.na(H)] <- 1
[…]

### 2020-11-21 scripts/dataSims/IPMSimData_nonest.R
change defaults

@@ -8,9 +8,9 @@
 
 library(nimble)
 
-n.years=10; n.data=c(50,50,50);init.age = c(10,10); 
-phi.1=0.85; phi.ad=0.76;p.1=0.98; p.ad=0.65;
-f.1=0.8;f.ad=0.8; p.sur=0.8; p.prod=0.77
+n.years=10; n.data=c(50,50,50);init.age = c(100,100); 
+phi.1=0.4; phi.ad=0.76;p.1=0.98; p.ad=0.65;
+p.sur=0.8; p.prod=0.77
 
 n.initiation.dates <- 31
 
@@ -79,6 +79,8 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
   for (i in 1:ti){
     les <- matrix(data = c(PHI1[i]*FL1[i], PHI1[i]*FL2[i], PHI2[i], PHI2[i]), byrow=T, nrow = length(Ni), ncol = length(Ni))
     N[,(i+1)]<-les%*%N[,i]
+    les
+    eigen(les)
   }
   no.anim <- sum(N)
   no.ani <- round(no.anim*2.5)

### 2020-11-22 scripts/dataSims/simulate data.R
updates to simulation study scenarios

@@ -53,17 +53,32 @@ scenarios <- scenarios %>%
   mutate(`MR Included` = ifelse(`MR Included` == "Y",1,0)) %>% 
   mutate(`Abund Included` = ifelse(`Abund Included` == "Y",1,0)) %>% 
   mutate(`Nests Included` = ifelse(`Nests Included` == "Y",1,0)) %>%
-  select(-`Sims per`)
+  select(-`Sims per`) %>%
+  mutate(`Fec` = 1/2 * `Mean Clutch Size` * `Daily nest survival`^30)
+
+lams <- apply(scenarios, 1, function(x) {eigen(matrix(data = c(x[8] * x[11], x[8] * x[11], x[7], x[7]), 
+                                              byrow = TRUE, nrow = 2, ncol = 2))$values[1]})
+
+scenarios <- cbind(scenarios, lams)
+View(scenarios)
+
+# AEB note
+# parameter values picked such that
+# parameter with most uncertainty - juv survival (one we observe most indirectly)
+  # how good at we are recovering that 
 
 # run simulations ######
 
 for (s in 1:n.scenarios) {
   for (i in 1:sims.per) {
     if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 1) {
+      # simulate datasets
       # run full model
     } else if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 0) {
+      # simulate datasets
       # run no nest model
     } else if (scenarios[s, `MR Included`] == 0 & scenarios[s, `Nests Includedd`] == 1) {
+      # simulate datasets
       # run no resight model
     }
     

### 2020-11-22 scripts/dataSims/IPMSimData_nonest.R
some fixes for Hinit hopefully???

@@ -8,9 +8,9 @@
 
 library(nimble)
 
-n.years=10; n.data=c(50,50,50);init.age = c(100,100); 
+n.years=10; n.data=c(50,50,100);init.age = c(100,100); 
 phi.1=0.4; phi.ad=0.76;p.1=0.98; p.ad=0.65;
-p.sur=0.8; p.prod=0.77
+p.prod=0.77; p.sur=0.8
 
 n.initiation.dates <- 31
 
@@ -31,8 +31,7 @@ n.sam<-3
 
 visit.interval <- 3
 
-simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
- f.1, f.ad, p.sur, 
+simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
  n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 
  season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval, n.sam){
   
@@ -43,8 +42,8 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
   phi1<-phi.1
   phi2<-phi.ad
   pjuv<-p.1
-  padults<-p.ad
   psur<-p.sur
+  padults<-p.ad
   n.initiation.dates <- n.initiation.dates
   max.nest.age <- max.nest.age
   first.initiation.date <- first.initiation.date
@@ -381,9 +380,11 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
           }
           
           first.nest[year, i] <- min(which(!is.na(H[year, i, ])))
-          if (is.nest[year, i]) {
+          if (is.nest[year, i] & last.status[year, i]) { # found active and succeeded
             last.nest[year, i] <- max(which(!is.na(H[year, i, ])))
-          } else {
+          } else if (is.nest[year, i] & !last.status[year, i]) { # found active but then failed
+            last.nest[year, i] <- min(which(!is.na(H[year, i, ]) & H[year, i, ] == 0))
+          } else { # found inactive
             last.nest[year, i] <- first.nest[year, i]
           }
             
@@ -403,9 +404,11 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,
           }
           
           first.nest[year, i] <- min(which(!is.na(H[year, i, ])))
-          if (is.nest[year, i]) {
[…]

### 2020-11-22 scripts/dataSims/Non_m_arrayCJSIPMNimble.R

@@ -102,7 +102,7 @@ IPMmod<-nimbleCode({
   #nest success model goes here
   
   # Priors and constraints #####
-  lambdaf ~ dunif(0, 10)
+  lambdaf ~ dunif(1, 5)
   
   #changed this prior, since if it is too low the population goes negative
   #and I gather from looking that this may be reasonable?
@@ -185,8 +185,8 @@ constants<-list(nyears = ncol(m),
 # function for H inits - set every NA to 1
 # ugly but whatever
 Hinits <- H
-Hinits[!is.na(H)] <- NA
-Hinits[is.na(H)] <- 1
+Hinits[which(!is.na(H))] <- NA
+Hinits[which(is.na(H))] <- 1
 
 inits <- list(mean.phi=runif(2,0,1),
               mean.p = runif(1, 0, 1), 
@@ -197,8 +197,8 @@ inits <- list(mean.phi=runif(2,0,1),
               n1.start=10,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
               nad.start=10,#sample(1:30,1),
               phi.nest = runif(1, 0.925, 1), 
-              lambdaf = runif(1, 1, 10),
-              H = Hinits
+              lambdaf = runif(1, 1, 10) #,
+              #H = Hinits
               )
 parameters <- c("mean.phi", 
                 "mean.p", "fec", "N1", "Nad", 
@@ -208,7 +208,7 @@ mod<-nimbleModel(IPMmod, constants=constants, data=datipm, inits=inits)
 conf<-configureMCMC(mod)
 conf$addMonitors(parameters)
 Rmcmc<-buildMCMC(conf)
-Cmodel<-compileNimble(mod)
+Cmodel<-compileNimble(mod, maxContractions = 1000)
 #for 
 Cmodel$setInits(inits)
 #Cmodel$setData(newdata)

### 2020-11-23 scripts/dataSims/IPMSimData_nonest.R
Small stuff

@@ -9,8 +9,8 @@
 library(nimble)
 
 n.years=10; n.data=c(50,50,100);init.age = c(100,100); 
-phi.1=0.4; phi.ad=0.76;p.1=0.98; p.ad=0.65;
-p.prod=0.77; p.sur=0.8
+phi.1=0.3; phi.ad=0.77;p.1=0.98; p.ad=0.3;
+p.sur=0.3
 
 n.initiation.dates <- 31
 
@@ -20,7 +20,7 @@ last.fledge.date <- n.initiation.dates + max.nest.age
 season.length <- last.fledge.date - first.initiation.date + 1 + 2
 
 # mean clutch size
-mean.clutch.size <- 3
+mean.clutch.size <- 2.5
 
 # daily nest survival
 phi.nest <- 0.975
@@ -222,6 +222,7 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
   IND_MR <- IND[,,resamp1]
   IND_Count <- IND[,,resamp2]
   IND_Nest <- IND[,,resamp3]
+  #only use the yoy and adults that are alive
   
   #######Mark resight data
   #pretend marking and resighting:
@@ -344,6 +345,7 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
   
   inds.nests.available <- resamp3
   
+  
   n.found <- length(inds.nests.available) #* prop.nests.found
   
   H <- array(data = NA, dim = c(ti+1, n.found, season.length))
@@ -449,7 +451,7 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
               first.nest = first.nesttidy, last.nest = last.nesttidy, 
               max.nest.age = max.nest.age, 
               n.nests = N.nests.found, n.succ.nests = N.nests.successful, 
-              n.sam = n.sam))
+              n.sam = n.sam, true.fec=true.fec))
 }
 df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
                n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 

### 2020-11-23 scripts/dataSims/Non_m_arrayCJSIPMNimble.R

@@ -102,7 +102,7 @@ IPMmod<-nimbleCode({
   #nest success model goes here
   
   # Priors and constraints #####
-  lambdaf ~ dunif(1, 5)
+  lambdaf ~ dunif(0, 5)
   
   #changed this prior, since if it is too low the population goes negative
   #and I gather from looking that this may be reasonable?
@@ -194,26 +194,26 @@ inits <- list(mean.phi=runif(2,0,1),
               p.surv=runif(1,0,1),
               z=z.state,
               p.surv=runif(1,0,1),
-              n1.start=10,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
-              nad.start=10,#sample(1:30,1),
+              n1.start=100,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
+              nad.start=100,#sample(1:30,1),
               phi.nest = runif(1, 0.925, 1), 
-              lambdaf = runif(1, 1, 10) #,
+              lambdaf = runif(1, 0, 5) #,
               #H = Hinits
               )
 parameters <- c("mean.phi", 
-                "mean.p", "fec", "N1", "Nad", 
-                "Ntot", "lambda", "p.surv", 
+                "mean.p", "fec", 
+                "p.surv", 
                 "phi.nest", "lambdaf")
 mod<-nimbleModel(IPMmod, constants=constants, data=datipm, inits=inits)
 conf<-configureMCMC(mod)
 conf$addMonitors(parameters)
 Rmcmc<-buildMCMC(conf)
-Cmodel<-compileNimble(mod, maxContractions = 1000)
-#for 
+Cmodel<-compileNimble(mod)#, maxContractions = 1000)
+#for re-running without compiling all the above
 Cmodel$setInits(inits)
 #Cmodel$setData(newdata)
 Cmcmc<-compileNimble(Rmcmc, project=Cmodel)
-Cmcmc$run(thin=10, reset=T, niter=100000, nburnin=50000)
+Cmcmc$run(thin=10, reset=T, niter=10000, nburnin=5000)
 out<-as.data.frame(as.matrix(Cmcmc$mvSamples))
 
 # TODO

### 2020-11-23 scripts/combogenerator.R
big repo reorg

@@ -1,20 +1,0 @@
-abundinc <- c("Y", "N")
-MRinc <- c("Y", "N")
-nestinc <- c("Y", "N")
-
-param1 <- c("L", "H")
-param2 <- c("L", "H")
-param3 <- c("L", "H")
-param4 <- c("L", "H")
-param5 <- c("L", "H")
-
-paramcombs <- expand.grid(param1, param2, param3, param4, param5)
-paramcombs <- cbind(paramcombs, paramcombs[, 5])
-
-combs <- expand.grid(abundinc, MRinc, nestinc, param1, param2, param3, param4, param5)
-
-combs <- cbind(combs, combs[, 8])
-
-
-
-write.csv(combs, )

### 2020-11-24 scripts/dataSims/IPMSimData_markasYoYandAd.R
New files for simulation and fitting, with marking YoY and Adults but never chicks

@@ -1,0 +1,476 @@
+#HAS - need to comment throughout, but got the MR data (I think) how it should be
+
+
+#altered IPM sim with placeholder for nest success
+#population projections and count code is from BPA WinBUGS appendix 2
+#code adapted from WinBugsBPA Appendix 2, code to simulate IPM
+# link to code is here: https://www.vogelwarte.ch/assets/files/publications/BPA/Web%20appendix%202.txt
+
+library(nimble)
+
+n.years=10; n.data=c(50,50,100);init.age = c(100,100); 
+phi.1=0.4; phi.ad=0.77; p.ad=0.75;#p.1=0.98;
+p.sur=0.5
+
+n.initiation.dates <- 31
+
+max.nest.age <- 30
+first.initiation.date <- 1
+last.fledge.date <- n.initiation.dates + max.nest.age
+season.length <- last.fledge.date - first.initiation.date + 1 + 2
+
+# mean clutch size
+mean.clutch.size <- 2.5
+
+# daily nest survival
+phi.nest <- 0.975
+
+prop.nests.found <- 0.8
+
+n.sam<-3
+
+visit.interval <- 3
+
+simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
+ n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 
+ season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval, n.sam){
+  
+  ti<-n.years
+  ni<-n.years-1
+  nd<-n.data
+  Ni<-init.age
+  phi1<-phi.1
+  phi2<-phi.ad
+  pjuv<-p.1
+  psur<-p.sur
+  padults<-p.ad
+  n.initiation.dates <- n.initiation.dates
+  max.nest.age <- max.nest.age
+  first.initiation.date <- first.initiation.date
+  last.fledge.date <- n.initiation.dates + max.nest.age
+  season.length <- last.fledge.date - first.initiation.date + 1 + 2
+  mean.clutch.size <- mean.clutch.size
+  phi.nest <- phi.nest
+  prop.nests.found <- prop.nests.found
+  visit.interval <- visit.interval
+  n.sam <- n.sam
+  
+  # DERIVE TRUE FECUNDITY
[…]

### 2020-11-24 documents/DraftReport.Rmd
added equations for count and productivity data
CJS equations
Draft equations and methods text, though still working.

@@
 ---
-title: 'Title'
+title: ''
 output:
@@
 ```{r include = FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)
+#for reference: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
 
-#set default performance for chunks and then modify individually if necessary
-#for reference: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
-# echo == show code
-# warning/message == show warnings/messages
-# results == type of output
-# eval == whether to evaluate code or not, defaults to TRUE
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)
 ```
@@
 #source a plot theme for figures
-source(here::here('scripts', 'PlotTheme.R'))
-
-```
-
-```{r load data or model output, echo = FALSE}
-
-# #load data
-# rain <- data.frame(rain_std = rnorm(100, 10, 10), 
-#                        site = factor(rep(1:5, each = 20)))
-# 
-# #handy objects for summary stats?
-# avg_rain <- rain %>%
-#   group_by(site) %>%
-#   dplyr::summarize(avg_rain = mean(rain_std))
-
-```
-
-```{r create all figures and tables, echo = FALSE, results = 'hide'}
-
-# #Figures - there might be a more eloquent way, but in the past I've found you have to create objects above (figures and tables) so you can reference them in-text and then call them below where you actually want them to appear in the document
-# 
-# plotA <- ggplot(rain, aes(x = site, y = rain_std, col = site, group = site)) +
-#   geom_point() +
-#   xlab('') + ylab('') +
-#   plot_theme()
-# 
-# plotB <- ggplot(rain, aes(x = site, y = rain_std, group = site)) +
-#   geom_boxplot() +
-#   xlab('') + ylab('') +
-#   plot_theme()
-# 
-# plots <- plot_grid(plotA, plotB, labels = "auto", align = 'v', ncol = 1, hjust = 0, label_size = 10)
-# 
-# #use the 'figs' function created above
-# figs("plots", caption = "Scatter and boxplots of rain per site.")
-# 
-# 
-# #create table
-# rain_table <- avg_rain
+# source(here::here('scripts', 'PlotTheme.R'))
 
@@
 
-Authors
+Ideas: 
+- Performance of integrated population models across varying data landscapes  
+- When and how far can we push the integrated population model framework? Estimating demographic parameters under variable data landscapes  
+- Pushing the limits: examining minimal data requirements for estimating demographic parameters in an integrated population model framework
 
-*Affiliations*
-Text
-
-**Abstract**
-
-Don't forget that it's easy to call summaries from the data.
-For example, the sample size for this project is `r dim(rain)[1]`, and the average rainfall at site `r unique(as.numeric(rain[rain$site == 1,'site']))` over the study period is `r round(mean(rain[rain$site == 1,'rain_std']),1)`.
-<!-- AJW: make comments like this -->
+Abby Bratt, Caroline Cappello, Amelia DuVall, Hannah Sipe, Amanda Warlick
 
@@
 
+1. Briefly describe the issue, conclude with teaser sentence about what this project does
+IPMs as an increasingly popular tool that can provide tangible benefits
+Non-ID parameters, improved precision
+Expanded spatio-temporal inference
+These benefits have motivated interesting expansions and applications of this technique
+Examples
+First (?) joint analysis of mark-recovery and census data (Besbeas et al. 2002)
+Application to species (bats) w/ demographic data deficiencies (Schaub et al. 2007)
+Estimating hidden demographic processes (e.g., proportion of female lambs on the island and temporal variation) (Tavecchia et al. 2009)
+“We experimentally evaluated the flexibility and limitations of the integrated approach by omitting portions of the data and comparing the results from these “incomplete” data sets with those obtained from the entire data set and from the analysis of the mark-resighting-recovery data only”
+Inclusion of multivariate observations w/ estimation of transition probabilities within multi-state systems (McCrea et al. 2012)
+Inclusion of environmental stochasticity (Oppel et al. 2014)
+However, this growth in usage has occurred with relatively poor understanding of how to assess goodness of fit, and therefore less is also known about the bounds/limits of the IPM framework to estimate demographic parameters in situations where available data might be sparse.
+Study summary: In this study, we use simulated data to examine IPM performance across a range of scenarios where data informing key demographic parameters may be either sparse or missing.
+
+2. With the expansion of the use of IPMs, few studies have robustly examined IPM performance through data simulation. 
+In those that have used data simulation, the focus has been on:
+Proof of concept (improved precision and reduced bias)
+McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al. 2017, Ahrestani et al. 2017, Bowler et al. 2019
+Model assumptions
+Independence assumption (Abadi et al. 2010, Schaub & Fletcher 2015, Weegman et al. 2020)
+MR assumptions (survival heterogeneity, marker-induced bias; Riecke et al. 2019)
+None/few to date have looked at ability to return parameter values or power to detect trends when survival not informed by data or when a given data source is sparse, biased, or highly uncertain, but this is an interesting, complicated, and important topic
+IPM performance is often context-specific: depends on life history, sample size, effect sizes, likelihood structures, study/inference goals, so what can we learn
+Contributions of different datasets to the likelihood - helps us understand the underlying connections/mechanisms of the IPM, and to use this knowledge to think about IPM performance across life histories, survey designs, etc.
+
+3. The “typical” IPM has gotten a lot of mileage, but can we push the limits?
+The typical (survival and counts) has taught a lot, and been extended
+Examples of a few cool extensions: spatially explicit IPMs, source-sink dynamics, environmental stochasticity
+But wouldn’t it also be cool and open doors if we could apply this approach in situations where this “typical” data wasn’t available or was sparse? 
+Situations where only know fecundity and counts
+Situations where some of our data streams are biased and/or uncertain?
+Best to worst: unbiased/low variance, unbiased/high variance, biased/low variance, biased/high variance
+Sun et al. 2019 (adding sites vs. changing detection)
+Simmonds et al. 2020; Farr et al. 2020 (similar to our goals but for SDMs)
+We can’t confidently go in these directions without refining our understanding of the mechanisms underlying IPMs in these situations and where and when IPMs may not lead to improved precision or where they actually can’t estimate that unidentifiable parameter
+This study fills that knowledge gap by X (restate purpose/approach) with the aim of improving our understanding of this important statistical approach, particularly for passerine bird life history
+
+4. Why this study is useful for practitioners
+Study design: minimizing cost/effort and maximizing usefulness
+Surveys per week, hours per survey, number of sites, etc
+Intentional survey design vs opportunistic/community science program data  
+Sanderlin et al. 2014
+Helping ecologists understand what is possible with available data, particularly given the tendency to maximize what you have 
+Importance of unbiased estimates for conservation and management
+
 **Methods**
 
-*Data simulation and life history*  
-We will simulate data for a hypothetical passerine (or seabird) species that is best described with two age-classes: young-of-year (YOY) and adults.
-We will assume an even sex ratio between males and females, but we will use a female-only model because we assume that the population dynamics of our population are driven by female abundance and demography.
-We assume that the population is surveyed annually just prior to chicks hatching.
-Young and adult survival could be assumed to be the same.
-Adults produce new chicks at rate f, and those chicks survive to be YOY at rate $\phi_1$.
-Current YOY will also breed the following year, and also produce surviving chicks at rate f*phi1$.
-A life cycle diagram for this population is given below:
+*Data simulation: life history*  
+We simulated data for an avian species assuming life history stages and traits of a typical passerine.
+Our simulated hypothetical species has two age-classes: young-of-year (YOY) and adults.
+Each age class has a unique survival rate ($S_1$, $S_{ad}$) and adults produce new chicks at rate *f*, and those chicks survive to be YOY at rate $S_1$.
+Current YOY will also breed the following year to produce surviving chicks at rate $f \times S_1$.
+(Fig 1).
+We assume an annual breeding cycle, with females nesting once per year.
+The period from nest initiation to fledging is 30 days regardless of clutch size.
 
-Our hypothetical bird species nests once per year and we assume all adult females are breeders.
-The population is quite small, so we are able to monitor it over its entire range.
-Therefore we assume no immigration or emigration.
+As is often the convention, we use a female-only model because we assume that the population dynamics of our species are driven largely by female abundance and demography.
+All demographic rates are assumed constant over a 15-year study period.
+The population is quite small and the survey extends across the entire range, therefore we assume no immigration or emigration.
 
 *Model framework and assumptions*
-Productivity: Nests are monitored according to the procedures described in Martin and Geupel (1993).
-In summary, adults are carefully monitored for behavioral clues that indicate nesting behavior, and followed until a nest is found.
-We assume that the nests found are independent and representative of the population.
-After locating a nest, we revisit the nest every 2-3 days to check its status, until it either fledges or fails.
-We assume our observations of the nests are perfect.
-It takes 30 days from nest initiation for chicks to fledge, regardless of clutch size.
-We assume that nests can be correctly aged when first found.
-We also assume the daily survival probability is the same regardless of nest stage (i.e.
-laying, incubating, nestlings).
-We use a daily nest survival model to obtain unbiased estimates for the probability of nest success (i.e.
-at least 1 chick fledges), and a Poisson model to estimate the average number of chicks fledged conditional on a nest being successful.
 
-Equations
+*Count model*  
+We simulated data such that the population is surveyed annually in a 15-year pre-breeding census.
+Count surveys are conducted three times per year in quick succession such that we assume the population is closed between secondary sampling periods.
+We use an N-mixture model to estimate annual abundance from these repeated counts, where we assume no individual is double counted and that individual detections are independent and occur with equal probability.
+The underlying state process represents the latent variables of stage-specific abundance at each occasion:
 
-MR: Chicks are marked in the nest just prior to fledging with a unique combination of colored leg bands such that they can be individually identified over the course of their life.
-Every chick in every nest that fledges is banded.
-We perform 1 exhaustive mark-resight survey per breeding season.
-We will use a Cormack-Jolly-Seber (CJS) model to estimate juvenile and adult survival based on the resulting individual encounter histories.
-In accordance with the CJS model, we assume that each marked individual is independent with the same probability of detection and same probability of survival in and between sampling occasions.
-We also assume that marks are not lost and they are perfectly observed.
+$$N_{1, t+1} \sim Poisson(N_{1,t}, S_{juv,t}, \frac{f_t}{2} + N_{ad,t}S_{ad,t},\frac{f_t}{2})$$  
+$$N_{ad,t+1} \sim Binomial(N_{1,t} + N_{ad,t}, S_{ad,t})$$
 
-Equations
+where the number of juveniles ($N_1$) is estimated using a Poisson distribution with probability based on the sum of stage-specific individuals that survive and reproduce.
+A binomial distribution is used to estimate the number of adults based on the stage-specific survival and abundance in the previous occasion.
 
-Counts: Count surveys are conducted 3 times per year just prior to breeding.
-Since we assume no immigration and emigration, and these surveys are done in quick succession, we assume the population is closed during this period of time.
-We assume no individual is double counted on a survey, and that individual detections are independent and occur with equal probability.
-Therefore we will use an N-mixture model to estimate annual population abundance from this repeated count data.
+The observation process model,
 
-Equations
+$$y_t = (N_{tot, t}) + \epsilon$$
+$$\epsilon \sim Normal(0, \sigma^2_y)$$
 
+describes counts *y* at time *t* given given stage-specific abundance as binomially distributed with sample size $N_t$ and detection probability $p$, with normally distributed observation error $\epsilon$ estimated with mean = 0 with the prior for $sigma_y$ being U(0,10).
+<!-- from amanda: help - prior for observation error?  -->
+
+The overall likelihood for this hierarchical state-space model can be described as the product of the state and observation processe models:
+$$L_{SS}(\text{y}|\text{N}, \text{S_juv}, \text{S_ad}, \text{f}, \sigma^2_y) = L_O(\text{y}|\text{N},\sigma^2_y) \times L_S(\text{N}|\text{S_juv}, \text{S_ad}, \text{f})$$
+
+*Nest success productivity model*
+
+In this simulation, nests are monitored according to the procedures described in Martin and Geupel (1993), where adults are observed for behavioral clues that indicate nesting behavior and followed until a nest is found.
+After locating a nest, it is revisited every 2-3 days to check its status until it either fledges or fails.
+We assume no state uncertainty in our nest observations, that nests can be correctly aged when first found, and that monitored nests are independent and representative of the population.
+We assume the daily survival probability is uniform across nest stages (*i.e.*, laying, incubating, nestlings).
+We estimate daily nest survival to obtain unbiased estimates for the probability of nest success (*i.e.*, at least 1 chick fledges): We use a Poisson model to estimate the average number of chicks fledged conditional on a nest being successful.
+
+$$J_t \sim Poisson(\rho_t) $$
+$$\rho_t = R_tf_t $$
+
+where the number of nestlings counted $J$ in year *t* is modeled using a Poisson distribution with the product of the number of surveyed broods $R_t$ and fecundity $f_t$ at each time *t*.
+Fecundity is estimated using a uniform prior (0,20) and the likelihood of this daily nest survival model can be described as $L_P(\text{J}, \text{R}|\text{f})$.
+
+*Mark-resight model*  
+In this simulation study, YOY chicks are marked in the nest just prior to fledging with a unique combination of colored leg bands such that they can be individually identified over the course of their lives.
+Every chick in every nest that fledges is banded from a single annual comprehensive mark-resight survey per breeding season.
+To estimate stage-specific survival based on the individual encounter histories, we use a Cormack-Jolly-Seber (citation) model where we assume that marks are not lost or misread and that each marked individual has independent and identical survival and detection probabilities across occasions.
+The ecological state process *z* describes the probability that individual *i* at time *t* is alive or dead, namely:
+
+$$ z_{i,t}|z_{i,t-1} \sim bernoulli(z_{i,t-1} S_{1,ad}) $$
+
+where the state at time *t* is modeled with a Bernouli distribution based on the state at the previous occasion and state-specific survival probabilities $S_1$ and $S_{ad}$.
+Similarly, the observation process model
+
+$$ y_{i,t}|z_{i,t} \sim  bernoulli(z_{i,t} p)$$
+
+describes observations *y* using a Bernouli distribution based on the state of the individual in the current occasion and detection probability *p*.
+Mean survival and detection probabilities are modeled on the probability scale using uniform priors, $ S_{1,ad} \sim uniform(0,1)$, $ p \sim uniform(0,1)$, respectively.
+The overall likelihood of the mark-resight model estimating survival probabilities is the product of the state and observation process models: $L_{SS}(\text{y}|\text{z}, \text{S_juv}, \text{S_ad}, \text{p}) = L_O(\text{y}|\text{z},\text{p})     \times L_S(\text{z}|\text{S_juv}, \text{S_ad})$.
+
+*Integrated population model*
+
+To estimate abundance of the overall population $N_{tot}$ at each time *t*, the parameter estimates from the productivity and mark-resight models are combined with the stage-specific abundane estimated using the N-mixture model in a stochastic leslie population matrix model.
+
+$$ \left[\begin{array}
+{l}
+N_{1,t+1} \sim Binomial(N_{1,t}S_{1,t}) \\
+N_{ad,t+1} \sim Binomial(N_{ad,t}S_{ad,t}) \\
+N_{Tot, t+1} = N_{1,t} + N_{ad,t}
+\end{array}\right] 
+$$
+
+The joint likelihood for the integrated population model is the product of the likelihoods for the three subcomponent models.
+In instances when one dataset is removed, informative priors are used in place of the data to inform each of the relevant parameters (Table 1).
 Integration: leslie matrix/pop model; joint likelihood equations
@@
 
-plots
+# plots
 
@@
 
-kable(rain_table, align = 'c', format.args = list(big.mark = ","))
+# kable(rain_table, align = 'c', format.args = list(big.mark = ","))
 

### 2020-11-25 documents/DraftReport.Rmd
Added intro outline, did some methods clean-up.

@@
 1. Briefly describe the issue, conclude with teaser sentence about what this project does
-IPMs as an increasingly popular tool that can provide tangible benefits
-Non-ID parameters, improved precision
-Expanded spatio-temporal inference
-These benefits have motivated interesting expansions and applications of this technique
-Examples
-First (?) joint analysis of mark-recovery and census data (Besbeas et al. 2002)
-Application to species (bats) w/ demographic data deficiencies (Schaub et al. 2007)
-Estimating hidden demographic processes (e.g., proportion of female lambs on the island and temporal variation) (Tavecchia et al. 2009)
-“We experimentally evaluated the flexibility and limitations of the integrated approach by omitting portions of the data and comparing the results from these “incomplete” data sets with those obtained from the entire data set and from the analysis of the mark-resighting-recovery data only”
-Inclusion of multivariate observations w/ estimation of transition probabilities within multi-state systems (McCrea et al. 2012)
-Inclusion of environmental stochasticity (Oppel et al. 2014)
-However, this growth in usage has occurred with relatively poor understanding of how to assess goodness of fit, and therefore less is also known about the bounds/limits of the IPM framework to estimate demographic parameters in situations where available data might be sparse.
-Study summary: In this study, we use simulated data to examine IPM performance across a range of scenarios where data informing key demographic parameters may be either sparse or missing.
+
+- IPMs as an increasingly popular tool that can provide tangible benefits  
+-- Non-ID parameters, improved precision  
+-- Expanded spatio-temporal inference  
+-- These benefits have motivated interesting expansions and applications of this technique  
+--- Examples   
+- Brief history of IPM development  
+-- First (?) joint analysis of mark-recovery and census data (Besbeas et al. 2002)  
+-- Application to species (bats) w/ demographic data deficiencies (Schaub et al. 2007)  
+-- Estimating hidden demographic processes (e.g., proportion of female lambs on the island and temporal variation) (Tavecchia et al. 2009)  
+*“We experimentally evaluated the flexibility and limitations of the integrated approach by omitting portions of the data and comparing the results from these “incomplete” data sets with those obtained from the entire data set and from the analysis of the mark-resighting-recovery data only”*
+-- Inclusion of multivariate observations w/ estimation of transition probabilities within multi-state systems (McCrea et al. 2012)  
+-- Inclusion of environmental stochasticity (Oppel et al. 2014)  
+- However, this growth in usage has occurred with relatively poor understanding of how to assess goodness of fit, and therefore less is also known about the bounds/limits of the IPM framework to estimate demographic parameters in situations where available data might be sparse.
+- Study summary: In this study, we use simulated data to examine IPM performance across a range of scenarios where data informing key demographic parameters may be either sparse or missing.
 
 2. With the expansion of the use of IPMs, few studies have robustly examined IPM performance through data simulation. 
-In those that have used data simulation, the focus has been on:
-Proof of concept (improved precision and reduced bias)
-McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al. 2017, Ahrestani et al. 2017, Bowler et al. 2019
-Model assumptions
-Independence assumption (Abadi et al. 2010, Schaub & Fletcher 2015, Weegman et al. 2020)
-MR assumptions (survival heterogeneity, marker-induced bias; Riecke et al. 2019)
-None/few to date have looked at ability to return parameter values or power to detect trends when survival not informed by data or when a given data source is sparse, biased, or highly uncertain, but this is an interesting, complicated, and important topic
-IPM performance is often context-specific: depends on life history, sample size, effect sizes, likelihood structures, study/inference goals, so what can we learn
-Contributions of different datasets to the likelihood - helps us understand the underlying connections/mechanisms of the IPM, and to use this knowledge to think about IPM performance across life histories, survey designs, etc.
+- In those that have used data simulation, the focus has been on:
+-- Proof of concept (improved precision and reduced bias): McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al. 2017, Ahrestani et al. 2017, Bowler et al. 2019  
+- Model assumptions  
+-- Independence assumption (Abadi et al. 2010, Schaub & Fletcher 2015, Weegman et al. 2020)  
+-- MR assumptions (survival heterogeneity, marker-induced bias; Riecke et al. 2019)  
+- None/few to date have looked at ability to return parameter values or power to detect trends when survival not informed by data or when a given data source is sparse, biased, or highly uncertain, but this is an interesting, complicated, and important topic
+- IPM performance is often context-specific: depends on life history, sample size, effect sizes, likelihood structures, study/inference goals, so what can we learn  
+- Contributions of different datasets to the likelihood - helps us understand the underlying connections/mechanisms of the IPM, and to use this knowledge to think about IPM performance across life histories, survey designs, etc.
 
 3. The “typical” IPM has gotten a lot of mileage, but can we push the limits?  
-The typical (survival and counts) has taught a lot, and been extended
-Examples of a few cool extensions: spatially explicit IPMs, source-sink dynamics, environmental stochasticity
-But wouldn’t it also be cool and open doors if we could apply this approach in situations where this “typical” data wasn’t available or was sparse? 
-Situations where only know fecundity and counts
-Situations where some of our data streams are biased and/or uncertain?
-Best to worst: unbiased/low variance, unbiased/high variance, biased/low variance, biased/high variance
-Sun et al. 2019 (adding sites vs. changing detection)
-Simmonds et al. 2020; Farr et al. 2020 (similar to our goals but for SDMs)
-We can’t confidently go in these directions without refining our understanding of the mechanisms underlying IPMs in these situations and where and when IPMs may not lead to improved precision or where they actually can’t estimate that unidentifiable parameter
-This study fills that knowledge gap by X (restate purpose/approach) with the aim of improving our understanding of this important statistical approach, particularly for passerine bird life history
+- The typical (survival and counts) has taught a lot, and been extended  
+- Examples of a few cool extensions: spatially explicit IPMs, source-sink dynamics, environmental stochasticity
+- But wouldn’t it also be cool and open doors if we could apply this approach in situations where this “typical” data wasn’t available or was sparse?  
+-- Situations where only know fecundity and counts  
+-- Situations where some of our data streams are biased and/or uncertain?  
+-- Best to worst: unbiased/low variance, unbiased/high variance, biased/low variance, biased/high variance
+--- Sun et al. 2019 (adding sites vs. changing detection)  
+--- Simmonds et al. 2020; Farr et al. 2020 (similar to our goals but for SDMs)  
+- We can’t confidently go in these directions without refining our understanding of the mechanisms underlying IPMs in these situations and where and when IPMs may not lead to improved precision or where they actually can’t estimate that unidentifiable parameter.
+This study fills that knowledge gap by X (restate purpose/approach) with the aim of improving our understanding of this important statistical approach, particularly for passerine bird life history.
 
 4. Why this study is useful for practitioners    
-Study design: minimizing cost/effort and maximizing usefulness
-Surveys per week, hours per survey, number of sites, etc
-Intentional survey design vs opportunistic/community science program data  
-Sanderlin et al. 2014
-Helping ecologists understand what is possible with available data, particularly given the tendency to maximize what you have 
-Importance of unbiased estimates for conservation and management
+- Study design: minimizing cost/effort and maximizing usefulness
+- Surveys per week, hours per survey, number of sites, etc  
+- Intentional survey design vs opportunistic/community science program data: Sanderlin et al. 2014
+- Helping ecologists understand what is possible with available data, particularly given the tendency to maximize what you have 
+- Importance of unbiased estimates for conservation and management
 
@@
 Each age class has a unique survival rate ($S_1$, $S_{ad}$) and adults produce new chicks at rate *f*, and those chicks survive to be YOY at rate $S_1$.
-Current YOY will also breed the following year to produce surviving chicks at rate $f \times S_1$.
-(Fig 1).
+Current YOY will also breed the following year to produce surviving chicks at rate $f \times S_1$ (Fig 1).
 We assume an annual breeding cycle, with females nesting once per year.
@@
 Count surveys are conducted three times per year in quick succession such that we assume the population is closed between secondary sampling periods.
-We use an N-mixture model to estimate annual abundance from these repeated counts, where we assume no individual is double counted and that individual detections are independent and occur with equal probability.
+We use an N-mixture model (citations) to estimate annual abundance from these repeated counts, where we assume no individual is double counted and that individual detections are independent and occur with equal probability.
 The underlying state process represents the latent variables of stage-specific abundance at each occasion:
 
-$$N_{1, t+1} \sim Poisson(N_{1,t}, S_{juv,t}, \frac{f_t}{2} + N_{ad,t}S_{ad,t},\frac{f_t}{2})$$  
-$$N_{ad,t+1} \sim Binomial(N_{1,t} + N_{ad,t}, S_{ad,t})$$
+$$N_{1, t+1} \sim \text{Poisson}(N_{1,t}, S_{1,t}, \frac{f_t}{2} + N_{ad,t}S_{ad,t},\frac{f_t}{2})$$
 
-where the number of juveniles ($N_1$) is estimated using a Poisson distribution with probability based on the sum of stage-specific individuals that survive and reproduce.
+$$N_{ad,t+1} \sim \text{Binomial}(N_{1,t} + N_{ad,t}, S_{ad,t})$$
+
+where the number of juveniles ($N_1$) is estimated using a Poisson distribution with probability based on the sum of stage-specific individuals that survive ($S_{1,ad}$) and reproduce ($f_t$).
 A binomial distribution is used to estimate the number of adults based on the stage-specific survival and abundance in the previous occasion.
@@
 
-$$y_t = (N_{tot, t}) + \epsilon$$
-$$\epsilon \sim Normal(0, \sigma^2_y)$$
+$$y_t = N_{tot, t} + \epsilon$$
+$$\epsilon \sim \text{Normal}(0, \sigma^2_y)$$
 
-describes counts *y* at time *t* given given stage-specific abundance as binomially distributed with sample size $N_t$ and detection probability $p$, with normally distributed observation error $\epsilon$ estimated with mean = 0 with the prior for $sigma_y$ being U(0,10).
+describes counts *y* at time *t* given given stage-specific abundance as binomially distributed with sample size $N_t$ and detection probability $p$, with normally distributed observation error $\epsilon$ estimated with mean = 0 with the prior for $\sigma_y$ being U(0,10).
 <!-- from amanda: help - prior for observation error?  -->
 
-The overall likelihood for this hierarchical state-space model can be described as the product of the state and observation processe models:
-$$L_{SS}(\text{y}|\text{N}, \text{S_juv}, \text{S_ad}, \text{f}, \sigma^2_y) = L_O(\text{y}|\text{N},\sigma^2_y) \times L_S(\text{N}|\text{S_juv}, \text{S_ad}, \text{f})$$
+The overall likelihood for this hierarchical state-space model can be described as the product of the state and observation process models:
+$$L_{SS}(\text{y}|\text{N}, S_1, S_{ad}, f, \sigma^2_y) = L_O(\text{y}|\text{N},\sigma^2_y) \times L_S(\text{N}|S_1, S_{ad}, f)$$
 
@@
 
-$$J_t \sim Poisson(\rho_t) $$
+$$J_t \sim \text{Poisson}(\rho_t) $$
 $$\rho_t = R_tf_t $$
 
-where the number of nestlings counted $J$ in year *t* is modeled using a Poisson distribution with the product of the number of surveyed broods $R_t$ and fecundity $f_t$ at each time *t*.
-Fecundity is estimated using a uniform prior (0,20) and the likelihood of this daily nest survival model can be described as $L_P(\text{J}, \text{R}|\text{f})$.
+where the number of nestlings counted $J$ in year *t* is modeled using a Poisson distribution with the product of the number of surveyed broods $R_t$ and fecundity $f_t$ at each occasion.
+Fecundity is estimated using a uniform prior (0, 20) and the likelihood of this daily nest survival model can be described as $L_P(\text{J}, \text{R}|\text{f})$.
 
@@
 
-$$ z_{i,t}|z_{i,t-1} \sim bernoulli(z_{i,t-1} S_{1,ad}) $$
+$$ z_{i,t}|z_{i,t-1} \sim \text{Bernoulli}(z_{i,t-1} S_{1,ad}) $$
 
@@
 
-$$ y_{i,t}|z_{i,t} \sim  bernoulli(z_{i,t} p)$$
+$$ y_{i,t}|z_{i,t} \sim  \text{Bernoulli}(z_{i,t} p)$$
 
 describes observations *y* using a Bernouli distribution based on the state of the individual in the current occasion and detection probability *p*.
-Mean survival and detection probabilities are modeled on the probability scale using uniform priors, $ S_{1,ad} \sim uniform(0,1)$, $ p \sim uniform(0,1)$, respectively.
-The overall likelihood of the mark-resight model estimating survival probabilities is the product of the state and observation process models: $L_{SS}(\text{y}|\text{z}, \text{S_juv}, \text{S_ad}, \text{p}) = L_O(\text{y}|\text{z},\text{p})     \times L_S(\text{z}|\text{S_juv}, \text{S_ad})$.
+Mean survival and detection probabilities are modeled on the probability scale using uniform priors, $S_{1,ad} \sim \text{Uniform}(0,1)$ and $p \sim \text{Uniform}(0,1)$, respectively.
+The overall likelihood of the mark-resight model estimating survival probabilities is the product of the state and observation process models: $L_{SS}(\text{y}|\text{z}, ~S_1, ~S_{ad}, ~\text{p}) = L_O(\text{y}|\text{z},~\text{p})     \times L_S(\text{z}|S_1,~S_{ad})$.
 
@@
 
-$$ \left[\begin{array}
-{l}
-N_{1,t+1} \sim Binomial(N_{1,t}S_{1,t}) \\
-N_{ad,t+1} \sim Binomial(N_{ad,t}S_{ad,t}) \\
-N_{Tot, t+1} = N_{1,t} + N_{ad,t}
-\end{array}\right] 
-$$
+<!-- From amanda: help: not sure if we need something like this since we have the equations above? -->
+<!-- $$ \left[\begin{array} -->
+<!-- {l} -->
+<!-- N_{1,t+1} \sim \text{Binomial}(N_{1,t},~S_{1,t}) \\ -->
+<!-- N_{ad,t+1} \sim \text{Binomial}(N_{ad,t},~S_{ad,t}) \\ -->
+<!-- N_{Tot, t+1} = N_{1,t} + N_{ad,t} -->
+<!-- \end{array}\right]  -->
+<!-- $$  -->
 
@@
 In instances when one dataset is removed, informative priors are used in place of the data to inform each of the relevant parameters (Table 1).
-Integration: leslie matrix/pop model; joint likelihood equations
+
+[IPM joint likelihood equation]
 
@@
 
+Review project aim and main findings
+
+Elaborate on interesting finding 1  
+- Implications of not exploring this in an application?  
+- Future areas of work?  
+- Extension to more complex life history? Data scenarios?
+
+Elaborate on interesting finding 2
+- Implications of not exploring this in an application?  
+- Future areas of work?  
+- Extension to more complex life history? Data scenarios?
+
+Broaden to implications of work, study design, existing IPMs
+
 \newpage
@@
 **Literature cited**
+
+Abadi, F., Gimenez, O., Arlettaz, R., Schaub, M., 2010.
+An assessment of integrated population models: bias, accuracy, and violation of the assumption of independence.
+Ecology 91, 7–14.
+https://doi.org/10.1890/08-2235.1
+
+Ahrestani, F.S., Saracco, J.F., Sauer, J.R., Pardieck, K.L., Royle, J.A., 2017.
+An integrated population model for bird monitoring in North America.
+Ecol Appl 27, 916–924.
+https://doi.org/10.1002/eap.1493
+
+Bled, F., Belant, J.L., Daele, L.J.V., Svoboda, N., Gustine, D., Hilderbrand, G., Barnes, V.G., n.d.
+Using multiple data types and integrated population models to improve our knowledge of apex predator population dynamics.
+Ecology and Evolution 7, 9531–9543.
+https://doi.org/10.1002/ece3.3469
+
+Bowler, D.E., Nilsen, E.B., Bischof, R., O’Hara, R.B., Yu, T.T., Oo, T., Aung, M., Linnell, J.D.C., 2019.
+Integrating data from different survey types for population monitoring of an endangered species: the case of the Eld’s deer.
+Scientific Reports 9, 7766.
+https://doi.org/10.1038/s41598-019-44075-9
+
+Chandler, R.B., Clark, J.D., 2014. Spatially explicit integrated population models. Methods in Ecology and Evolution 5, 1351–1360. https://doi.org/10.1111/2041-210X.12153
+
+Farr, M.T., Green, D.S., Holekamp, K.E., Zipkin, E.F., n.d. Integrating distance sampling and presence-only data to estimate species abundance. Ecology n/a, e03204. https://doi.org/10.1002/ecy.3204
+
+McCrea, R.S., Morgan, B.J.T., Gimenez, O., Besbeas, P., Lebreton, J.-D., Bregnballe, T., 2010.
+Multi-Site Integrated Population Modelling.
+Journal of Agricultural, Biological and Environmental Statistics 15, 539–561.
+https://doi.org/10.1007/s13253-010-0027-5
+
+Riecke, T.V., Williams, P.J., Behnke, T.L., Gibson, D., Leach, A.G., Sedinger, B.S., Street, P.A., Sedinger, J.S., 2019.
+Integrated population models: model assumptions and inference.
+Methods Ecol Evol 2041–210X.13195.
+https://doi.org/10.1111/2041-210X.13195
+
+Schaub, M., Fletcher, D., 2015. Estimating immigration using a Bayesian integrated population model: choice of 
+parametrization and priors. Environ Ecol Stat 22, 535–549. https://doi.org/10.1007/s10651-015-0309-8
+
+Simmonds, E.G., Jarvis, S.G., Henrys, P.A., Isaac, N.J.B., O’Hara, R.B., 2020.
+Is more data always better?
+A simulation study of benefits and limitations of integrated distribution models.
+Ecography 43, 1413–1422.
+https://doi.org/10.1111/ecog.05146
+
+Sun, C.C., Royle, J.A., Fuller, A.K., 2019. Incorporating citizen science data in spatially explicit integrated population models. Ecology 100, e02777. https://doi.org/10.1002/ecy.2777
+
+Tavecchia, G., Besbeas, P., Coulson, T., Morgan, B.J.T., Clutton‐Brock, T.H., 2009.
+Estimating Population Size and Hidden Demographic Parameters with State‐Space Modeling.
+The American Naturalist 173, 722–733.
+https://doi.org/10.1086/598499
+
+Weegman, M.D., Arnold, T.W., Clark, R.G., Schaub, M., 2020.
+Partial and complete dependency among data sets has minimal consequence on estimates from integrated population models.
+Ecological Applications.
+https://doi.org/10.1002/eap.2258
 
@@
 
-`r figs("plots", caption = "Scatter and boxplots of rain per site.")`
+`r figs("plots", caption = "Plots.")`
 
@@
 
-`r tbls("rain_table", caption = "Table of rain by site.")`
+`r tbls("rain_table", caption = "Table of X, Y, Z.")`
 ```{r}
 
-# kable(rain_table, align = 'c', format.args = list(big.mark = ","))
+# kable(table, align = 'c', format.args = list(big.mark = ","))
 

### 2020-11-30 scripts/current version/IPMNimble_markedasYoYandAD.R
some changes but still doesn't work :(

@@ -4,11 +4,14 @@
 #IPM model for sim data
 library(here)
 #source(here("scripts", "dataSims","BPA_IPMsimData.R")) #full on BPA WinBUGS code
-source(here("scripts", "dataSims","IPMsimData_nonest.R")) #changed the MR data
+source(here("scripts", "current version","IPMsimData_markasYoYandAd.R")) #changed the MR data
+source(here("scripts", "current version","productivityDataSim.R"))
 
 # 11.3. Example of a simple IPM (counts, capture-recapture, reproduction)
 # 11.3.1. Load data
 # Population counts (from years 1 to n.years)
+
+# AEB NOTE - if 1000 individuals, why are the counts so low?
 y <- df$SUR
 n.sam <- df$n.sam
 
@@ -23,13 +26,13 @@ age_ch <- df$age_ch
 #J<-nestlings#df$nestlings
 #R<-R#df$R
  
-H <- df$H
-Fledged <- df$Fledged
-first.nest <- df$first.nest
-last.nest <- df$last.nest
-max.nest.age <- df$max.nest.age
-n.nests <- df$n.nests
-n.succ.nests <- df$n.succ.nests
+H <- prod$observed.nest.status
+Fledged <- prod$clutch.sizes
+first.nest <- prod$first.nest
+last.nest <- prod$last.nest
+max.nest.age <- prod$max.nest.age
+n.nests <- prod$N.nests.found
+n.succ.nests <- prod$N.nests.successful
 
 library("nimble")
 IPMmod<-nimbleCode({
@@ -57,12 +60,12 @@ IPMmod<-nimbleCode({
     f[t] <- fec
   }
   #mean.fec ~ dunif(0, 20)
-  
+
   # 3.1. Likelihood for population population count data (state-space model)
   # 3.1.1 System process
   for (t in 2:nyears){
     #I think this was causing the problem, it was a little mis-specified
-    #Double check that this is right, but it doesnt cause an error in 
+    #Double check that this is right, but it doesnt cause an error in
     #Slice samplers this way
     #mean1[t]<-((f[t-1]*N1[t-1])+(f[t-1]*Nad[t-1]))*mean.phi[1]
     #for reference, this is how it was:
@@ -73,16 +76,16 @@ IPMmod<-nimbleCode({
[…]

### 2020-11-30 scripts/current version/IPMSimData_markasYoYandAd.R

@@ -1,0 +1,285 @@
+#HAS - need to comment throughout, but got the MR data (I think) how it should be
+
+
+#altered IPM sim with placeholder for nest success
+#population projections and count code is from BPA WinBUGS appendix 2
+#code adapted from WinBugsBPA Appendix 2, code to simulate IPM
+# link to code is here: https://www.vogelwarte.ch/assets/files/publications/BPA/Web%20appendix%202.txt
+
+library(nimble)
+
+n.years=10; n.data=c(1000,200);init.age = c(1000,1000); 
+phi.1=0.4; phi.ad=0.77; p.ad=0.75;p.1=0.98;
+p.sur=0.5
+
+
+max.nest.age <- 30
+
+# mean clutch size
+mean.clutch.size <- 2.5
+
+# daily nest survival
+phi.nest <- 0.975
+
+n.sam <- 3
+
+simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
+ max.nest.age, mean.clutch.size, phi.nest, n.sam){
+  
+  ti<-n.years
+  ni<-n.years-1
+  nd<-n.data
+  Ni<-init.age
+  phi1<-phi.1
+  phi2<-phi.ad
+  pjuv<-p.1
+  psur<-p.sur
+  padults<-p.ad
+  max.nest.age <- max.nest.age
+  mean.clutch.size <- mean.clutch.size
+  phi.nest <- phi.nest
+  n.sam <- n.sam
+  
+  # DERIVE TRUE FECUNDITY
+  true.fec <- 1/2 * mean.clutch.size * phi.nest^max.nest.age
+  
+  # Combine the values to vectors
+  PHI1 <- rep(phi1, ti)
+  PHI2 <- rep(phi2, ti)
+  PJUV <- rep(pjuv, ti)
+  PADULTS <- rep(padults, ti)
+  PSUR <- rep(psur, ti)
+  FL1 <- rep(true.fec, ti+1) # THIS MATCHES WITH PROD DATA SIM
+  FL2 <- rep(true.fec, ti+1) # THIS MATCHES WITH PROD DATA SIM
+  
+  ###########################
+  # Population simulation
+  ##########################
+  
+  # 1. Matrix model to estimate roughly the population size with a Leslie matrix
+  
+  N <- matrix(data = NA, nrow = length(Ni), ncol = ti+1)
+  N[,1] <- Ni
[…]

### 2020-11-30 scripts/current version/productivityDataSim.R

@@ -13,8 +13,14 @@ library(nimble)
 
 # parameters ######
 
+# TODO
+# add structure for multiple years!!!!!
+
+getNestDat <- function(nyears = 10, phi = 0.975, mean.clutch.size = 2.5) {
+  
+
 # set season length and stuff
-n.initiation.dates <- 120
+n.initiation.dates <- 31
 max.nest.age <- 30
 first.initiation.date <- 1
 last.fledge.date <- n.initiation.dates + max.nest.age
@@ -22,7 +28,8 @@ last.fledge.date <- n.initiation.dates + max.nest.age
 season.length <- last.fledge.date - first.initiation.date + 1 + 2
 
 # nests for population
-N.nests.total <- 100
+# AEB note - this is currently a crude way to do this
+N.nests.total <- 100 * nyears # year does not matter because there is no year effect
 prop.nests.found <- 0.8
 
 # TODO
@@ -32,11 +39,11 @@ prop.nests.found <- 0.8
 # so we get a random number every year
 
 # mean clutch size
-mean.clutch.size <- 5
+#mean.clutch.size <- 5
 # AEB note - additional assumption that all eggs hatch and all nestlings fledge
 
 # daily nest survival
-phi <- 0.975
+#phi <- 0.975
 
 # true fecundity (female chicks produced per nest)
 # assuming 50/50 sex ratio at hatching
@@ -46,6 +53,9 @@ true.fec <- 1/2 * mean.clutch.size * phi^max.nest.age
 visit.interval <- 3
 
 # start data sim ####
+
+# TODO 
+# change to arrays
 
 # all nests - this is latent
 total.nests.age <- matrix(NA, nrow = N.nests.total, ncol = season.length)
@@ -136,4 +146,12 @@ last[-which.successful] <- apply(observed.nest.status[-which.successful, ], 1, f
 # there should not be zeroes in here
 clutch.sizes <- rpois(N.nests.successful, mean.clutch.size)
 
-
+return(list(observed.nest.status = observed.nest.status, 
+            N.nests.found = N.nests.found,
+            clutch.sizes = clutch.sizes,
+            N.nests.successful = N.nests.successful, 
+            first.nest = first, 
+            last.nest = last, 
+            max.nest.age = max.nest.age))
+}
+prod <- getNestDat()

### 2020-11-30 scripts/dataSims/IPMSimData_markasYoYandAd.R

@@ -1,476 +1,0 @@
-#HAS - need to comment throughout, but got the MR data (I think) how it should be
-
-
-#altered IPM sim with placeholder for nest success
-#population projections and count code is from BPA WinBUGS appendix 2
-#code adapted from WinBugsBPA Appendix 2, code to simulate IPM
-# link to code is here: https://www.vogelwarte.ch/assets/files/publications/BPA/Web%20appendix%202.txt
-
-library(nimble)
-
-n.years=10; n.data=c(50,50,100);init.age = c(100,100); 
-phi.1=0.4; phi.ad=0.77; p.ad=0.75;#p.1=0.98;
-p.sur=0.5
-
-n.initiation.dates <- 31
-
-max.nest.age <- 30
-first.initiation.date <- 1
-last.fledge.date <- n.initiation.dates + max.nest.age
-season.length <- last.fledge.date - first.initiation.date + 1 + 2
-
-# mean clutch size
-mean.clutch.size <- 2.5
-
-# daily nest survival
-phi.nest <- 0.975
-
-prop.nests.found <- 0.8
-
-n.sam<-3
-
-visit.interval <- 3
-
-simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
- n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 
- season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval, n.sam){
-  
-  ti<-n.years
-  ni<-n.years-1
-  nd<-n.data
-  Ni<-init.age
-  phi1<-phi.1
-  phi2<-phi.ad
-  pjuv<-p.1
-  psur<-p.sur
-  padults<-p.ad
-  n.initiation.dates <- n.initiation.dates
-  max.nest.age <- max.nest.age
-  first.initiation.date <- first.initiation.date
-  last.fledge.date <- n.initiation.dates + max.nest.age
-  season.length <- last.fledge.date - first.initiation.date + 1 + 2
-  mean.clutch.size <- mean.clutch.size
-  phi.nest <- phi.nest
-  prop.nests.found <- prop.nests.found
-  visit.interval <- visit.interval
-  n.sam <- n.sam
-  
-  # DERIVE TRUE FECUNDITY
[…]

### 2020-11-30 scripts/current version/IPMNimble_markedasYoYandAD.R
think i fixed this

@@ -76,7 +76,7 @@ IPMmod<-nimbleCode({
     Nad[t] ~ dbin(mean.phi[2],(Ntot[t-1]))
   }
   for (t in 1:nyears){
-    Ntot[t] <- Nad[t] + N1[t]
+    Ntot[t] <- round(Nad[t] + N1[t])
   }
 
   # 3.1.2 Observation process
@@ -190,22 +190,22 @@ Hinits <- H
 Hinits[!is.na(H)] <- NA
 Hinits[is.na(H)] <- 1
 
-inits <- list(mean.phi=runif(2,0,1),
-              mean.p = runif(1, 0, 1), 
+inits <- list(mean.phi=c(0.4, 0.77),
+              mean.p = 0.5, 
               #mean.fec = runif(1, 0, 10), 
-              p.surv=runif(1,0,1),
+              p.surv=0.5,
               z=z.state,
-              p.surv=runif(1,0,1),
-              n1.start=10,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
-              nad.start=10,#sample(1:30,1),
-              phi.nest = runif(1, 0.925, 1), 
-              lambdaf = runif(1, 0, 5) ,
+              n1.start=1000,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
+              nad.start=1000,#sample(1:30,1),
+              phi.nest = 0.975, 
+              lambdaf = 2.5 ,
               H = Hinits
               )
 parameters <- c("mean.phi", 
                 "mean.p", "fec", 
                 "p.surv", 
-                "phi.nest", "lambdaf")
+                "phi.nest", "lambdaf", 
+                "N1", "Nad")
 mod<-nimbleModel(IPMmod, constants=constants, data=datipm, inits=inits)
 conf<-configureMCMC(mod)
 conf$addMonitors(parameters)
@@ -221,6 +221,3 @@ out<-as.data.frame(as.matrix(Cmcmc$mvSamples))
 # TODO
 # slice sampler reached maximum number of contractions for Ns :(
 
-# choke points
-# data sim takes some time
-# compile nimble takes some time

### 2020-11-30 scripts/current version/IPMSimData_markasYoYandAd.R

@@ -279,7 +279,8 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
   
   #Get all the data together, just MR and Survey Counts until nest in here
   return(list(ch=ch, SUR=SUR, age_ch=age_ch, first=first, last=last,
-              n.sam = n.sam, true.fec=true.fec))
+              n.sam = n.sam, true.fec=true.fec, 
+              N = N))
 }
 df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
                max.nest.age, mean.clutch.size, phi.nest, n.sam)  

### 2020-11-30 scripts/current version/productivityDataSim.R
updates to simulation running code

@@ -154,4 +154,3 @@ return(list(observed.nest.status = observed.nest.status,
             last.nest = last, 
             max.nest.age = max.nest.age))
 }
-prod <- getNestDat()

### 2020-11-30 scripts/current version/run_scenarios.R

@@ -7,19 +7,25 @@ library(lubridate)
 library(hms)
 library(stringr)
 library(tidyr)
+library(nimble)
 
 # source files ######
 
-# source data simulation function
-# source model function
+# source data simulation functions
+source(here("scripts", "current version","IPMsimData_markasYoYandAd.R")) #changed the MR data
+source(here("scripts", "current version","productivityDataSim.R"))
+# source model functions
+source(here("scripts", "current version","IPMNimble_markedasYoYandAD.R")) #changed the MR data
+# source initial value functions
+source(here("scripts", "current version","IPMinitvalues.R")) #changed the MR data
 
 # load scenarios ####
 
 # site
-paramlevels <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx"), range = "E1:K4")
+paramlevels <- read_excel(here("data", "IPM sim spreadsheet.xlsx"), range = "E1:K4")
 paramlevels[, 1] <- c("L", "M", "H")
 colnames(paramlevels)[1] <- "Level"
-scenarios <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx"), range = "A5:K22")
+scenarios <- read_excel(here("data", "IPM sim spreadsheet.xlsx"), range = "A5:K22")
 scenarios <- scenarios[-c(6, 12), ]
 
 n.scenarios <- max(scenarios$`Scenario Number`)
@@ -69,30 +75,223 @@ View(scenarios)
 
 # run simulations ######
 
+# THINGS THAT NEVER CHANGE
+phi.nest <- 0.975
+mean.clutch.size <- 2.5
+n.sam <- 3
+max.nest.age <- 30
+n.years=10
+n.data=c(1000,200)
+init.age = c(1000,1000)
+phi.ad = 0.77
+p.1 = 0.98
+
+# THIS IS FOR COMPLETE IPM
+
 for (s in 1:n.scenarios) {
+  # THINGS THAT DO CHANGE
+  phi.1 = scenarios[s, "Juv Surv"]
+  p.ad = scenarios[s, "MR detection"]
+  p.sur = scenarios[s, "Abund detection"]
   for (i in 1:sims.per) {
-    if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 1) {
+    if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Included`] == 1) {
       # simulate datasets
[…]

### 2020-11-30 scripts/current version/IPMNimble_markedasYoYandAD.R
updates to simulation running code

@@ -1,38 +1,8 @@
-# TODO
-# abby put nest stuff in here
-
-#IPM model for sim data
-library(here)
-#source(here("scripts", "dataSims","BPA_IPMsimData.R")) #full on BPA WinBUGS code
-source(here("scripts", "current version","IPMsimData_markasYoYandAd.R")) #changed the MR data
-source(here("scripts", "current version","productivityDataSim.R"))
-
 # 11.3. Example of a simple IPM (counts, capture-recapture, reproduction)
 # 11.3.1. Load data
 # Population counts (from years 1 to n.years)
 
-# AEB NOTE - if 1000 individuals, why are the counts so low?
-y <- df$SUR
-n.sam <- df$n.sam
-
-# Capture-recapture data (in m-array format, from years 1 to n.years)
-m <- df$ch
-first <- df$first
-age_ch <- df$age_ch
-
-#Productivity/nest success not in the data sim right now
-#these will need to be changed once we have the nest success data
-# Productivity data (from years 1 to n.years-1)
-#J<-nestlings#df$nestlings
-#R<-R#df$R
- 
-H <- prod$observed.nest.status
-Fledged <- prod$clutch.sizes
-first.nest <- prod$first.nest
-last.nest <- prod$last.nest
-max.nest.age <- prod$max.nest.age
-n.nests <- prod$N.nests.found
-n.succ.nests <- prod$N.nests.successful
+###### FULL IPM ######
 
 library("nimble")
 IPMmod<-nimbleCode({
@@ -128,96 +98,173 @@ IPMmod<-nimbleCode({
   
 })
 
+###### NO NEST ######
 
-
-
-#Provide known state as data  
-state.data <- function(EH){
-  state <- EH
-  for (i in 1:dim(EH)[1]){
-    first <- min(which(EH[i,]==1))
-    last <- max(which(EH[i,]==1))
-    state[i,first:last] <- 1
+# create model versions that omit some of the data
+noNests<-nimbleCode({
+  #-------------------------------------------------
+  #  Integrated population model
+  #  - Age structured model with 2 age classes: 
+  #		1-year old and adult (at least 2 years old)
+  #  - Age at first breeding = 1 year
+  #  - Prebreeding census, female-based
+  #  - All vital rates assumed to be constant
[…]

### 2020-11-30 scripts/current version/IPMSimData_markasYoYandAd.R

@@ -8,21 +8,6 @@
 
 library(nimble)
 
-n.years=10; n.data=c(1000,200);init.age = c(1000,1000); 
-phi.1=0.4; phi.ad=0.77; p.ad=0.75;p.1=0.98;
-p.sur=0.5
-
-
-max.nest.age <- 30
-
-# mean clutch size
-mean.clutch.size <- 2.5
-
-# daily nest survival
-phi.nest <- 0.975
-
-n.sam <- 3
-
 simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
  max.nest.age, mean.clutch.size, phi.nest, n.sam){
   
@@ -282,5 +267,3 @@ simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
               n.sam = n.sam, true.fec=true.fec, 
               N = N))
 }
-df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
-               max.nest.age, mean.clutch.size, phi.nest, n.sam)  

### 2020-11-30 scripts/current version/IPMinitvalues.R

@@ -1,0 +1,41 @@
+#Provide known state as data  
+state.data <- function(EH){
+  state <- EH
+  for (i in 1:dim(EH)[1]){
+    first <- min(which(EH[i,]==1))
+    last <- max(which(EH[i,]==1))
+    state[i,first:last] <- 1
+  }
+  state[state==0] <- NA
+  return(state)
+}
+
+#age for age specific survival probabilities
+ageunknown<-function(age_ch){
+  allages<-age_ch
+  f1<-l1<-numeric(dim(allages)[1])
+  t<-dim(age_ch)[2]
+  for(i in 1:dim(allages)[1]){
+    f1[i]<-min(which(allages[i,]>=2))
+    l1[i]<-max(which(allages[i,]>0))
+    if((f1[i]+1)<=t){
+      allages[i,(f1[i]+1)]<-3
+    } else{}
+    if((f1[i]+2)<=t){
+      allages[i,(f1[i]+2):t]<-3
+    } else{}
+  }
+  allages<-allages-1
+  allages[which(allages==-1)]<-0
+  #allages[which(allages==1)]<-0
+  # allages[which(allages==2)]<-1
+  # allages[which(allages==3)]<-2
+  return(allages)
+}
+
+getHinits <- function(H) {
+  Hinits <- H
+  Hinits[!is.na(H)] <- NA
+  Hinits[is.na(H)] <- 1
+  return(Hinits)
+}

### 2020-11-30 scripts/current version/run_scenarios.R

@@ -7,19 +7,25 @@ library(lubridate)
 library(hms)
 library(stringr)
 library(tidyr)
+library(nimble)
 
 # source files ######
 
-# source data simulation function
-# source model function
+# source data simulation functions
+source(here("scripts", "current version","IPMsimData_markasYoYandAd.R")) #changed the MR data
+source(here("scripts", "current version","productivityDataSim.R"))
+# source model functions
+source(here("scripts", "current version","IPMNimble_markedasYoYandAD.R")) #changed the MR data
+# source initial value functions
+source(here("scripts", "current version","IPMinitvalues.R")) #changed the MR data
 
 # load scenarios ####
 
 # site
-paramlevels <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx"), range = "E1:K4")
+paramlevels <- read_excel(here("data", "IPM sim spreadsheet.xlsx"), range = "E1:K4")
 paramlevels[, 1] <- c("L", "M", "H")
 colnames(paramlevels)[1] <- "Level"
-scenarios <- read_excel(here("scripts", "dataSims", "IPM sim spreadsheet.xlsx"), range = "A5:K22")
+scenarios <- read_excel(here("data", "IPM sim spreadsheet.xlsx"), range = "A5:K22")
 scenarios <- scenarios[-c(6, 12), ]
 
 n.scenarios <- max(scenarios$`Scenario Number`)
@@ -69,30 +75,223 @@ View(scenarios)
 
 # run simulations ######
 
+# THINGS THAT NEVER CHANGE
+phi.nest <- 0.975
+mean.clutch.size <- 2.5
+n.sam <- 3
+max.nest.age <- 30
+n.years=10
+n.data=c(1000,200)
+init.age = c(1000,1000)
+phi.ad = 0.77
+p.1 = 0.98
+
+# THIS IS FOR COMPLETE IPM
+
 for (s in 1:n.scenarios) {
+  # THINGS THAT DO CHANGE
+  phi.1 = scenarios[s, "Juv Surv"]
+  p.ad = scenarios[s, "MR detection"]
+  p.sur = scenarios[s, "Abund detection"]
   for (i in 1:sims.per) {
     if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 1) {
       # simulate datasets
+      df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
[…]

### 2020-11-30 scripts/current version/run_scenarios.R
fix typo

@@ -94,7 +94,7 @@ for (s in 1:n.scenarios) {
   p.ad = scenarios[s, "MR detection"]
   p.sur = scenarios[s, "Abund detection"]
   for (i in 1:sims.per) {
-    if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 1) {
+    if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Included`] == 1) {
       # simulate datasets
       df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
                      max.nest.age, mean.clutch.size, phi.nest, n.sam) 
@@ -152,7 +152,7 @@ for (s in 1:n.scenarios) {
       CmcmcIPM$run(thin=10, reset=T, niter=10000, nburnin=5000)
       out<-as.data.frame(as.matrix(CmcmcIPM$mvSamples))
       
-    } else if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Includedd`] == 0) {
+    } else if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Included`] == 0) {
       # simulate datasets
       df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
                      max.nest.age, mean.clutch.size, phi.nest, n.sam) 
@@ -211,7 +211,7 @@ for (s in 1:n.scenarios) {
       out<-as.data.frame(as.matrix(Cmcmcnonest$mvSamples))
       
       
-    } else if (scenarios[s, `MR Included`] == 0 & scenarios[s, `Nests Includedd`] == 1) {
+    } else if (scenarios[s, `MR Included`] == 0 & scenarios[s, `Nests Included`] == 1) {
       # simulate datasets
       df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
                      max.nest.age, mean.clutch.size, phi.nest, n.sam) 

### 2020-12-01 scripts/current version/IPMNimble_markedasYoYandAD.R
fixed initial population size error

@@ -17,8 +17,10 @@ IPMmod<-nimbleCode({
   
   #population projection part
   # Initial population sizes
-  N1[1]<-n1.start
-  Nad[1]<-nad.start
+  n1.start ~ dunif(0, 1000)
+  nad.start ~ dunif(0, 1000)
+  N1[1] <- n1.start
+  Nad[1] <- nad.start
   # Population growth rate
   for (t in 1:(nyears-1)){
     lambda[t] <- Ntot[t+1] / Ntot[t]
@@ -113,8 +115,10 @@ noNests<-nimbleCode({
   
   #population projection part
   # Initial population sizes
-  N1[1]<-n1.start
-  Nad[1]<-nad.start
+  n1.start ~ dunif(0, 1000)
+  nad.start ~ dunif(0, 1000)
+  N1[1] <- n1.start
+  Nad[1] <- nad.start
   # Population growth rate
   for (t in 1:(nyears-1)){
     lambda[t] <- Ntot[t+1] / Ntot[t]
@@ -197,8 +201,10 @@ noMR<-nimbleCode({
   
   #population projection part
   # Initial population sizes
-  N1[1]<-n1.start
-  Nad[1]<-nad.start
+  n1.start ~ dunif(0, 1000)
+  nad.start ~ dunif(0, 1000)
+  N1[1] <- n1.start
+  Nad[1] <- nad.start
   # Population growth rate
   for (t in 1:(nyears-1)){
     lambda[t] <- Ntot[t+1] / Ntot[t]

### 2020-12-01 scripts/current version/run_scenarios.R

@@ -81,7 +81,7 @@ mean.clutch.size <- 2.5
 n.sam <- 3
 max.nest.age <- 30
 n.years=10
-n.data=c(1000,200)
+n.data=c(200,1000)
 init.age = c(1000,1000)
 phi.ad = 0.77
 p.1 = 0.98
@@ -132,17 +132,18 @@ for (s in 1:n.scenarios) {
       inits <- list(mean.phi=c(0.4, 0.77),
                     mean.p = 0.5, 
                     #mean.fec = runif(1, 0, 10), 
-                    p.surv=0.5,
+                    p.surv=0.9,
                     z=z.state,
-                    n1.start=1000,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
-                    nad.start=1000,#sample(1:30,1),
+                    n1.start=round(mean(y[,1]) * 1.5),#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
+                    nad.start=round(mean(y[,1]) * 1.5),#sample(1:30,1),
                     phi.nest = 0.975, 
                     lambdaf = 2.5 ,
                     H = Hinits
       )
       # run full model
       # IPM
-      parameters <- c("mean.phi", "mean.p", "fec", "p.surv", "phi.nest", "lambdaf", "N1", "Nad")
+      # AEB note - weird stuff happening with p.surv
+      parameters <- c("mean.phi", "mean.p", "fec", "p.surv", "phi.nest", "lambdaf", "lambda")
       modIPM<-nimbleModel(IPMmod, constants=constants, data=datipm, inits=inits)
       confIPM<-configureMCMC(modIPM) 
       confIPM$addMonitors(parameters)
@@ -190,17 +191,17 @@ for (s in 1:n.scenarios) {
       inits <- list(mean.phi=c(0.4, 0.77),
                     mean.p = 0.5, 
                     #mean.fec = runif(1, 0, 10), 
-                    p.surv=0.5,
+                    p.surv=0.9,
                     z=z.state,
-                    n1.start=1000,#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
-                    nad.start=1000,#sample(1:30,1),
+                    n1.start=round(mean(y[,1]) * 1.5),#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
+                    nad.start=round(mean(y[,1]) * 1.5),#sample(1:30,1),
[…]

### 2020-12-01 scripts/current version/run_scenarios.R
add easier tracking

@@ -86,14 +86,13 @@ init.age = c(1000,1000)
 phi.ad = 0.77
 p.1 = 0.98
 
-# THIS IS FOR COMPLETE IPM
-
 for (s in 1:n.scenarios) {
   # THINGS THAT DO CHANGE
   phi.1 = scenarios[s, "Juv Surv"]
   p.ad = scenarios[s, "MR detection"]
   p.sur = scenarios[s, "Abund detection"]
   for (i in 1:sims.per) {
+    print(paste("scenario: ", s, "; simulation number: ", i, sep = ""))
     if (scenarios[s, `MR Included`] == 1 & scenarios[s, `Nests Included`] == 1) {
       # simulate datasets
       df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,

### 2020-12-01 scripts/current version/run_scenarios.R
created R object for scenarios to run

@@ -21,52 +21,54 @@ source(here("scripts", "current version","IPMinitvalues.R")) #changed the MR dat
 
 # load scenarios ####
 
-# site
-paramlevels <- read_excel(here("data", "IPM sim spreadsheet.xlsx"), range = "E1:K4")
-paramlevels[, 1] <- c("L", "M", "H")
-colnames(paramlevels)[1] <- "Level"
-scenarios <- read_excel(here("data", "IPM sim spreadsheet.xlsx"), range = "A5:K22")
-scenarios <- scenarios[-c(6, 12), ]
-
-n.scenarios <- max(scenarios$`Scenario Number`)
-sims.per <- max(scenarios$`Sims per`)
-
-scenarios <- scenarios %>% 
-  mutate(`MR detection` = case_when(`MR detection` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "MR detection"]),
-                                    `MR detection` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "MR detection"]),
-                                    `MR detection` == "H" ~ as.numeric(paramlevels[paramlevels$Level == "H", "MR detection"]))
-         ) %>% 
-  mutate(`Abund detection` = case_when(`Abund detection` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "Abund detection"]),
-                                    `Abund detection` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "Abund detection"]),
-                                    `Abund detection` == "H" ~ as.numeric(paramlevels[paramlevels$Level == "H", "Abund detection"]))
-  ) %>% 
-  mutate(`Adult Surv` = case_when(`Adult Surv` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "Adult Surv"]),
-                                    `Adult Surv` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "Adult Surv"]),
-                                    `Adult Surv` == "H" ~ as.numeric(paramlevels[paramlevels$Level == "H", "Adult Surv"]))
-  ) %>% 
-  mutate(`Juv Surv` = case_when(`Juv Surv` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "Juv Surv"]),
-                                    `Juv Surv` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "Juv Surv"]),
[…]

### 2020-12-01 scripts/current version/run_scenarios.R
fixes for running on ursus

@@ -2,10 +2,10 @@
 
 library(here)
 library(dplyr)
-library(readxl)
-library(lubridate)
-library(hms)
-library(stringr)
+#library(readxl)
+#library(lubridate)
+#library(hms)
+#library(stringr)
 library(tidyr)
 library(nimble)
 
@@ -28,8 +28,6 @@ source(here("scripts", "current version","IPMinitvalues.R")) #changed the MR dat
 # scenarios <- read_excel(here("data", "IPM sim spreadsheet.xlsx"), range = "A5:K22")
 # scenarios <- scenarios[-c(6, 12), ]
 # 
-# n.scenarios <- max(scenarios$`Scenario Number`)
-# sims.per <- max(scenarios$`Sims per`)
 # 
 # scenarios <- scenarios %>% 
 #   mutate(`MR detection` = case_when(`MR detection` == "L" ~ as.numeric(paramlevels[paramlevels$Level == "L", "MR detection"]),
@@ -56,9 +54,9 @@ source(here("scripts", "current version","IPMinitvalues.R")) #changed the MR dat
 #                                     `Daily nest survival` == "M" ~ as.numeric(paramlevels[paramlevels$Level == "M", "Daily nest survival"]),
 #                                     `Daily nest survival` == "H" ~ as.numeric(paramlevels[paramlevels$Level == "H", "Daily nest survival"]))
 #   ) %>% 
-#   mutate(`MR Included` = ifelse(`MR Included` == "Y",1,0)) %>% 
+#   mutate("MR Included" = ifelse("MR Included" == "Y",1,0)) %>% 
 #   mutate(`Abund Included` = ifelse(`Abund Included` == "Y",1,0)) %>% 
-#   mutate(`Nests Included` = ifelse(`Nests Included` == "Y",1,0)) %>%
+#   mutate("Nests Included" = ifelse("Nests Included" == "Y",1,0)) %>%
 #   select(-`Sims per`) %>%
 #   mutate(`Fec` = 1/2 * `Mean Clutch Size` * `Daily nest survival`^30)
 # 
@@ -69,6 +67,9 @@ source(here("scripts", "current version","IPMinitvalues.R")) #changed the MR dat
 # #View(scenarios)
 # saveRDS(scenarios, here("data", "scenarios.Rdata"))
 scenarios <- readRDS(here("data", "scenarios.Rdata"))
+
[…]

### 2020-12-01 documents/DraftReport.Rmd
added draft exploratory figure for posterior medians

@@
 # source(here::here('scripts', 'PlotTheme.R'))
+
+```
+
+```{r load results, results = 'hide', message = FALSE}
+
+test <- readRDS(here::here('data', 'out1_1.Rdata'))
+
+#for investigating, particularly scenarios where survivals are low
+sc <- c(9)
+niter <- 25
+
+test <- data.frame()
+for (s in sc) {
+  for (i in 1:niter) {
+
+temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
+temp.med <- data.frame(t(apply(temp, 2, median)))
+temp.med$iter <- i
+temp.med$scenario <- s
+test <- rbind(test, temp.med)
+  }
+}
+
+#posterior medians for fec, mean.p, mean.phi[1:2], p.surv, phi.nest
+sc <- c(1:5, 11:15)
+niter <- 25
+
+meds <- data.frame()
+for (s in sc) {
+  for (i in 1:niter) {
+
+temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
+temp.med <- data.frame(t(apply(temp, 2, median)))
+temp.med$iter <- i
+temp.med$scenario <- s
+
+meds <- rbind(meds, temp.med)
+  }
+}
+
+sc <- c(6:10)
+niter <- 25
+
+meds2 <- data.frame()
+for (s in sc) {
+  for (i in 1:niter) {
+
+temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
+temp.med <- data.frame(t(apply(temp, 2, median)))
+temp.med$iter <- i
+temp.med$scenario <- s
+
+meds2 <- rbind(meds2, temp.med)
+  }
+}
+
+meds <- bind_rows(meds, meds2)
+
+post.meds <- meds %>% transform(scenario = factor(scenario)) %>%
+  reshape2::melt(id.vars = c('scenario', 'iter')) %>%
+    filter(variable %in% c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest'))
+
+true_vals <- data.frame(scenario = rep(1:15, times = 6),
+                        variable = rep(c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest'), 
+                                       each = 15),
+                        value = c(rep(c(0.8, 0.3, 0.8, 0.3, 0.5), times = 3), #adult phi
+                                  rep(c(0.7, 0.2, 0.7, 0.2, 0.4), times = 3), #juv phi
+                                  c(0.8, 0.8, 0.3, 0.3, 0.5, rep(NA, 5), c(0.8, 0.8, 0.3, 0.3, 0.5)), #mean.p
+                                  rep(NA, 15), #fec: what goes here?
+                                  rep(c(0.85, 0.85, 0.3, 0.3, 0.5), times = 3), #p.surv
+                                  rep(c(0.995, 0.975, 0.995, 0.975, 0.99), times = 3))) #phi.nest
+
+ggplot(post.meds, aes(y = value, x = scenario, group = scenario)) +
+  geom_boxplot() +
+  geom_point(size = 0.7, col = 'darkgrey') + 
+  geom_point(data = true_vals, aes(y = value, x = scenario, group = scenario), col = 'darkred') +
+  xlab('Scenario') + ylab('Posterior medians') +
+  facet_wrap(~variable, scales = 'free_y') +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90))
 

### 2020-12-01 scripts/current version/IPMNimble_markedasYoYandAD.R

@@ -209,7 +209,7 @@ noMR<-nimbleCode({
   for (t in 1:(nyears-1)){
     lambda[t] <- Ntot[t+1] / Ntot[t]
   }
-  #prior for survey detectin probability
+  #prior for survey detection probability
   p.surv~dunif(0,1)
   #  productivity
   for (t in 1:(nyears-1)){

### 2020-12-02 documents/DraftReport.Rmd
posterior median figure

--- 2020-11-25 documents/DraftReport.Rmd
@@
 # source(here::here('scripts', 'PlotTheme.R'))
+
+```
+
+```{r load results, results = 'hide', message = FALSE}
+
+# test <- readRDS(here::here('data', 'out1_1.Rdata'))
+
+#true vals
+truth <- readRDS(here::here('data', 'scenarios.Rdata')) 
+colnames(truth) <- gsub(' ', '', colnames(truth))
+true.vals <- truth %>%
+  dplyr::select(ScenarioNumber, MRdetection, Abunddetection, AdultSurv, JuvSurv, Dailynestsurvival, Fec) %>%
+  dplyr::rename(scenario = ScenarioNumber) %>%
+  reshape2::melt(id.vars = c('scenario'), value.name = 'Truth') %>%
+  transform(variable = factor(variable,
+                              levels = c('AdultSurv', 'JuvSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'),
+                              labels = c('AdultSurv', 'JuvSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival')))
+
+#posterior medians for fec, mean.p, mean.phi[1:2], p.surv, phi.nest
+
+#read in scenarios separately without mean.p below
+sc <- c(1:5, 11:15)
+niter <- 25
+
+meds <- data.frame()
+for (s in sc) {
+  for (i in 1:niter) {
+
+temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
+temp.med <- data.frame(t(apply(temp, 2, median)))
+temp.med$iter <- i
+temp.med$scenario <- s
+
+meds <- rbind(meds, temp.med)
+  }
+}
+
+sc <- c(6:10)
+niter <- 25
+
+meds2 <- data.frame()
+for (s in sc) {
+  for (i in 1:niter) {
+
+temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
+temp.med <- data.frame(t(apply(temp, 2, median)))
+temp.med$iter <- i
+temp.med$scenario <- s
+
+meds2 <- rbind(meds2, temp.med)
+  }
+}
+
+meds <- bind_rows(meds, meds2)
+
+post.meds <- meds %>% transform(scenario = factor(scenario)) %>%
+  reshape2::melt(id.vars = c('scenario', 'iter')) %>%
+  filter(variable %in% c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest')) %>%
+  transform(variable = factor(variable,
+                              levels = c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest'),
+                              labels = c('JuvSurv', 'AdultSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival')))
+
+ggplot(post.meds, aes(y = value, x = scenario, group = scenario)) +
+  geom_boxplot() +
+  geom_point(size = 0.7, col = 'darkgrey') + 
+  geom_point(data = true.vals, aes(y = Truth, x = scenario, group = scenario), col = 'darkred') +
+  xlab('Scenario') + ylab('Posterior medians') +
+  facet_wrap(~variable, scales = 'free_y') +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90))
 

### 2020-12-02 documents/DraftReport.Rmd
bias figures

--- 2020-11-25 documents/DraftReport.Rmd
@@
 #for reference: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
+
+cols <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")
 
@@
 # source(here::here('scripts', 'PlotTheme.R'))
+
+```
+
+```{r load results, results = 'hide', message = FALSE}
+
+# test <- readRDS(here::here('data', 'out1_1.Rdata'))
+
+#true vals
+truth <- readRDS(here::here('data', 'scenarios.Rdata')) 
+colnames(truth) <- gsub(' ', '', colnames(truth))
+true.vals <- truth %>%
+  dplyr::select(ScenarioNumber, MRdetection, Abunddetection, AdultSurv, JuvSurv, Dailynestsurvival, Fec) %>%
+  dplyr::rename(scenario = ScenarioNumber) %>%
+  reshape2::melt(id.vars = c('scenario'), value.name = 'Truth') %>%
+  transform(variable = factor(variable,
+                              levels = c('AdultSurv', 'JuvSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'),
+                              labels = c('AdultSurv', 'JuvSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'))) %>%
+  transform(Data = ifelse(scenario < 6, 'All', 
+                          ifelse(scenario < 11 & scenario > 5, 'Counts + Productivity', 'Counts + MR'))) %>%
+  transform(Truth = ifelse(Data == 'Counts + Productivity' & variable == 'MRdetection', NA, Truth))
+
+#posterior medians for fec, mean.p, mean.phi[1:2], p.surv, phi.nest
+
+#read in scenarios separately without mean.p below
+sc <- c(1:5, 11:15)
+niter <- 25
+
+meds <- data.frame()
+for (s in sc) {
+  for (i in 1:niter) {
+
+temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
+temp.med <- data.frame(t(apply(temp, 2, median)))
+temp.med$iter <- i
+temp.med$scenario <- s
+
+meds <- rbind(meds, temp.med)
+  }
+}
+
+sc <- c(6:10)
+niter <- 25
+
+meds2 <- data.frame()
+for (s in sc) {
+  for (i in 1:niter) {
+
+temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
+temp.med <- data.frame(t(apply(temp, 2, median)))
+temp.med$iter <- i
+temp.med$scenario <- s
+
+meds2 <- rbind(meds2, temp.med)
+  }
+}
+
+meds <- bind_rows(meds, meds2)
+
+post.meds <- meds %>% #transform(scenario = factor(scenario)) %>%
+  reshape2::melt(id.vars = c('scenario', 'iter')) %>%
+  filter(variable %in% c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest')) %>%
+  transform(variable = factor(variable,
+                              levels = c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest'),
+                              labels = c('JuvSurv', 'AdultSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'))) %>%
+  transform(Data = ifelse(scenario < 6, 'All', 
+                          ifelse(scenario < 11 & scenario > 5, 'Counts + Productivity', 'Counts + MR')))
+
+ggplot(post.meds, aes(y = value, x = scenario, group = scenario)) +
+  geom_boxplot(aes(color = Data)) +
+  geom_point(size = 0.7, col = 'darkgrey') + 
+  geom_point(data = true.vals, aes(y = Truth, x = scenario, group = scenario), col = 'darkred') +
+  xlab('Scenario') + ylab('Posterior medians') +
+  facet_wrap(~variable, scales = 'free_y') +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+        legend.position = 'top') +
+  scale_x_continuous(breaks = 1:15) +
+  scale_color_manual(values = cols[c(2,3,5)])
+
+#relative bias
+
+bias.vals <- post.meds %>% 
+  merge(true.vals %>% dplyr::select(scenario, variable, Truth), by = c('scenario', 'variable')) %>%
+  transform(bias = value-Truth, rel_bias = (value-Truth)/Truth)
+
+ggplot(bias.vals, aes(y = rel_bias, x = scenario, group = scenario)) +
+  geom_boxplot(aes(color = Data)) +
+  geom_point(size = 0.7, col = 'darkgrey') + 
+  geom_hline(aes(yintercept = 0), linetype = 'dashed', col = 'darkred') +
+  xlab('Scenario') + ylab('Relative bias') +
+  facet_wrap(~variable, scales = 'free_y') +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+        legend.position = 'top') +
+  scale_x_continuous(breaks = 1:15) +
+  scale_color_manual(values = cols[c(2,3,5)])
 

### 2020-12-03 scripts/current version/run_scenarios.R
updates for running multiple chains

@@ -101,181 +101,184 @@ for (s in 1:1) {
       df<-simIPMdata(n.years, n.data, init.age, phi.1, phi.ad, p.1, p.ad, p.sur,
                      max.nest.age, mean.clutch.size, phi.nest, n.sam) 
       prod <- getNestDat()
-      # abund data
-      y <- df$SUR
-      n.sam <- df$n.sam
-      # Capture-recapture data (in m-array format, from years 1 to n.years)
-      m <- df$ch
-      first <- df$first
-      age_ch <- df$age_ch
-      # Nest data
-      H <- prod$observed.nest.status
-      Fledged <- prod$clutch.sizes
-      first.nest <- prod$first.nest
-      last.nest <- prod$last.nest
-      max.nest.age <- prod$max.nest.age
-      n.nests <- prod$N.nests.found
-      n.succ.nests <- prod$N.nests.successful
-      # initial values
-      age<-ageunknown(age_ch)
-      z.state <- state.data(m)
-      Hinits <- getHinits(H)
-      
-      datipm <- list(ch.y = m, y = y, 
-                     H = H, 
-                     Fledged = Fledged)
-      constants<-list(nyears = ncol(m), 
-                      n.ind=nrow(m), first=first, age=age, n.sam=n.sam,
-                      n.nests = n.nests, 
-                      n.succ.nests = n.succ.nests, 
-                      first.nest = first.nest, 
-                      last.nest = last.nest, 
-                      max.nest.age = max.nest.age)
-      inits <- list(mean.phi=c(0.4, 0.77),
-                    mean.p = 0.5, 
-                    #mean.fec = runif(1, 0, 10), 
-                    p.surv=0.9,
-                    z=z.state,
-                    n1.start=round(mean(y[,1]) * 1.5),#sample(1:30,1),#super sensitive to these values, tried rpois(1,30) and it dodnt work
-                    nad.start=round(mean(y[,1]) * 1.5),#sample(1:30,1),
-                    phi.nest = 0.975, 
-                    lambdaf = 2.5 ,
-                    H = Hinits
-      )
-      # run full model
-      # IPM
-      # AEB note - weird stuff happening with p.surv
-      parameters <- c("mean.phi", "mean.p", "fec", "p.surv", "phi.nest", "lambdaf", "lambda")
-      modIPM<-nimbleModel(IPMmod, constants=constants, data=datipm, inits=inits)
[…]

### 2020-12-03 documents/DraftReport.Rmd
Added introduction

@@
 
-1. Briefly describe the issue, conclude with teaser sentence about what this project does
+Understanding the processes that control population dynamics in wild populations is a critical component of conservation but is frequently constrained by missing demographic information, uncertainty in data, and temporally or spatially misaligned datasets (citation).
+Integrated population models (IPMs) are an increasingly popular tool in ecology that offer release from these limitations by providing a framework to combine disparate datasets in a single, unified analysis (Schaub & Abadi 2011).
+This approach can provide several advantages like estimation of otherwise non-identifiable parameters, improved precision in parameter estimation, and [spatio-temporal inference] (citation).
+These advances have practical benefits for resource managers that must make decisions based on limited available information.
+As IPMs become more widely used, however, it is critical that they are robustly evaluated.
 
-- IPMs as an increasingly popular tool that can provide tangible benefits  
--- Non-ID parameters, improved precision  
--- Expanded spatio-temporal inference  
--- These benefits have motivated interesting expansions and applications of this technique  
---- Examples   
-- Brief history of IPM development  
--- First (?) joint analysis of mark-recovery and census data (Besbeas et al. 2002)  
--- Application to species (bats) w/ demographic data deficiencies (Schaub et al. 2007)  
--- Estimating hidden demographic processes (e.g., proportion of female lambs on the island and temporal variation) (Tavecchia et al. 2009)  
-*“We experimentally evaluated the flexibility and limitations of the integrated approach by omitting portions of the data and comparing the results from these “incomplete” data sets with those obtained from the entire data set and from the analysis of the mark-resighting-recovery data only”*
--- Inclusion of multivariate observations w/ estimation of transition probabilities within multi-state systems (McCrea et al. 2012)  
--- Inclusion of environmental stochasticity (Oppel et al. 2014)  
-- However, this growth in usage has occurred with relatively poor understanding of how to assess goodness of fit, and therefore less is also known about the bounds/limits of the IPM framework to estimate demographic parameters in situations where available data might be sparse.
-- Study summary: In this study, we use simulated data to examine IPM performance across a range of scenarios where data informing key demographic parameters may be either sparse or missing.
+Since the introduction of IPMs to ecology in the early 2000s and following the work of Besbeas and colleagues with passerines (Besbeas et al.
+2002, 2005), the common approach has been to estimate parameters by combining mark-recapture and abundance data into a joint likelihood.
+Researchers have has since expanded the use of IPMs to include a variety of taxa (e.g.
+in bats Rhinolophus ferrumequinum: Schaub et al.
+2007, in Svalbard reindeer Rangifer tarandus platyrhynchus: Lee et al.
+2015, Antarctic toothfish Dissostichus mawsoni: Mormede et al.
+2019, wild boar Sus scrofa: Gamelon et al.
+2019), [include multivariate observations with estimation of transition probabilities within multi-state and multi-site systems] (McCrea et al.
+2012), evaluate management efforts (Cruz et al.
+2018), and to include environmental processes and stochasticity (Oppel et al.
+2014, Woodworth et al.
+2017).
+This growth in usage has occurred despite a relatively poor understanding of how to assess goodness of fit however, and we lack knowledge of the limits of the IPM framework in situations where available data are sparse, of poor quality, or when certain data sources are absent.
+For example, many field studies may have survey and reproductive data while missing survival estimates from mark-recapture datasets.
 
-2. With the expansion of the use of IPMs, few studies have robustly examined IPM performance through data simulation. 
-- In those that have used data simulation, the focus has been on:
--- Proof of concept (improved precision and reduced bias): McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al. 2017, Ahrestani et al. 2017, Bowler et al. 2019  
-- Model assumptions  
--- Independence assumption (Abadi et al. 2010, Schaub & Fletcher 2015, Weegman et al. 2020)  
--- MR assumptions (survival heterogeneity, marker-induced bias; Riecke et al. 2019)  
-- None/few to date have looked at ability to return parameter values or power to detect trends when survival not informed by data or when a given data source is sparse, biased, or highly uncertain, but this is an interesting, complicated, and important topic
-- IPM performance is often context-specific: depends on life history, sample size, effect sizes, likelihood structures, study/inference goals, so what can we learn  
-- Contributions of different datasets to the likelihood - helps us understand the underlying connections/mechanisms of the IPM, and to use this knowledge to think about IPM performance across life histories, survey designs, etc.
+Data simulation studies provide an avenue for robust evaluation of the IPM framework, but few such studies exist.
+Those that have used data simulation largely focus on proof of 1) concept (i.e.
+improved precision and reduced bias McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al.
+2017, Ahrestani et al.
+2017, Bowler et al.
+2019), and 2) model assumptions such as independence of datasets (Abadi et al.
+2010, Schaub & Fletcher 2015, Weegman et al.
+2020).
+Riecke et al.
+(2019) found that parameter estimates from IPMs were highly sensitive to violations of mark-recapture model assumptions (i.e.
+marker-induced bias and survival heterogeneity).
+These findings further motivate studies to determine the reliability of IPMs in parameter estimation and in detecting trends when survival is not informed by data or when a data source is sparse, biased, or highly uncertain.
 
-3. The “typical” IPM has gotten a lot of mileage, but can we push the limits?  
-- The typical (survival and counts) has taught a lot, and been extended  
-- Examples of a few cool extensions: spatially explicit IPMs, source-sink dynamics, environmental stochasticity
-- But wouldn’t it also be cool and open doors if we could apply this approach in situations where this “typical” data wasn’t available or was sparse?  
--- Situations where only know fecundity and counts  
--- Situations where some of our data streams are biased and/or uncertain?  
--- Best to worst: unbiased/low variance, unbiased/high variance, biased/low variance, biased/high variance
---- Sun et al. 2019 (adding sites vs. changing detection)  
---- Simmonds et al. 2020; Farr et al. 2020 (similar to our goals but for SDMs)  
-- We can’t confidently go in these directions without refining our understanding of the mechanisms underlying IPMs in these situations and where and when IPMs may not lead to improved precision or where they actually can’t estimate that unidentifiable parameter.
-This study fills that knowledge gap by X (restate purpose/approach) with the aim of improving our understanding of this important statistical approach, particularly for passerine bird life history.
-
-4. Why this study is useful for practitioners    
-- Study design: minimizing cost/effort and maximizing usefulness
-- Surveys per week, hours per survey, number of sites, etc  
-- Intentional survey design vs opportunistic/community science program data: Sanderlin et al. 2014
-- Helping ecologists understand what is possible with available data, particularly given the tendency to maximize what you have 
-- Importance of unbiased estimates for conservation and management
+We aim to improve our understanding of this important statistical approach, particularly for passerine bird life history.
+In this study, we simulate data to examine IPM performance across a range of scenarios where data informing key demographic parameters are sparse or missing.
+We simulate three datasets for a for a theoretical passerine: 1) census counts, 2) reproductive output, and 3) mark-recapture (MR) of banded individuals.
+We then systematically vary the uncertainty in our abundance and MR data and run the integrated models under scenarios where demographic parameters like survival and fecundity are high or low.
+By identifying scenarios where IPMs do not lead to improved precision or cannot estimate unidentifiable parameters, we refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool.
 

### 2020-12-03 scripts/current version/run_scenarios.R
save every chain

@@ -89,7 +89,7 @@ init.age = c(1000,1000)
 phi.ad = 0.77
 p.1 = 0.98
 
-for (s in 1:1) {
+for (s in 15:15) {
   # THINGS THAT DO CHANGE
   phi.1 = scenarios[s, "Juv Surv"]
   p.ad = scenarios[s, "MR detection"]
@@ -155,6 +155,12 @@ for (s in 1:1) {
         CmcmcIPM<-compileNimble(RmcmcIPM, project=CmodelIPM)
         CmcmcIPM$run(thin=10, reset=T, niter=45000, nburnin=5000)
         out<-as.data.frame(as.matrix(CmcmcIPM$mvSamples))
+        
+        # save results to file and to environment
+        outcopy <- out
+        assign(paste("out", s, "_",  i, "-", c, sep = ""), outcopy)
+        saveRDS(out, here("data", paste("out", s, "_",  i, "-", c, ".Rdata", sep = "")))
+        rm(out, outcopy)
       }
     } else if (scenarios[s, "MR Included"] == 1 & scenarios[s, "Nests Included"] == 0) {
       # simulate datasets
@@ -214,6 +220,12 @@ for (s in 1:1) {
         Cmcmcnonest<-compileNimble(Rmcmcnonest, project=Cmodelnonest)
         Cmcmcnonest$run(thin=10, reset=T, niter=10000, nburnin=5000)
         out<-as.data.frame(as.matrix(Cmcmcnonest$mvSamples))
+        
+        # save results to file and to environment
+        outcopy <- out
+        assign(paste("out", s, "_",  i, "-", c, sep = ""), outcopy)
+        saveRDS(out, here("data", paste("out", s, "_",  i, "-", c, ".Rdata", sep = "")))
+        rm(out, outcopy)
       }
     } else if (scenarios[s, "MR Included"] == 0 & scenarios[s, "Nests Included"] == 1) {
       # simulate datasets
@@ -272,14 +284,14 @@ for (s in 1:1) {
         CmcmcnoMR<-compileNimble(RmcmcnoMR, project=CmodelnoMR)
         CmcmcnoMR$run(thin=10, reset=T, niter=10000, nburnin=5000)
         out<-as.data.frame(as.matrix(CmcmcnoMR$mvSamples))
+        
+        # save results to file and to environment
+        outcopy <- out
+        assign(paste("out", s, "_",  i, "-", c, sep = ""), outcopy)
[…]

### 2020-12-03 documents/DraftReport.Rmd
to-dos

@@
 ---
+
+To-do's
+
+Presentation
+- caroline/amelia
+
+Write-up
+- intro (amanda re-reads and checks for possible additions)
+- methods/equation check: change YOTY to juv (hannah)
+- results (abby/hannah)
+
+- discussion (TBD)
+-- summarize high-level finding
+
+-- why we think we see what we see
+--- crux of the issue.... 
+(1) how the life cycle model relies on different data sources
+(2) parameter identifiability
+(3) ?
+
+-- limitations/caveats of this particular study (what can/can't we make inference to/about)
+--- life histories
+
+-- future directions
+--- degrading quantity/quality
+
+-- importance of the work
+(1) understanding IPMs better
+(2) understanding weaknesses of IPMs: when does the integration almost hurt instead of help?
+(3) survey design
+
+Figures/analysis/etc
+- posterior densities (amanda)
+- life cycle/survey timing diagram (amelia)
+- figure captions/printing (amanda)
+- 3 dags (amelia)
+
+Code review
+- annotate data simulation (amanda! caroline! amelia!)
+
+Displaying results
+- posterior densities
+- table (true values, data included, posterior mean, 95% CI, mean bias, mean RMSE)
 

### 2020-12-03 documents/DraftReport.Rmd
to-dos

@@
 - methods/equation check: change YOTY to juv (hannah)
+- simulation write-up (amelia)
 - results (abby/hannah)

### 2020-12-04 scripts/current version/IPMNimble_markedasYoYandAD.R
should use zero truncated poisson (because nest success is defined as fledging at least 1 chick)

@@ -90,7 +90,8 @@ IPMmod<-nimbleCode({
   } # n
   
   for (c in 1:n.succ.nests) {
-    Fledged[c] ~ dpois(lambdaf) 
+    # AEB note - now using zero truncated poisson
+    Fledged[c] ~ T(dpois(lambdaf), 1, Inf)
   } # c
 
   # Derived quantities #####

### 2020-12-05 scripts/current version/IPMNimble_markedasYoYandAD.R
many an oops

@@ -90,8 +90,9 @@ IPMmod<-nimbleCode({
   } # n
   
   for (c in 1:n.succ.nests) {
-    # AEB note - now using zero truncated poisson
-    Fledged[c] ~ T(dpois(lambdaf), 1, Inf)
+    # AEB note - should be using zero truncated poisson
+    #Fledged[c] ~ T(dpois(lambdaf), 1, Inf)
+    Fledged[c] ~ dpois(lambdaf)
   } # c
 
   # Derived quantities #####

### 2020-12-05 scripts/current version/run_scenarios.R

@@ -89,7 +89,7 @@ init.age = c(1000,1000)
 phi.ad = 0.77
 p.1 = 0.98
 
-for (s in 15:15) {
+for (s in 1:1) {
   # THINGS THAT DO CHANGE
   phi.1 = scenarios[s, "Juv Surv"]
   p.ad = scenarios[s, "MR detection"]

### 2020-12-05 scripts/current version/check-trace.R
starting to delete stuff

@@ -1,0 +1,7 @@
+# example code for checking traces
+for (i in 1:18) {
+  plot(aaa[, i], type = "l")
+  lines(bbb[, i], col = 2)
+  lines(ccc[, i], col = 3)
+}
+colnames(aaa)

### 2020-12-05 scripts/current version/run_scenarios.R

@@ -89,7 +89,7 @@ init.age = c(1000,1000)
 phi.ad = 0.77
 p.1 = 0.98
 
-for (s in 1:1) {
+for (s in 2:2) {
   # THINGS THAT DO CHANGE
   phi.1 = scenarios[s, "Juv Surv"]
   p.ad = scenarios[s, "MR detection"]
@@ -218,7 +218,7 @@ for (s in 1:1) {
         Rmcmcnonest<-buildMCMC(confnonest) 
         Cmodelnonest<-compileNimble(modnonest)
         Cmcmcnonest<-compileNimble(Rmcmcnonest, project=Cmodelnonest)
-        Cmcmcnonest$run(thin=10, reset=T, niter=10000, nburnin=5000)
+        Cmcmcnonest$run(thin=10, reset=T, niter=45000, nburnin=5000)
         out<-as.data.frame(as.matrix(Cmcmcnonest$mvSamples))
         
         # save results to file and to environment
@@ -282,7 +282,7 @@ for (s in 1:1) {
         RmcmcnoMR<-buildMCMC(confnoMR) 
         CmodelnoMR<-compileNimble(modnoMR)
         CmcmcnoMR<-compileNimble(RmcmcnoMR, project=CmodelnoMR)
-        CmcmcnoMR$run(thin=10, reset=T, niter=10000, nburnin=5000)
+        CmcmcnoMR$run(thin=10, reset=T, niter=45000, nburnin=5000)
         out<-as.data.frame(as.matrix(CmcmcnoMR$mvSamples))
         
         # save results to file and to environment

### 2020-12-06 documents/DraftReport.Rmd
Edited introduction

@@
 
-Understanding the processes that control population dynamics in wild populations is a critical component of conservation but is frequently constrained by missing demographic information, uncertainty in data, and temporally or spatially misaligned datasets (citation).
+Understanding the processes that control population dynamics in wild populations is a critical component of conservation but is frequently constrained by missing demographic information, uncertainty in data, and temporally or spatially misaligned datasets (Schaub & Abadi 2011, Zipkin & Saunders 2017).
 Integrated population models (IPMs) are an increasingly popular tool in ecology that offer release from these limitations by providing a framework to combine disparate datasets in a single, unified analysis (Schaub & Abadi 2011).
@@
 In this study, we simulate data to examine IPM performance across a range of scenarios where data informing key demographic parameters are sparse or missing.
-We simulate three datasets for a for a theoretical passerine: 1) census counts, 2) reproductive output, and 3) mark-recapture (MR) of banded individuals.
-We then systematically vary the uncertainty in our abundance and MR data and run the integrated models under scenarios where demographic parameters like survival and fecundity are high or low.
-By identifying scenarios where IPMs do not lead to improved precision or cannot estimate unidentifiable parameters, we refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool.
+We simulate three datasets for a for a theoretical passerine: 1) count surveys, 2) reproductive output, and 3) mark-recapture (MR) of banded individuals.
+We then run integrated models systematically 1) omitting nest or MR data, 2) varying the detection probabilities of the abundance and MR datasets, and 3) creating scenarios where demographic parameters like survival and fecundity are high or low.
+By examining model performance, we aim to identify scenarios where IPMs do not lead to improved precision or cannot estimate unidentifiable parameters.
+In doing so, we refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool.
 

### 2020-12-06 documents/DraftReport.Rmd
lots of work on figures for multiple chains, and summary tables. Did a bit to create figs/tables automated output, but didn't want to finish until we have fuller results.

--- 2020-12-03 documents/DraftReport.Rmd
@@
 library(ggplot2)
+library(coda)
 library(captioner)
@@
 
-```{r load results, results = 'hide', message = FALSE}
-
-# test <- readRDS(here::here('data', 'out1_1.Rdata'))
+```{r figures - multiple chains, results = 'hide', message = FALSE}
 
@@
 colnames(truth) <- gsub(' ', '', colnames(truth))
-true.vals <- truth %>%
+true_vals <- truth %>%
   dplyr::select(ScenarioNumber, MRdetection, Abunddetection, AdultSurv, JuvSurv, Dailynestsurvival, Fec) %>%
@@
 
-#posterior medians for fec, mean.p, mean.phi[1:2], p.surv, phi.nest
+# ~~~~~~~~~~~~~~~~~~~~~~~~~~ #
 
-#read in scenarios separately without mean.p below
+### assessing convergence
+
 sc <- c(1:5, 11:15)
-niter <- 25
+nsim <- 25
+# sc <- c(1:2)
+# nsim <- 3
+pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'p.surv', 'phi.nest')
 
-meds <- data.frame()
+out_meds1 <- data.frame() #storage
 for (s in sc) {
-  for (i in 1:niter) {
+  for (i in 1:nsim) { 
+chains <- as.mcmc.list(list(as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 1, '.Rdata')))), 
+                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 2, '.Rdata')))), 
+                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 3, '.Rdata'))))))
 
-temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
-temp.med <- data.frame(t(apply(temp, 2, median)))
-temp.med$iter <- i
-temp.med$scenario <- s
+max_rhat <- max(gelman.diag(chains[,pars], multivariate=F, autoburnin=F)$psrf[,1], na.rm = T)
 
-meds <- rbind(meds, temp.med)
-  }
-}
+## posterior medians
+out_temp <- data.frame(t(c(summary(chains)$q[pars,"50%"], max_rhat, i, s)))
+colnames(out_temp) <- c(names(summary(chains)$q[pars,"50%"]), 'max_rhat', 'sim', 'scenario')
 
+out_meds1 <- rbind(out_meds1, out_temp)
+
+  } #sims
+} #scenarios
+
+#scenarios without p.surv
 sc <- c(6:10)
-niter <- 25
+pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'phi.nest')
 
-meds2 <- data.frame()
+out_meds2 <- data.frame() #storage
 for (s in sc) {
-  for (i in 1:niter) {
+  for (i in 1:nsim) { 
+chains <- as.mcmc.list(list(as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 1, '.Rdata')))), 
+                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 2, '.Rdata')))), 
+                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 3, '.Rdata'))))))
 
-temp <- data.frame(readRDS(file = here::here('data', paste0('out', s, '_', i, '.Rdata')))) 
-temp.med <- data.frame(t(apply(temp, 2, median)))
-temp.med$iter <- i
-temp.med$scenario <- s
+max_rhat <- max(gelman.diag(chains[,pars], multivariate=F, autoburnin=F)$psrf[,1], na.rm = T)
 
-meds2 <- rbind(meds2, temp.med)
-  }
-}
+## posterior medians
+out_temp <- data.frame(t(c(summary(chains)$q[pars,"50%"], max_rhat, i, s)))
+colnames(out_temp) <- c(names(summary(chains)$q[pars,"50%"]), 'max_rhat', 'sim', 'scenario')
 
-meds <- bind_rows(meds, meds2)
+out_meds2 <- rbind(out_meds2, out_temp)
 
-post.meds <- meds %>% #transform(scenario = factor(scenario)) %>%
-  reshape2::melt(id.vars = c('scenario', 'iter')) %>%
-  filter(variable %in% c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest')) %>%
+  } #sims
+} #scenarios
+
+out_meds <- bind_rows(out_meds1, out_meds2)
+
+post_meds <- out_meds %>% 
+  filter(max_rhat < 1.1) %>%
+  dplyr::select(-c(max_rhat, lambdaf)) %>%
+  reshape2::melt(id.vars = c('scenario', 'sim')) %>%
+  filter(variable %in% c('mean.phi[1]', 'mean.phi[2]', 'mean.p', 'fec', 'p.surv', 'phi.nest')) %>%
   transform(variable = factor(variable,
-                              levels = c('mean.phi.1.', 'mean.phi.2.', 'mean.p', 'fec', 'p.surv', 'phi.nest'),
+                              levels = c('mean.phi[1]', 'mean.phi[2]', 'mean.p', 'fec', 'p.surv', 'phi.nest'),
                               labels = c('JuvSurv', 'AdultSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'))) %>%
@@
 
-ggplot(post.meds, aes(y = value, x = scenario, group = scenario)) +
+meds_plot <- ggplot(post_meds, aes(y = value, x = scenario, group = scenario)) +
   geom_boxplot(aes(color = Data)) +
   geom_point(size = 0.7, col = 'darkgrey') + 
-  geom_point(data = true.vals, aes(y = Truth, x = scenario, group = scenario), col = 'darkred') +
+  geom_point(data = true_vals, aes(y = Truth, x = scenario, group = scenario), col = 'darkred') +
   xlab('Scenario') + ylab('Posterior medians') +
@@
 #relative bias
-
-bias.vals <- post.meds %>% 
-  merge(true.vals %>% dplyr::select(scenario, variable, Truth), by = c('scenario', 'variable')) %>%
+bias_vals <- post_meds %>% 
+  merge(true_vals %>% dplyr::select(scenario, variable, Truth), by = c('scenario', 'variable')) %>%
   transform(bias = value-Truth, rel_bias = (value-Truth)/Truth)
 
-ggplot(bias.vals, aes(y = rel_bias, x = scenario, group = scenario)) +
+bias_plot <- ggplot(bias_vals, aes(y = rel_bias, x = scenario, group = scenario)) +
   geom_boxplot(aes(color = Data)) +
@@
   scale_color_manual(values = cols[c(2,3,5)])
+
+#rmse
+rmse_vals <- post_meds %>% 
+  merge(true_vals %>% dplyr::select(scenario, variable, Truth), by = c('scenario', 'variable')) %>%
+  transform(rmse.temp = (value-Truth)^2) %>%
+  group_by(Data, scenario, variable) %>%
+  dplyr::summarize(rmse = sqrt(mean(rmse.temp)))
+
+rmse_plot <- ggplot(rmse_vals, aes(y = rmse, x = scenario, group = scenario)) +
+  geom_boxplot(aes(color = Data)) +
+  geom_point(size = 0.7, col = 'darkgrey') +
+  geom_hline(aes(yintercept = 0), linetype = 'dashed', col = 'darkred') +
+  xlab('Scenario') + ylab('RMSE') +
+  facet_wrap(~variable, scales = 'free_y') +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+        legend.position = 'top') +
+  scale_x_continuous(breaks = 1:15) +
+  scale_color_manual(values = cols[c(2,3,5)])
+
+figs("rmse_plot", caption = "Figure of RMSE per scenario.")
+
+```
+
+```{r summary table}
+
+posterior_summary <- post_meds %>%
+  group_by(scenario, variable) %>%
+  dplyr::summarize(post_med = round(median(value), 2)) %>%
+  dcast(scenario ~ variable, value.var = 'post_med')
+
+bias_summary <- bias_vals %>%
+  group_by(scenario, variable) %>%
+  dplyr::summarize(mean_bias = round(mean(rel_bias), 2)) %>%
+  dcast(scenario ~ variable, value.var = 'mean_bias')
+
+rmse_summary <- rmse_vals %>%
+  group_by(scenario, variable) %>%
+  dplyr::summarize(mean_rmse = round(mean(rmse), 2)) %>%
+  dcast(scenario ~ variable, value.var = 'mean_rmse')
 
@@
 
-```{r, fig.height = 8.5, fig.width = 7, fig.align = 'center'}
+```{r plots, fig.height = 8.5, fig.width = 7, fig.align = 'center'}
 
 # plots
+rmse_plot
 
@@
 
-`r figs("plots", caption = "Plots.")`
+`r figs("rmse_plot", caption = "Plot of RMSE per scenario.")`
 
@@
 
-`r tbls("rain_table", caption = "Table of X, Y, Z.")`
-```{r}
+`r tbls("rmse_summary", caption = "Table of RMSE by scenario.")`
+```{r table}
 
-# kable(table, align = 'c', format.args = list(big.mark = ","))
+kable(rmse_summary, align = 'c', format.args = list(big.mark = ","))
 

### 2020-12-07 documents/DraftReport.Rmd
clean-up
Fleshed out the introduction

--- 2020-12-06 documents/DraftReport.Rmd
@@
 
-```{r load packages, results = 'hide', message = FALSE}
+```{r load packages, results = 'hide'}
 
@@
 
-```{r figures - multiple chains, results = 'hide', message = FALSE}
+```{r figures - multiple chains, results = 'hide'}
 
@@
 nsim <- 25
-# sc <- c(1:2)
-# nsim <- 3
 pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'p.surv', 'phi.nest')
@@
 
-Understanding the processes that control population dynamics in wild populations is a critical component of conservation but is frequently constrained by missing demographic information, uncertainty in data, and temporally or spatially misaligned datasets (citation).
-Integrated population models (IPMs) are an increasingly popular tool in ecology that offer release from these limitations by providing a framework to combine disparate datasets in a single, unified analysis (Schaub & Abadi 2011).
-This approach can provide several advantages like estimation of otherwise non-identifiable parameters, improved precision in parameter estimation, and [spatio-temporal inference] (citation).
+Understanding the processes that control population dynamics in wild populations is a critical component of conservation and management but is frequently limited by missing demographic information, uncertainty, and temporally or spatially misaligned datasets (Schaub & Abadi 2011, Zipkin & Saunders 2017).
+Integrated population models (IPMs; Besbeas et al.
+2002, Brooks et al.
+2004) are an increasingly popular tool in ecology that can help overcome these limitations by combining disparate datasets in a single, unified analysis (Schaub & Abadi 2011).
+This approach can provide several advantages, such as improved precision (Schaub et al.
+2007, Tavecchia 2009, Abadi et al.
+2010) and the estimation of parameters that would otherwise be unidentifiable (Besbeas 2005, Schaub et al.
+2007, Veran & Lebreton 2008), which is often the case for key demographic rates such as recruitment or immigration.
 These advances have practical benefits for resource managers that must make decisions based on limited available information.
@@
 2002, 2005), the common approach has been to estimate parameters by combining mark-recapture and abundance data into a joint likelihood.
-Researchers have has since expanded the use of IPMs to include a variety of taxa (e.g.
-in bats Rhinolophus ferrumequinum: Schaub et al.
-2007, in Svalbard reindeer Rangifer tarandus platyrhynchus: Lee et al.
-2015, Antarctic toothfish Dissostichus mawsoni: Mormede et al.
-2019, wild boar Sus scrofa: Gamelon et al.
-2019), [include multivariate observations with estimation of transition probabilities within multi-state and multi-site systems] (McCrea et al.
-2012), evaluate management efforts (Cruz et al.
-2018), and to include environmental processes and stochasticity (Oppel et al.
+Researchers have has since expanded the use of IPMs to include a variety of taxa (e.g., bats *Rhinolophus ferrumequinum* (Schaub et al.
+2007); Svalbard reindeer *Rangifer tarandus platyrhynchus* (Lee et al.
+2015), Antarctic toothfish *Dissostichus mawsoni* (Mormede et al.
+2019), wild boar *Sus scrofa* (Gamelon et al.
+2019)).
+Applications have also been developed to estimate transition probabilities within multi-state and multi-site systems (McCrea et al.
+2012), evaluate management efficacy and trade-offs (Cruz et al.
+2018), and to estimate the effects of environmental processes and stochasticity on demography (Oppel et al.
 2014, Woodworth et al.
 2017).
-This growth in usage has occurred despite a relatively poor understanding of how to assess goodness of fit however, and we lack knowledge of the limits of the IPM framework in situations where available data are sparse, of poor quality, or when certain data sources are absent.
-For example, many field studies may have survey and reproductive data while missing survival estimates from mark-recapture datasets.
+This proliferation of integrated population modeling has occurred despite a relatively poor understanding of how to assess goodness of fit for the joint likelihood.
+Additionally, the limitations of the IPM framework are still underexplored, particularly in situations where available data are sparse, of poor quality, or when data to inform specific parameters are entirely absent (i.e., most field studies may have count surveys, but not all have data informing both recruitment and survival).
 
-Data simulation studies provide an avenue for robust evaluation of the IPM framework, but few such studies exist.
-Those that have used data simulation largely focus on proof of 1) concept (i.e.
-improved precision and reduced bias McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al.
+Data simulation studies provide an avenue for robust evaluations of the IPM framework, but few such studies exist.
+Those that have used data simulation largely focus on (1) proof of concept (i.e., improved precision and reduced bias of parameter estimation) or (2) violation of assumptions within the subcomponent models (i.e., marker-induced bias or heterogeneity in survival) or the joint likelihood (i.e., dataset independence).
+In the former, IPMs generally have led to improved precision and unbiased estimates of parameters that would otherwise be unidentifiable using traditional approaches (McCrea et al.
+2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al.
 2017, Ahrestani et al.
 2017, Bowler et al.
-2019), and 2) model assumptions such as independence of datasets (Abadi et al.
+2019), though not always depending on sample sizes, dominance of one dataset over another, or the likelihood structures (citation).
+In examining the effects of violating model assumptions on bias and precision, results have been mixed and again largely depend on sample sizes, effect sizes, parameterization, and the dominance of given datasets.
+In some cases, dependence across datasets has been shown to result in biased estimates or at least overly optimistic estimates of uncertainty (Goodman et al.
+2004, Besbeas et al.
+2009) while in others the effects on bias and precision are minimal (Abadi et al.
 2010, Schaub & Fletcher 2015, Weegman et al.
@@
 Riecke et al.
-(2019) found that parameter estimates from IPMs were highly sensitive to violations of mark-recapture model assumptions (i.e.
-marker-induced bias and survival heterogeneity).
-These findings further motivate studies to determine the reliability of IPMs in parameter estimation and in detecting trends when survival is not informed by data or when a data source is sparse, biased, or highly uncertain.
+(2019) found that estimates for previously unidentifiable parameters were highly sensitive to violations of mark-recapture model assumptions (i.e., marker-induced bias and survival heterogeneity).
 
-We aim to improve our understanding of this important statistical approach, particularly for passerine bird life history.
-In this study, we simulate data to examine IPM performance across a range of scenarios where data informing key demographic parameters are sparse or missing.
-We simulate three datasets for a for a theoretical passerine: 1) census counts, 2) reproductive output, and 3) mark-recapture (MR) of banded individuals.
-We then systematically vary the uncertainty in our abundance and MR data and run the integrated models under scenarios where demographic parameters like survival and fecundity are high or low.
-By identifying scenarios where IPMs do not lead to improved precision or cannot estimate unidentifiable parameters, we refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool.
+These studies motivate the further examination of the strengths and weaknesses IPMs across a range of model structures and data landscapes, particularly when certain demographic rates are not directly informed by data or when a data source is sparse, biased, or highly uncertain.
+This type of analysis has been conducted for species distribution models in a few cases (Farr et al.
+2020, Simmonds et al.
+2020), but has not yet been rigorously explored for an IPM.
+Through a simulation analysis, we aim to gain insights about this statistical framework, particularly as applied to a passerine bird life history.
+We compare parameter estimation for a traditional IPM with data informing abundance (count data), survival (mark-recapture data), and reproductive output (nest monitoring data) to situations when either of the latter two datasets are omitted.
+In future iterations of this analysis, we also plan to compare model performance in situations when datasets informing survival and reproduction are sparse (reduced survey frequency) or highly uncertain (reduced detection probability).
+
+Through this study, we will identify scenarios where IPMs may not lead to improved precision or return unbiased estimates of parameters that would otherwise be unidentifiable when certain datasets are excluded.
+In doing so, we will refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool or might not be necessary or warrant the collection of additional data.
+Improving our understanding of the strengths and weaknesses of the IPM framework can provide helpful insights for the design and implementation of field survey programs so as to most efficiently make use of limited resources to inform management efforts aimed at monitoring or recovering species that are deemed ecosystem indicators or depleted and in need of conservation intervention.
 

### 2020-12-07 documents/DraftReport.Rmd
More intro edits.

@@
 We compare parameter estimation for a traditional IPM with data informing abundance (count data), survival (mark-recapture data), and reproductive output (nest monitoring data) to situations when either of the latter two datasets are omitted.
-In future iterations of this analysis, we also plan to compare model performance in situations when datasets informing survival and reproduction are sparse (reduced survey frequency) or highly uncertain (reduced detection probability).
+In each of these three cases, we also examine parameter estimation when juvenile survival varies and under situations with low detection to reflect a higher degree of uncertainty.
+In future iterations of this analysis, we also plan to compare model performance in situations when datasets informing survival and reproduction are not only uncertain (low detection probability) but also sparse (reduced survey frequency), and the effects of having lower or variable adult survival and reproductive output.
 

### 2020-12-07 documents/DraftReport.Rmd
More intro edits.

@@
 Through a simulation analysis, we aim to gain insights about this statistical framework, particularly as applied to a passerine bird life history.
-We compare parameter estimation for a traditional IPM with data informing abundance (count data), survival (mark-recapture data), and reproductive output (nest monitoring data) to situations when either of the latter two datasets are omitted.
+We assume all model assumptions are met and compare parameter estimation for a traditional IPM with data informing abundance (count data), survival (mark-recapture data), and reproductive output (nest monitoring data) to situations when either of the latter two datasets are omitted.
 In each of these three cases, we also examine parameter estimation when juvenile survival varies and under situations with low detection to reflect a higher degree of uncertainty.

### 2020-12-07 documents/DraftReport.Rmd
small edits

@@
 
-cols <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")
-
 ```
@@
 tbls <- captioner(prefix = "Table")
+cols <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")
 
@@
 
-sc <- c(1:5, 11:15)
-nsim <- 25
+# sc <- c(1:5, 11:15)
+sc <- c(1:5)
+nsim <- 5
 pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'p.surv', 'phi.nest')
@@
 #scenarios without p.surv
-sc <- c(6:10)
+# sc <- c(6:10)
+sc <- c(6:9)
+
 pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'phi.nest')

### 2020-12-08 documents/DraftReport.Rmd
nicer figures, results summary tables

@@
 Presentation
-- caroline/amelia
+<!-- - caroline/amelia -->
+-designate/delegate sections/slides
 
 Write-up
-- intro (amanda re-reads and checks for possible additions)
+<!-- - intro (amanda re-reads and checks for possible additions) -->
 - methods/equation check: change YOTY to juv (hannah)
@@
 - posterior densities (amanda)
-- life cycle/survey timing diagram (amelia)
+<!-- - life cycle/survey timing diagram (amelia) -->
 - figure captions/printing (amanda)
-- 3 dags (amelia)
+<!-- - 3 dags (amelia) -->
 
@@
 - posterior densities
-- table (true values, data included, posterior mean, 95% CI, mean bias, mean RMSE)
+<!-- - table (true values, data included, posterior mean, 95% CI, mean bias, mean RMSE) -->
 
@@
 
-```{r figures - multiple chains, results = 'hide'}
+```{r figures - multiple chains, results = 'hide', cache = F}
 
@@
   transform(variable = factor(variable,
-                              levels = c('AdultSurv', 'JuvSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'),
-                              labels = c('AdultSurv', 'JuvSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'))) %>%
+                              levels = c('JuvSurv', 'AdultSurv', 'Fec', 'Dailynestsurvival', 'MRdetection', 'Abunddetection'),
+                              labels = c('Juvenile survival', 'Adult survival', 'Fecundity', 'Daily nest survival', 
+                                         'MR detection', 'Survey detection'))) %>%
   transform(Data = ifelse(scenario < 6, 'All', 
                           ifelse(scenario < 11 & scenario > 5, 'Counts + Productivity', 'Counts + MR'))) %>%
-  transform(Truth = ifelse(Data == 'Counts + Productivity' & variable == 'MRdetection', NA, Truth))
+  transform(Truth = ifelse(Data == 'Counts + Productivity' & variable == 'MR detection', NA, Truth))
 
@@
 
-# sc <- c(1:5, 11:15)
-sc <- c(1:5)
-nsim <- 5
+# test <-readRDS(file = here::here('data', paste0('out_3-1.Rdata')))[3500:4000,]
+
+sc <- c(1:5, 11:15)
+# sc <- c(1:5)
+nsim <- 25
 pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'p.surv', 'phi.nest')
@@
   for (i in 1:nsim) { 
-chains <- as.mcmc.list(list(as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 1, '.Rdata')))), 
-                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 2, '.Rdata')))), 
-                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 3, '.Rdata'))))))
+chains <- as.mcmc.list(list(
+  as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 1, '.Rdata')))[3500:4000,]), 
+  as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 2, '.Rdata')))[3500:4000,]), 
+  as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 3, '.Rdata')))[3500:4000,])))
 
-max_rhat <- max(gelman.diag(chains[,pars], multivariate=F, autoburnin=F)$psrf[,1], na.rm = T)
+max_rhat <- max(gelman.diag(chains[,pars[-6]], multivariate=F, autoburnin=F)$psrf[,1], na.rm = T)
 
@@
 #scenarios without p.surv
-# sc <- c(6:10)
-sc <- c(6:9)
+sc <- c(6:10)
+# sc <- c(6:9)
 
-pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'phi.nest')
+pars <- c('fec', 'lambdaf', 'mean.phi[1]', 'mean.phi[2]', 'p.surv', 'phi.nest')
 
@@
   for (i in 1:nsim) { 
-chains <- as.mcmc.list(list(as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 1, '.Rdata')))), 
-                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 2, '.Rdata')))), 
-                            as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 3, '.Rdata'))))))
+chains <- as.mcmc.list(list(
+  as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 1, '.Rdata')))[3500:4000,]), 
+  as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 2, '.Rdata')))[3500:4000,]), 
+  as.mcmc(readRDS(file = here::here('data', paste0('out', s, '_', i, '-', 3, '.Rdata')))[3500:4000,])))
 
@@
 post_meds <- out_meds %>% 
-  filter(max_rhat < 1.1) %>%
+  filter(max_rhat < 1.8) %>%
   dplyr::select(-c(max_rhat, lambdaf)) %>%
   reshape2::melt(id.vars = c('scenario', 'sim')) %>%
-  filter(variable %in% c('mean.phi[1]', 'mean.phi[2]', 'mean.p', 'fec', 'p.surv', 'phi.nest')) %>%
+  filter(variable %in% c('mean.phi[1]', 'mean.phi[2]', 'fec', 'phi.nest', 'p.surv', 'mean.p')) %>%
   transform(variable = factor(variable,
-                              levels = c('mean.phi[1]', 'mean.phi[2]', 'mean.p', 'fec', 'p.surv', 'phi.nest'),
-                              labels = c('JuvSurv', 'AdultSurv', 'MRdetection', 'Fec', 'Abunddetection', 'Dailynestsurvival'))) %>%
+                              levels = c('mean.phi[1]', 'mean.phi[2]', 'fec', 'phi.nest', 'mean.p', 'p.surv'),
+                              labels = c('Juvenile survival', 'Adult survival', 'Fecundity', 'Daily nest survival', 
+                                         'MR detection', 'Survey detection'))) %>%
   transform(Data = ifelse(scenario < 6, 'All', 
@@
   xlab('Scenario') + ylab('Posterior medians') +
-  facet_wrap(~variable, scales = 'free_y') +
+  facet_wrap(~variable, scales = 'free_y', nrow = 3) +
   theme_classic() +
@@
   xlab('Scenario') + ylab('Relative bias') +
-  facet_wrap(~variable, scales = 'free_y') +
+  facet_wrap(~variable, scales = 'free_y', nrow = 3) +
   theme_classic() +
@@
 
-rmse_plot <- ggplot(rmse_vals, aes(y = rmse, x = scenario, group = scenario)) +
-  geom_boxplot(aes(color = Data)) +
-  geom_point(size = 0.7, col = 'darkgrey') +
-  geom_hline(aes(yintercept = 0), linetype = 'dashed', col = 'darkred') +
-  xlab('Scenario') + ylab('RMSE') +
-  facet_wrap(~variable, scales = 'free_y') +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
-        legend.position = 'top') +
-  scale_x_continuous(breaks = 1:15) +
-  scale_color_manual(values = cols[c(2,3,5)])
-
-figs("rmse_plot", caption = "Figure of RMSE per scenario.")
+# rmse_plot <- ggplot(rmse_vals, aes(y = rmse, x = scenario, group = scenario)) +
+#   geom_boxplot(aes(color = Data)) +
+#   geom_point(size = 0.7, col = 'darkgrey') +
+#   # geom_hline(aes(yintercept = 0), linetype = 'dashed', col = 'darkred') +
+#   xlab('Scenario') + ylab('RMSE') +
+#   facet_wrap(~variable, scales = 'free_y', nrow = 3) +
+#   theme_classic() +
+#   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+#         legend.position = 'top') +
+#   scale_x_continuous(breaks = 1:15) +
+#   scale_color_manual(values = cols[c(2,3,5)])
+# 
+# figs("rmse_plot", caption = "Figure of RMSE per scenario.")
 
@@
   dcast(scenario ~ variable, value.var = 'mean_bias')
+write.csv(bias_summary, file = here::here('results', 'bias_summary.csv'))
 
@@
   dcast(scenario ~ variable, value.var = 'mean_rmse')
+write.csv(rmse_summary, file = here::here('results', 'rmse_summary.csv'))
+
+summary_tab <- bias_summary %>%
+  merge(rmse_summary, by = 'scenario', suffixes = c('.bias', '.rmse')) %>%
+  merge(true_vals %>% dplyr::select(scenario, Data) %>% distinct(), by = 'scenario') %>%
+  dplyr::select(Data, scenario, JuvSurv.bias, JuvSurv.rmse, AdultSurv.bias, AdultSurv.rmse, 
+                Fec.bias, Fec.rmse, Dailynestsurvival.bias, Dailynestsurvival.rmse, 
+                Abunddetection.bias, Abunddetection.rmse, MRdetection.bias, MRdetection.rmse)
+
+write.csv(summary_tab, file = here::here('results', 'summary_table.csv'))
 

### 2020-12-09 documents/DraftReport.Rmd
Rough draft of scenario text, not finalized

@@
 *Simulation scenarios* 
-
-Scenarios - see sheet
+We developed three versions of an IPM: 1) a "full" IPM that incorporates all three datasets (counts, productivity, mark-resight); 2) an IPM with the counts and productivity datasets only (excludes mark-resight dataset); and, 3) an IPM with the counts and mark-resight datasets only (excludes the productivity dataset).
+There are numerous scenarios we could run to vary detection and survival probabilities for each IPM version and test our objectives.
+We opted to run a select number of scenarios to provide preliminary findings to inform future work.
+For each version of the IPM, we ran 5 scenarios, for a total of 15 scenarios.
+Within each scenario, we varied the following parameters: mark-resight probability (0.3, 0.5, 0.8), count detection probability (0.3, 0.5, 0.85), and juvenile survival (0.3, 0.4, 0.5).
+The remaining parameters were kept constant for all scenarios: daily nest survival (0.975), mean clutch size (2.5), max nest age (30), initial age distribution (1000 juveniles, 1000 adults), and adult survival (0.77).
+For demographic parameters, we chose to vary juvenile survival because it is a hidden parameter, and these various values result in 5% annual decline, steady population growth, and 5% annual growth, respectively.
 

### 2020-12-10 documents/DraftReport.Rmd
Added references to lit cited

@@
 
+Besbeas P., Borysiewicz R.S., Morgan B.J., 2009.
+Completing the Ecological Jigsaw.
+In: Thomson D.L., Cooch E.G., Conroy M.J.
+(eds) Modeling Demographic Processes In Marked Populations.
+Environmental and Ecological Statistics, vol 3.
+Springer, Boston, MA.
+https://doi.org/10.1007/978-0-387-78151-8_22
+
+Besbeas, P., Freeman, S.N., Morgan, B.J.T., Catchpole, E.A., 2002. Integrating mark-recapture-recovery and census data to estimate animal abundance and demographic parameters. Biometrics 58, 540–547.
+
+Besbeas, P., Freeman, S.N., Morgan, B.J.T., 2005.
+The potential of integrated population modelling.
+Australian & New Zealand Journal of Statistics 47, 35-48.
+https://doi.org/10.1111/j.1467-842X.2005.00370.x
+
 Bled, F., Belant, J.L., Daele, L.J.V., Svoboda, N., Gustine, D., Hilderbrand, G., Barnes, V.G., n.d.
@@
 
+Brooks S.P., King R., Morgan B.J.T., 2004. A Bayesian approach to
+combining animal abundance and demographic data. Animal
+Biodiversity and Conservation 27, 515–529.
+
 Chandler, R.B., Clark, J.D., 2014. Spatially explicit integrated population models. Methods in Ecology and Evolution 5, 1351–1360. https://doi.org/10.1111/2041-210X.12153
 
+Cruz, J., Windels, S.K., Thogmartin, W.E., Crimmins, S.M., Grim, L.H., Zuckerberg, B., 2018.
+Managing individual nests promotes population recovery of a top predator.
+Journal of Applied Ecology 55, 1418–1429.
+
 Farr, M.T., Green, D.S., Holekamp, K.E., Zipkin, E.F., n.d. Integrating distance sampling and presence-only data to estimate species abundance. Ecology n/a, e03204. https://doi.org/10.1002/ecy.3204
+
+Gamelon, M., Baubet, E., Besnard, A., Gaillard, J.-M., Lebreton, J.-D., Touzot, L., Veylit, L., Gimenez, O., 2019.Efficient use of harvest data: An integrated population model for exploited animal populations.
+bioRxiv 776104.
+doi: https://doi.org/10.1101/776104
+
+Lee, A.M., Bjørkvoll, E.M., Hansen, B.B., Albon, S.D., Stien, A., Sæther, B.‐E., Engen, S., Veiberg, V., Loe, L.E., Grøtan, V., 2015.
+An integrated population model for a long‐lived ungulate: more efficient data use with Bayesian methods.
+Oikos 124, 806-816.
+https://doi.org/10.1111/oik.01924
 
@@
 
+Mormede, S., Parker, S.J., Pinkerton, M.H., 2020.
+Comparing spatial distribution modelling of fisheries data with single-area or spatially-explicit integrated population models, a case study of toothfish in the Ross Sea region.
+Fisheries Research 221, 105381. https://doi.org/10.1016/j.fishres.2019.105381
+
+Oppel, S., Hilton, G., Ratcliffe, N., Fenton, C., Daley, J., Gray, G., Vickery, J., Gibbons, D., 2014.
+Assessing population viability while accounting for demographic and environmental uncertainty.
+Ecology 95, 1809–1818.
+
 Riecke, T.V., Williams, P.J., Behnke, T.L., Gibson, D., Leach, A.G., Sedinger, B.S., Street, P.A., Sedinger, J.S., 2019.
@@
 
+Schaub, M., Abadi, F., 2011. Integrated population models: A novel analysis framework for deeper insights into population dynamics. Journal of Ornithology 152:S227–S237.
+
 Schaub, M., Fletcher, D., 2015. Estimating immigration using a Bayesian integrated population model: choice of 
 parametrization and priors. Environ Ecol Stat 22, 535–549. https://doi.org/10.1007/s10651-015-0309-8
+
+Schaub, M., Gimenez, O., Sierro, A., Arlettaz, R., 2007.
+Use of Integrated Modeling to Enhance Estimates of Population Dynamics Obtained from Limited Data.
+Conservation Biology 21, 945-955.
+https://doi.org/10.1111/j.1523-1739.2007.00743.x
 
@@
 
+Véran, S., Lebreton, J.D., 2008.
+The potential of integrated modelling in conservation biology: A case study of the black‐footed albatross (Phoebastria nigripes).
+Can J Statistics, 36, 85-98.
+https://doi.org/10.1002/cjs.5550360109
+
 Weegman, M.D., Arnold, T.W., Clark, R.G., Schaub, M., 2020.
@@
 https://doi.org/10.1002/eap.2258
+
+Woodworth, B., Wheelwright, N., Newman, A., Schaub, M., Norris, D.R., 2017.
+Winter temperatures limit population growth rate of a migratory songbird.
+Nat Commun 8, 14812.
+https://doi.org/10.1038/ncomms14812
+
+Zipkin, E.F., Saunders, S.P., 2018. Synthesizing multiple data types for biological conservation using integrated population models. Biological Conservation 217, 240–250.
 

### 2020-12-11 documents/DraftReport.Rmd
discussion text

@@
 
-Presentation
-<!-- - caroline/amelia -->
--designate/delegate sections/slides
-
 Write-up
-<!-- - intro (amanda re-reads and checks for possible additions) -->
 - methods/equation check: change YOTY to juv (hannah)
@@
 - results (abby/hannah)
-
-- discussion (TBD)
--- summarize high-level finding
-
--- why we think we see what we see
---- crux of the issue.... 
-(1) how the life cycle model relies on different data sources
-(2) parameter identifiability
-(3) ?
-
--- limitations/caveats of this particular study (what can/can't we make inference to/about)
---- life histories
-
--- future directions
---- degrading quantity/quality
-
--- importance of the work
-(1) understanding IPMs better
-(2) understanding weaknesses of IPMs: when does the integration almost hurt instead of help?
-(3) survey design
 
@@
 <!-- - 3 dags (amelia) -->
-
-Code review
-- annotate data simulation (amanda! caroline! amelia!)
 
@@
 
-Review project aim and main findings
+This simulation analysis aims to explore the benefits and weaknesses of data integration for population models across varying levels of data quality (degree of uncertainty) and quantity (sample size, number of available datasets) for a typical passerine bird species.
+In essence, one of the main goals of the study was to examine when and to what extent and under what circumstances developing IPMs is worth the potential additional resources needed for data collection.
+Or, similarly, when this type of data integration framework can facilitate the estimation of parameters not directly informed by available data.
+Our initial results reaffirm the benefits of data integration for avian population modeling and highlight avenues of future investigation.
 
-Elaborate on interesting finding 1  
-- Implications of not exploring this in an application?  
-- Future areas of work?  
-- Extension to more complex life history? Data scenarios?
+In our example of a typical passerine life history, the full IPM was the only data landscape in which parameters were successfully returned with relatively little to no bias.
+In this case, we can also see the importance of integrating all three datasets for the estimation of juvenile survival, which is not directly informed by any of the datasets.
+Even in the full IPM, however, we do see increased bias and decreased accuracy (as measured by higher RMSE) in survival (particularly juvenile) at low detection rates that is not improved upon when the survival rate is higher.
+This suggests that detection probability is a strong determinant in parameter uncertainty.
+An important future scenario to examine under this full data landscape is whether that uncertainty could be overcome by increasing sample sizes and/or survey frequency.
 
-Elaborate on interesting finding 2
-- Implications of not exploring this in an application?  
-- Future areas of work?  
-- Extension to more complex life history? Data scenarios?
+When examining model performance across the other reduced data landscapes, similar patterns arose when removing either mark-resight or productivity datasets.
+With only two datasets, bias increased and accuracy was reduced only for parameters directly informed by the dataset that was omitted, and this deterioration in model performance was much greater than from decreased detection probability in the full IPM.
+This pattern, if explored further, will start to reveal some of the trade-offs that are most interesting to more fully investigate using this framework.
+Specifically, it will be helpful to examine to what extent sample size would have to increase and uncertainty would have to be reduced in order to estimate the implicit parameters with a similar degree of accuracy and precision as with the IPM.
+And, when aiming to improve data quality and quantity, whether it can be accomplished more readily by increasing sample sizes and/or detection probability in mark-resight data versus productivity data.
+In other words, for our passerine, if, when, and to what extent does one dataset dominate another in terms of estimating abundance?
 
-Broaden to implications of work, study design, existing IPMs
+With the progression of these analyses, our aim is to additionally understand how some of these trade-offs within an IPM framework might vary for other life histories (e.g., additional life stages) and survey features.
+Here we assume that all datasets are independent and that mark-recapture model assumptions hold, but it would be possible to explore violations of any of these.
+We are also interested in examining the ability of our IPM to detect a trend in demographic rates or abundance across the three data landscapes or in the face of unmodeled heterogeneity.
+
+Most research applications using an IPM framework cite the benefits of improved parameter estimation, including reduced bias and the ability to estimate implicit parameters, though not all authors can or choose to confirm this through data simulation or compare their IPM estimates with those from traditional modeling frameworks.
+It is likely that the benefits of IPMs are fairly variable and context-specific, meaning that the degree to which IPMs improve parameter estimation depend on life history, model assumptions, sample sizes, parameter uncertainty, and likelihood structure.
+This type of study can enhance our growing understanding of IPM performance when datasets may be sparse, biased, or both.
+Depending on the goal of a given monitoring or conservation and management program, it can be helpful to understand how biases in existing data affect our ability to produce accurate estimates of abundance and trends, and how it might or might not be possible to overcome sparseness or bias by collecting additional data.
+This information can be key to designing and implementing monitoring surveys for avian populations.
 

### 2020-12-11 documents/DraftReport.Rmd
eval = F

@@
 
-```{r figures - multiple chains, results = 'hide', cache = F}
+```{r figures - multiple chains, eval = F, results = 'hide', cache = F}
 
@@
 
-```{r summary table}
+```{r summary table, eval = F}
 
@@
 
-```{r plots, fig.height = 8.5, fig.width = 7, fig.align = 'center'}
+```{r plots, eval = F, fig.height = 8.5, fig.width = 7, fig.align = 'center'}
 
@@
 `r tbls("rmse_summary", caption = "Table of RMSE by scenario.")`
-```{r table}
+```{r table, eval = F}
 

### 2020-12-13 documents/DraftReport.Rmd
Methods and equation check, changed some things to match the updated simulation and model. Changed the notation for some of the parameters

@@
 Write-up
-- methods/equation check: change YOTY to juv (hannah)
+- methods/equation check: change YOTY to juv (hannah)*done* I changed some of the writing about the model and the notation used, could use a second look over by another person!
+
 - simulation write-up (amelia)
@@
 We simulated data for an avian species assuming life history stages and traits of a typical passerine.
-Our simulated hypothetical species has two age-classes: young-of-year (YOY) and adults.
-Each age class has a unique survival rate ($S_1$, $S_{ad}$) and adults produce new chicks at rate *f*, and those chicks survive to be YOY at rate $S_1$.
-Current YOY will also breed the following year to produce surviving chicks at rate $f \times S_1$ (Fig 1).
+Our simulated hypothetical species has two age-classes: 1 year-old juveniles and adults (individuals older than 1 year).
+Each age class has a unique survival rate ($\phi_1$, $\phi_{ad}$) and adults produce new chicks at fecundity rate *f*, and those chicks survive to be juveniles at rate $\phi_1$.
+Current juveniles will also breed the following year to produce surviving chicks at rate $f \times \phi_1$ (Fig 1).
 We assume an annual breeding cycle, with females nesting once per year.
 The period from nest initiation to fledging is 30 days regardless of clutch size.
+We also assume that survival and fecundity are constant across years and individuals.
 
@@
 *Count model*  
-We simulated data such that the population is surveyed annually in a 15-year pre-breeding census.
+We simulated data such that the population is surveyed annually in a 10-year pre-breeding census.
 Count surveys are conducted three times per year in quick succession such that we assume the population is closed between secondary sampling periods.
 We use an N-mixture model (citations) to estimate annual abundance from these repeated counts, where we assume no individual is double counted and that individual detections are independent and occur with equal probability.
-The underlying state process represents the latent variables of stage-specific abundance at each occasion:
+The underlying state process represents the latent variables of stage-specific abundance at each occasion for 1 year-olds, $N_{1}$ is modeled as:
 
-$$N_{1, t+1} \sim \text{Poisson}(N_{1,t}, S_{1,t}, \frac{f_t}{2} + N_{ad,t}S_{ad,t},\frac{f_t}{2})$$
+$$N_{1, t+1} \sim \text{Poisson}(N_{1,t}\times \phi_{1}\times \frac{f}{2} + N_{ad,t}\times \phi_{ad}\times \frac{f}{2})$$  
+where the number of juveniles in the next year, $N_{1,t+1}$, is estimated using a Poisson distribution with probability based on the sum of stage-specific individuals at time $t$ that survive ($\phi_{1,ad}$) and reproduce ($f$).
 
-$$N_{ad,t+1} \sim \text{Binomial}(N_{1,t} + N_{ad,t}, S_{ad,t})$$
+A binomial distribution is used to estimate the number of adults in the next year, $N_{ad, t+1}$ based on the stage-specific survival and abundance in the previous occasion (i.e. $t$):
+$$N_{ad,t+1} \sim \text{Binomial}(N_{1,t} + N_{ad,t}, \phi_{ad,t})$$
 
-where the number of juveniles ($N_1$) is estimated using a Poisson distribution with probability based on the sum of stage-specific individuals that survive ($S_{1,ad}$) and reproduce ($f_t$).
-A binomial distribution is used to estimate the number of adults based on the stage-specific survival and abundance in the previous occasion.
+The observation process model for the count data in each year $t$ is modeled using a binomial distribution:
 
-The observation process model,
+$$y_t \sim \text{Binomial} (N_{tot, t}, p_s)$$
+with $N_{tot,t}=N_{1,t}+N_{ad,t}$.
+The observed counts in each year, $y_t$, are assumed to be binomially distributed with a sample size of $N_{tot,t}$ and probability of detection $p_s$.
+The probability of detection for the observation process, $p_s$, of the count model is assumed constant across years and surveys.
+The survey detection probability was modeled on the probability scale using a uniform prior, $p_{s} \sim \text{Uniform}(0,1)$.
 
-$$y_t = N_{tot, t} + \epsilon$$
-$$\epsilon \sim \text{Normal}(0, \sigma^2_y)$$
-
-describes counts *y* at time *t* given given stage-specific abundance as binomially distributed with sample size $N_t$ and detection probability $p$, with normally distributed observation error $\epsilon$ estimated with mean = 0 with the prior for $\sigma_y$ being U(0,10).
-<!-- from amanda: help - prior for observation error?  -->
-
-The overall likelihood for this hierarchical state-space model can be described as the product of the state and observation process models:
-$$L_{SS}(\text{y}|\text{N}, S_1, S_{ad}, f, \sigma^2_y) = L_O(\text{y}|\text{N},\sigma^2_y) \times L_S(\text{N}|S_1, S_{ad}, f)$$
+The likelihood of the N-mixture abundance model is the product of the state and observation process models: $L_{Abundance}(\text{y}| ~N_1, ~N_{ad}, ~{p_s}, \phi_1,~\phi_{ad}, f)$
 
@@
 We assume the daily survival probability is uniform across nest stages (*i.e.*, laying, incubating, nestlings).
-We estimate daily nest survival to obtain unbiased estimates for the probability of nest success (*i.e.*, at least 1 chick fledges): We use a Poisson model to estimate the average number of chicks fledged conditional on a nest being successful.
+We estimate daily nest survival to obtain unbiased estimates for the probability of nest success (*i.e.*, at least 1 chick fledges): We use a Poisson model to estimate the number of chicks fledged conditional on a nest being successful.
 
-$$J_t \sim \text{Poisson}(\rho_t) $$
-$$\rho_t = R_tf_t $$
+$$\text{Fledged}_t \sim \text{Poisson}(\lambda_{\text{fledglings}}) $$
+The fecundity rate, *f*, is derived from the rate at which chicks fledge,
+$$f=\frac{1}{2}\times\lambda_{\text{fledglings}}*\phi_{n}^{30}$$
+where $\phi_{n}$ is the nest survival probability and we assume that the maximum nest age, the number of days the nest was active, is 30. 
+The observation process for the daily nest surveys is modeled using a Bernoulli distribution, based on daily nest survival probabilities and the state of the nest during the previous survey occasion:
 
-where the number of nestlings counted $J$ in year *t* is modeled using a Poisson distribution with the product of the number of surveyed broods $R_t$ and fecundity $f_t$ at each occasion.
-Fecundity is estimated using a uniform prior (0, 20) and the likelihood of this daily nest survival model can be described as $L_P(\text{J}, \text{R}|\text{f})$.
+$$\text{NestStatus}_{n,d}\sim\text{Bernoulli}(\phi_{n}*\text{NestStatus}_{n,d-1})$$
+where $\text{NestStatus}_{n,d}$ is the status of nest $n$ at survey occasion $d$, $\phi_{n}$ is the nest surivial probability, and $\text{NestStatus}_{n,d-1}$ is the state of nest $n$ at the previous survey occasion $d-1$.
+
+The likelihood of the nest success productivity model is $L_{Productivity}(\text{NestStatus}, \text{Fledged} | ~\phi_{n}, ~\lambda_{\text{fledglings}})$
 
 *Mark-resight model*  
-In this simulation study, YOY chicks are marked in the nest just prior to fledging with a unique combination of colored leg bands such that they can be individually identified over the course of their lives.
-Every chick in every nest that fledges is banded from a single annual comprehensive mark-resight survey per breeding season.
+In this simulation study, juveniles and adults are marked just prior to the end of the breeding season with a unique combination of colored leg bands such that they can be individually identified over the course of their lives.
+Marked individuals are resighted during marking efforts in future years.
 To estimate stage-specific survival based on the individual encounter histories, we use a Cormack-Jolly-Seber (citation) model where we assume that marks are not lost or misread and that each marked individual has independent and identical survival and detection probabilities across occasions.
@@
 
-$$ z_{i,t}|z_{i,t-1} \sim \text{Bernoulli}(z_{i,t-1} S_{1,ad}) $$
+$$ z_{i,t}|z_{i,t-1} \sim \text{Bernoulli}(z_{i,t-1}\times \phi_{ad}) $$
 
-where the state at time *t* is modeled with a Bernouli distribution based on the state at the previous occasion and state-specific survival probabilities $S_1$ and $S_{ad}$.
-Similarly, the observation process model
+where the state at time *t* is modeled with a Bernoulli distribution based on the state at the previous occasion and adult survival probability $\phi_{ad}$. Similarly, the observation process model
 
-$$ y_{i,t}|z_{i,t} \sim  \text{Bernoulli}(z_{i,t} p)$$
+$$ CH_{i,t}|z_{i,t} \sim  \text{Bernoulli}(z_{i,t}\times p_{mr})$$
 
-describes observations *y* using a Bernouli distribution based on the state of the individual in the current occasion and detection probability *p*.
-Mean survival and detection probabilities are modeled on the probability scale using uniform priors, $S_{1,ad} \sim \text{Uniform}(0,1)$ and $p \sim \text{Uniform}(0,1)$, respectively.
-The overall likelihood of the mark-resight model estimating survival probabilities is the product of the state and observation process models: $L_{SS}(\text{y}|\text{z}, ~S_1, ~S_{ad}, ~\text{p}) = L_O(\text{y}|\text{z},~\text{p})     \times L_S(\text{z}|S_1,~S_{ad})$.
+describes capture history observations *CH* using a Bernouli distribution based on the state of the individual in the current occasion and detection probability $p_{mr}$.
+Mean survival and detection probabilities are modeled on the probability scale using uniform priors, $\phi_{ad} \sim \text{Uniform}(0,1)$ and $p_{mr} \sim \text{Uniform}(0,1)$, respectively.
+The likelihood of the mark-resight model estimating the adult survival probability: $L_{Survival}(\text{y}|\text{z}, ~\phi_{ad}, ~p_{mr})$
 
 *Integrated population model*  
-
 To estimate abundance of the overall population $N_{tot}$ at each time *t*, the parameter estimates from the productivity and mark-resight models are combined with the stage-specific abundane estimated using the N-mixture model in a stochastic leslie population matrix model.
@@
 [IPM joint likelihood equation]
+The overall likelihood for this hierarchical state-space model can be described as the product of the state and observation process models:
+$$L_{IPM}(\text{y}, ~CH, ~\text{NestStatus}, ~\text{Fledged}|~N_{1},~N_{ad}, phi_1, phi_{ad}, f, p_{mr}, p_s, ~\lambda_{\text{fledglings}}, ~\phi_n) 
+$$=$L_{Abundance}(\text{y}| ~N_1, ~N_{ad}, ~{p_s}, \phi_1,~\phi_{ad}, f)\times$$
+$$L_{Productivity}(\text{NestStatus}, \text{Fledged} | ~\phi_{n}, ~\lambda_{\text{fledglings}}) \times$$
+$$L_{Survival}(\text{y}|\text{z}, ~\phi_{ad}, ~p_{mr})$$
 

### 2020-12-13 documents/DraftReport.Rmd
DRAFT - seriously a draft - of results

@@
 **Results**  
+-something about lack of convergence?
+
+Fitting the 'full' IPM model, including all simulated data types (counts, productivity, and mark-resight), generally produced parameter estimates close to the true value.
+The median model parameter estimates for $\phi_1$, $\phi_ad$, $f$, $\phi_n$, and $\p_{mr}$ across the 25 simulated data sets and under scenarios 1-5 resulted in estimates where the true parameter values are contained within $50%$ of the estimates.
+The model parameter estimates for the survey detection probability, $p_s$, were returned under scenarios where $p_s$ was $0.85$ (i.e.
+scenarios 1 and 2), but with greater uncertainty in the model estimates for scenarios 3,4,5 where the true value of $p_s$ was $0.3$ or $0.5$ (posterior median boxplot figure).
+Relative bias under scenarios $1-5$ was generally low (less than $|0.1|$), with the exection of $p_s$ with high bias resulting in scenarios 3,4,5 where the model was not able to recover this parameter (boxplot and table for bias).
+RMSE for the full model was also generally low (less than $|0.1|$), with the exception of $\phi_1$ and $p_s$.
+The RMSE for $\phi_1$ was high for scenarios 3, which the RMSE for $p_s$ was high for scenarios 3,4,5 (Table for RMSE).
+
+The IPM fit with data only from count and productivity datasets (excluding mark-resight) provided median estimates of $\phi_n$ and $f$ that were close to truth under all scenarios (6-10).
+The median estimates for $\phi_1$ and $\phi_{ad}$ showed high variation in the model estimates for all scenarios and was unable to capture the true parameter values for scenarios 9 and 10 for both parameters.
+The survey detection probability median model estimates for scenarios 6,7,and 10 resulted in medians covering the true parameter estimates, but under scenarios 8 and 9 the model estimates over and underestimated the true parameter value, respectively (Boxplot Figure).
+The relative bias for $\phi_1$ under scenarios 7,9,and 10 was high (greater than $|0.1|$), the relative bias for $\phi_ad$ under scenario 9 was also high.
+The relative bias for $p_s$ was high under scenarios 8 and 9 (Boxplot and table for relative bias).
+The RMSE of the model estimates for $\phi_n$ and $f$ under scenarios 6-10 was close to 0.
+The RMSE for $\phi_1$ was high (greater than 0.1) for all scenarios (6-10), the RMSE for $\phi_{ad}$ was high for scenarios 6,7,9,and 10.
+The RMSE for $p_s$ was high for scenarios 8,9,and 10 (Table for RMSE).
+
+The IPM fit with data only from count and mark-resight datasets (excluding productivity data) resulted in median estimates of $\phi_{ad}$ and $p_{mr}$ that were close to truth for scenarios 11-15.
+The model excluding the productivity data resulted in medians that underestimated $f$, $\phi_ad$, and $\phi_1$ under all scenarios.
+The model was able to capture the true value of $p_{s}$ under scenarios 11, 12, 13, and 15 (Boxplot figure).
+The relative bias and RMSE for $\phi_1$ and $f$ were high.
+Whereas the relative bias and RMSE for all other parameters in scenarios 11-15 was low (less than |0.1|; bias boxplot and table).
 

### 2020-12-14 documents/DraftReport.Rmd
Comment
Small stylistic edits to equations (thanks Hannah!)

@@
 - simulation write-up (amelia)
-- results (abby/hannah)
+- results (abby/hannah) - Hannah: wrote out a rough draft of results, will revisit tomorrow, but got something down to start with
 
@@
 The fecundity rate, *f*, is derived from the rate at which chicks fledge,
-$$f=\frac{1}{2}\times\lambda_{\text{fledglings}}*\phi_{n}^{30}$$
+$$f=\frac{1}{2}\times\lambda_{\text{fledglings}}\times\phi_{n}^{30}$$
 where $\phi_{n}$ is the nest survival probability and we assume that the maximum nest age, the number of days the nest was active, is 30. 
@@
 
-$$\text{NestStatus}_{n,d}\sim\text{Bernoulli}(\phi_{n}*\text{NestStatus}_{n,d-1})$$
+$$\text{NestStatus}_{n,d}\sim\text{Bernoulli}(\phi_{n}\times\text{NestStatus}_{n,d-1})$$
 where $\text{NestStatus}_{n,d}$ is the status of nest $n$ at survey occasion $d$, $\phi_{n}$ is the nest surivial probability, and $\text{NestStatus}_{n,d-1}$ is the state of nest $n$ at the previous survey occasion $d-1$.
@@
 *Integrated population model*  
-To estimate abundance of the overall population $N_{tot}$ at each time *t*, the parameter estimates from the productivity and mark-resight models are combined with the stage-specific abundane estimated using the N-mixture model in a stochastic leslie population matrix model.
+To estimate abundance of the overall population $N_{tot}$ at each time *t*, the parameter estimates from the productivity and mark-resight models are combined with the stage-specific abundance estimated using the N-mixture model in a stochastic leslie population matrix model.
 
@@
 The overall likelihood for this hierarchical state-space model can be described as the product of the state and observation process models:
-$$L_{IPM}(\text{y}, ~CH, ~\text{NestStatus}, ~\text{Fledged}|~N_{1},~N_{ad}, phi_1, phi_{ad}, f, p_{mr}, p_s, ~\lambda_{\text{fledglings}}, ~\phi_n) 
-$$=$L_{Abundance}(\text{y}| ~N_1, ~N_{ad}, ~{p_s}, \phi_1,~\phi_{ad}, f)\times$$
-$$L_{Productivity}(\text{NestStatus}, \text{Fledged} | ~\phi_{n}, ~\lambda_{\text{fledglings}}) \times$$
-$$L_{Survival}(\text{y}|\text{z}, ~\phi_{ad}, ~p_{mr})$$
+$$L_{IPM}(\text{y}, ~CH, ~\text{NestStatus}, ~\text{Fledged}|~N_{1},~N_{ad}, ~\phi_1, ~\phi_{ad}, ~f, ~p_{mr}, ~p_s, ~\lambda_{\text{fledglings}}, ~\phi_n) \ = $$
+$$L_{Abundance}(\text{y}| ~N_1, ~N_{ad}, ~{p_s}, ~\phi_1,~\phi_{ad}, ~f)\ \times$$
+$$L_{Productivity}(\text{NestStatus}, ~\text{Fledged} | ~\phi_{n}, ~\lambda_{\text{fledglings}})\ \times$$
+$$L_{Survival}(\text{y}|~\text{z}, ~\phi_{ad}, ~p_{mr})$$
 

### 2020-12-14 documents/DraftReport.Rmd
small edits to scenarios and estimation/diagnostics sections

@@
 
-Metrics to monitor  
-CV (precision)  
-RMSE (accuracy)  
-Relative bias
+<!-- amelia - maybe easiest to create table in word doc when we also add figures? -->
 
-Vary data quantity/combinations and demographic rates  
-Low MR detection vs high  
-Low daily nest survival vs high  
-Low abund detection vs high
+<!-- amelia - I think these are all old notes? -->
+<!-- Metrics to monitor   -->
+<!-- CV (precision)   -->
+<!-- RMSE (accuracy)   -->
+<!-- Relative bias   -->
 
-Best scenario, using all three datasets  
-Potential add-on: vary life history traits:  
-Survival: 0.3 vs. 0.8  
-Fecundity: 0.3 vs. 0.5  
-Remove nest monitoring entirely - use other two datasets (MR and counts)  
-Model adjustments: informed priors?   
-Sensitivity to priors?  
-Remove MR data entirely - use other two datasets (nest monitoring and counts)  
-Model adjustments: informed priors?  
-Sensitivity to priors?   
-Poorer quality MR  
-Quantity: decrease N (20 individuals vs. 100 resights per year)  
-Quality: decrease p (0.1, 0.5, 0.8) or annual surveys (5 vs. 10)
+<!-- Vary data quantity/combinations and demographic rates   -->
+<!-- Low MR detection vs high   -->
+<!-- Low daily nest survival vs high   -->
+<!-- Low abund detection vs high   -->
+
+<!-- Best scenario, using all three datasets   -->
+<!-- Potential add-on: vary life history traits:   -->
+<!-- Survival: 0.3 vs. 0.8   -->
+<!-- Fecundity: 0.3 vs. 0.5   -->
+<!-- Remove nest monitoring entirely - use other two datasets (MR and counts)   -->
+<!-- Model adjustments: informed priors?    -->
+<!-- Sensitivity to priors?   -->
+<!-- Remove MR data entirely - use other two datasets (nest monitoring and counts)   -->
+<!-- Model adjustments: informed priors?   -->
+<!-- Sensitivity to priors?    -->
+<!-- Poorer quality MR   -->
+<!-- Quantity: decrease N (20 individuals vs. 100 resights per year)   -->
+<!-- Quality: decrease p (0.1, 0.5, 0.8) or annual surveys (5 vs. 10)   -->
 
 *Estimation and diagnostics*   
-Nimble; iterations, burn-in, etc  
+<!-- amelia - just a quick draft, but is this right? -->
+Models were ran in NIMBLE using program R (NIMBLE 2020).
+We ran 45,000 iterations with the first 5,000 as a burn-in.
+We thinned samples to reduce autocorrelation and retained every 10th sample, for a total of 4,000 samples.
+Model convergence was assessed visually using trace plots.
+
 Examined relative bias, RMSE, etc
@@
 
-Abadi, F., Gimenez, O., Arlettaz, R., Schaub, M., 2010.
+Abadi, F., Gimenez, O., Arlettaz, R., Schaub, M., 2010a.
 An assessment of integrated population models: bias, accuracy, and violation of the assumption of independence.
@@
 Fisheries Research 221, 105381. https://doi.org/10.1016/j.fishres.2019.105381
+
+NIMBLE Development Team.
+2020.
+NIMBLE: MCMC, Particle Filtering, and Programmable Hierarchical Modeling.
+doi: 10.5281/zenodo.1211190.
+R package version 0.10.1, https://cran.r-project.org/package=nimble.
 

### 2020-12-14 documents/DraftReport.Rmd
Added Brooks-Gelman-Rubin statistic note to Estimation & Diagnostics (but need more info)
abby changes

@@
 
-Write-up
-- methods/equation check: change YOTY to juv (hannah)*done* I changed some of the writing about the model and the notation used, could use a second look over by another person!
+Figures/analysis/etc
 
-- simulation write-up (amelia)
-- results (abby/hannah) - Hannah: wrote out a rough draft of results, will revisit tomorrow, but got something down to start with
+*my understanding is that these are all done - just need to be pasted in to the final doc*
 
-Figures/analysis/etc
 - posterior densities (amanda)
@@
 Ideas: 
-- Performance of integrated population models across varying data landscapes  
+- Performance of integrated population models across varying data landscapes *happy with this one for now because it's consistent with our presentation*
 - When and how far can we push the integrated population model framework? Estimating demographic parameters under variable data landscapes  
@@
 As is often the convention, we use a female-only model because we assume that the population dynamics of our species are driven largely by female abundance and demography.
-All demographic rates are assumed constant over a 15-year study period.
+All demographic rates are assumed constant over a 10-year study period.
 The population is quite small and the survey extends across the entire range, therefore we assume no immigration or emigration.
+
+*insert life cycle diagram here - figure 1*
 
@@
 
+*insert figure for observation timeline - figure 2*
+
 *Integrated population model*  
@@
 
-<!-- From amanda: help: not sure if we need something like this since we have the equations above? -->
+<!-- From amanda: help: not sure if we need something like this since we have the equations above?
+aeb - we can use begin{align} if we want and put $~ or &= where we want things to be centered around, but I think this is good enough for now-->
 <!-- $$ \left[\begin{array} -->
@@
 The joint likelihood for the integrated population model is the product of the likelihoods for the three subcomponent models.
-In instances when one dataset is removed, informative priors are used in place of the data to inform each of the relevant parameters (Table 1).
+In instances when one dataset is removed, noninformative priors are used in place of the data to inform each of the relevant parameters.
 
-[IPM joint likelihood equation]
+[IPM joint likelihood equation] *is this how we want this?*
 The overall likelihood for this hierarchical state-space model can be described as the product of the state and observation process models:
 $$L_{IPM}(\text{y}, ~CH, ~\text{NestStatus}, ~\text{Fledged}|~N_{1},~N_{ad}, ~\phi_1, ~\phi_{ad}, ~f, ~p_{mr}, ~p_s, ~\lambda_{\text{fledglings}}, ~\phi_n) \ = $$
-$$L_{Abundance}(\text{y}| ~N_1, ~N_{ad}, ~{p_s}, ~\phi_1,~\phi_{ad}, ~f)\ \times$$
+$$L_{Abundance}(\text{y}| ~N_1, ~N_{ad}, ~{p_s},~\phi_{ad}, ~f)\ \times$$
 $$L_{Productivity}(\text{NestStatus}, ~\text{Fledged} | ~\phi_{n}, ~\lambda_{\text{fledglings}})\ \times$$
@@
 *Simulation scenarios*  
-We developed three versions of an IPM: 1) a "full" IPM that incorporates all three datasets (counts, productivity, mark-resight); 2) an IPM with the counts and productivity datasets only (excludes mark-resight dataset); and, 3) an IPM with the counts and mark-resight datasets only (excludes the productivity dataset).
+We developed three versions of an IPM: 1) a "full" IPM that incorporates all three datasets (counts, productivity, mark-resight); 2) an IPM with the counts and productivity datasets only (excludes mark-resight dataset); and, 3) an IPM with the counts and mark-resight datasets only (excludes the productivity dataset) (Figure 3).
 There are numerous scenarios we could run to vary detection and survival probabilities for each IPM version and test our objectives.
 We opted to run a select number of scenarios to provide preliminary findings to inform future work.
-For each version of the IPM, we ran 5 scenarios, for a total of 15 scenarios.
+For each version of the IPM, we ran 5 scenarios, for a total of 15 scenarios (Table for scenarios).
 Within each scenario, we varied the following parameters: mark-resight probability (0.3, 0.5, 0.8), count detection probability (0.3, 0.5, 0.85), and juvenile survival (0.3, 0.4, 0.5).
 The remaining parameters were kept constant for all scenarios: daily nest survival (0.975), mean clutch size (2.5), max nest age (30), initial age distribution (1000 juveniles, 1000 adults), and adult survival (0.77).
-For demographic parameters, we chose to vary juvenile survival because it is a hidden parameter, and these various values result in 5% annual decline, steady population growth, and 5% annual growth, respectively.
+For demographic parameters, we chose to vary juvenile survival because it is a hidden parameter, and these various values result in 5% annual decline, a stable population size, and 5% annual growth, respectively.
+
+*need a table here - table 1 for true values, table 2 for varied params*
+
+*also have the three dags next to each other - figure 3*
 
@@
 *Estimation and diagnostics*   
-<!-- amelia - just a quick draft, but is this right? -->
+<!-- amelia - just a quick draft, but is this right? AEB checked - made one small fix -->
 Models were ran in NIMBLE using program R (NIMBLE 2020).
-We ran 45,000 iterations with the first 5,000 as a burn-in.
-We thinned samples to reduce autocorrelation and retained every 10th sample, for a total of 4,000 samples.
-Model convergence was assessed visually using trace plots.
-
-Examined relative bias, RMSE, etc
+We ran 45,000 iterations with the first 35,000 as a burn-in.
+We thinned samples to reduce autocorrelation and retained every 10th sample, for a total of 1,000 samples.
+Model convergence was assessed visually using trace plots and the Brooks-Gelman-Rubin statistic (Rhat).
+We examined model performance using relative bias and root mean square error (RMSE).
 
@@
 
+- figures here
+- *posterior medians - figure 4*
+- *relative bias - figure 5*
+- *rmse table - table 2*
+
+*Full model*
 Fitting the 'full' IPM model, including all simulated data types (counts, productivity, and mark-resight), generally produced parameter estimates close to the true value.
@@
 The model parameter estimates for the survey detection probability, $p_s$, were returned under scenarios where $p_s$ was $0.85$ (i.e.
-scenarios 1 and 2), but with greater uncertainty in the model estimates for scenarios 3,4,5 where the true value of $p_s$ was $0.3$ or $0.5$ (posterior median boxplot figure).
-Relative bias under scenarios $1-5$ was generally low (less than $|0.1|$), with the exection of $p_s$ with high bias resulting in scenarios 3,4,5 where the model was not able to recover this parameter (boxplot and table for bias).
+scenarios 1 and 2), but with greater uncertainty in the model estimates for scenarios 3,4,5 where the true value of $p_s$ was $0.3$ or $0.5$ (Figure 4).
+Relative bias under scenarios $1-5$ was generally low (less than $|0.1|$), with the exection of $p_s$ with high bias resulting in the low detection scenarios 3,4,5 where the model was not able to recover this parameter (Figure 5).
 RMSE for the full model was also generally low (less than $|0.1|$), with the exception of $\phi_1$ and $p_s$.
-The RMSE for $\phi_1$ was high for scenarios 3, which the RMSE for $p_s$ was high for scenarios 3,4,5 (Table for RMSE).
+The RMSE for $\phi_1$ was high for scenarios 3, which the RMSE for $p_s$ was high for scenarios 3,4,5 (Table 3).
 
+*Counts + Productivity*
 The IPM fit with data only from count and productivity datasets (excluding mark-resight) provided median estimates of $\phi_n$ and $f$ that were close to truth under all scenarios (6-10).
 The median estimates for $\phi_1$ and $\phi_{ad}$ showed high variation in the model estimates for all scenarios and was unable to capture the true parameter values for scenarios 9 and 10 for both parameters.
-The survey detection probability median model estimates for scenarios 6,7,and 10 resulted in medians covering the true parameter estimates, but under scenarios 8 and 9 the model estimates over and underestimated the true parameter value, respectively (Boxplot Figure).
+The survey detection probability median model estimates for scenarios 6,7,and 10 resulted in medians covering the true parameter estimates, but under scenarios 8 and 9 the model estimates over and underestimated the true parameter value, respectively (Figure 4).
 The relative bias for $\phi_1$ under scenarios 7,9,and 10 was high (greater than $|0.1|$), the relative bias for $\phi_ad$ under scenario 9 was also high.
-The relative bias for $p_s$ was high under scenarios 8 and 9 (Boxplot and table for relative bias).
+The relative bias for $p_s$ was high under scenarios 8 and 9 (Figure 5).
 The RMSE of the model estimates for $\phi_n$ and $f$ under scenarios 6-10 was close to 0.
 The RMSE for $\phi_1$ was high (greater than 0.1) for all scenarios (6-10), the RMSE for $\phi_{ad}$ was high for scenarios 6,7,9,and 10.
-The RMSE for $p_s$ was high for scenarios 8,9,and 10 (Table for RMSE).
+The RMSE for $p_s$ was high for scenarios 8,9,and 10 (Table 3).
 
+*Counts + Mark-resight*
 The IPM fit with data only from count and mark-resight datasets (excluding productivity data) resulted in median estimates of $\phi_{ad}$ and $p_{mr}$ that were close to truth for scenarios 11-15.
 The model excluding the productivity data resulted in medians that underestimated $f$, $\phi_ad$, and $\phi_1$ under all scenarios.
-The model was able to capture the true value of $p_{s}$ under scenarios 11, 12, 13, and 15 (Boxplot figure).
+The model was able to capture the true value of $p_{s}$ under scenarios 11, 12, 13, and 15 (Figure 4).
 The relative bias and RMSE for $\phi_1$ and $f$ were high.
-Whereas the relative bias and RMSE for all other parameters in scenarios 11-15 was low (less than |0.1|; bias boxplot and table).
+Whereas the relative bias and RMSE for all other parameters in scenarios 11-15 was low (less than |0.1|; Figure 5).
 

### 2020-12-18 documents/DraftReport.Rmd

--- 2020-12-14 documents/DraftReport.Rmd
@@
 We opted to run a select number of scenarios to provide preliminary findings to inform future work.
-For each version of the IPM, we ran 5 scenarios, for a total of 15 scenarios.
+For each version of the IPM, we ran 5 scenarios, for a total of 15 scenarios (Table for scenarios).
 Within each scenario, we varied the following parameters: mark-resight probability (0.3, 0.5, 0.8), count detection probability (0.3, 0.5, 0.85), and juvenile survival (0.3, 0.4, 0.5).
@@
 We thinned samples to reduce autocorrelation and retained every 10th sample, for a total of 4,000 samples.
-Model convergence was assessed visually using trace plots.
-
-Examined relative bias, RMSE, etc
+Model convergence was assessed visually using trace plots and the Brooks-Gelman-Rubin statistic (Rhat).
+We examined model performance using relative bias and root mean square error (RMSE).
 

### 2021-04-26 scripts/EuringPosterFigures.Rmd
markdown file for euring poster figures.

---
title: "EuringPosterFigures"
author: ""
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)

library(tidyr)
library(dplyr)
library(cowplot)
library(ggplot2)
library(coda)
library(captioner)
library(knitr)
library(reshape2)
library(here)

```

```{r data load}

load(file = here::here('results', 'EuringPoster', 'processedIPMOutput.RData'))

row.low <- do.call(rbind, lapply( ls(patt='lowout'), get) )
row.med <- do.call(rbind, lapply( ls(patt='medout'), get) )
row.high <- do.call(rbind, lapply( ls(patt='highout'), get) )
rm(list=grep('highout|medout|lowout',ls(),value=TRUE,invert=FALSE))

low <- row.low %>%
  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`)

low.params <- readRDS(file = here::here('data', 'low.lam.params.rds')) %>%
  transform(scenario = 1:25)

med <- row.med %>%
  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`)

med.params <- readRDS(file = here::here('data', 'med.lam.params.rds')) %>%
  transform(scenario = 1:25)

high <- row.high %>%
  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`)

high.params <- readRDS(file = here::here('data', 'high.lam.params.rds')) %>%
  transform(scenario = 1:25)

#combine to get true values integrated in results
low.dat <- low %>%
  inner_join(low.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Decreasing')

med.dat <- med %>%
  inner_join(med.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Stable')

high.dat <- high %>%
  inner_join(high.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Increasing')

all.dat <- bind_rows(low.dat, med.dat, high.dat)

```

```{r bias}

#get posterior medians
all.meds <- all.dat %>%
  dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
                            'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'sims', 'scenario', 'simscenarios')) %>%
  group_by(lambda.scenario, scenario, simscenarios, variable) %>%
  summarize(median = median(value), .groups = 'keep') %>%
  reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median')

#just bias
rel.bias <- all.meds %>%
  #calculate relative bias, mean across scenarios
  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
            fec.bias = (fec.obs-fec.true)/fec.true) %>%
  dplyr::select('lambda.scenario', 'scenario', 'simscenarios', 
                'phi1.bias', 'phiad.bias', 'fec.bias') %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
  group_by(lambda.scenario, simscenarios, variable) %>%
  dplyr::summarize(mean.rel.bias = mean(value), .groups = 'keep') %>%
  transform(simscenarios = factor(simscenarios, levels = c(1:9), 
                                  labels = rep(c('L', 'M', 'H'), 3)))

ggplot(rel.bias, aes(x = factor(simscenarios), y = variable, fill = mean.rel.bias)) +
  geom_tile() +
  xlab('Detection level') + ylab('') +
  facet_grid(~lambda.scenario, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "Relative bias",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))

```

```{r rmse}

#for the love of buddha, someone please check this calculation of rmse
rmse.vals <- all.meds %>%
  transform(rmse.fec = (fec.obs-fec.true)^2,
            rmse.phiad = (phiad.obs-phiad.true)^2,
            rmse.phi1 = (phi1.obs-phi1.true)^2) %>%
  dplyr::select(lambda.scenario, scenario, simscenarios, rmse.fec, rmse.phiad, rmse.phi1) %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
  group_by(simscenarios, lambda.scenario, variable) %>%
  dplyr::summarize(rmse = sqrt(mean(value)), .groups = 'keep') %>%
  transform(simscenarios = factor(simscenarios, levels = c(1:9), 
                                  labels = rep(c('L', 'M', 'H'), 3)))

#dunno about this - looks suspiciously similar across detection levels :(
ggplot(rmse.vals, aes(x = factor(simscenarios), y = variable, fill = rmse)) +
  geom_tile() +
  xlab('Detection level') + ylab('') +
  facet_grid(~lambda.scenario, drop = T, scales = 'free_x') +
  scale_fill_gradient2(name = "RMSE",
                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
  theme_bw() +
  theme(legend.position = 'top',
             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))

```

```{r not using but keep for now}

### Keep to look at difference in bias across scenarios
# #get posterior medians
# low.meds <- low.dat %>%
#   dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
#                             'sims', 'scenario', 'simscenarios') %>%
#   reshape2::melt(id.vars = c('sims', 'scenario', 'simscenarios')) %>%
#   group_by(scenario, simscenarios, variable) %>%
#   summarize(median = median(value)) %>%
#   reshape2::dcast(scenario + simscenarios ~ variable, value.var = 'median') %>%
#   #calculate relative bias
#   transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true, 
#             phiad.bias = (phiad.obs-phiad.true)/phiad.true, 
#             fec.bias = (fec.obs-fec.true)/fec.true)
# 
# #just bias
# low.bias <- low.meds %>%
#   dplyr::select('scenario', 'simscenarios', 'phi1.bias', 'phiad.bias', 'fec.bias') %>%
#   reshape2::melt(id.vars = c('scenario', 'simscenarios')) %>%
#   transform(simscenarios = factor(simscenarios, labels = c('lam = low; p=L', 'M', 'H')))
# 
# ggplot(low.bias, aes(x = factor(scenario), y = variable, fill = value)) +
#   geom_tile() +
#   xlab('Scenario') + ylab('Parameter') +
#   facet_grid(~simscenarios, drop = T, scales = 'free_x') +
#   scale_fill_gradient2(name = "Relative bias",
#                       mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))

```

### 2021-04-27 scripts/EuringPosterFigures.Rmd
bias across life history scenarios

@@
 
-```{r not using but keep for now}
+```{r bias across life history scenarios}
 
-### Keep to look at difference in bias across scenarios
-# #get posterior medians
-# low.meds <- low.dat %>%
-#   dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
-#                             'sims', 'scenario', 'simscenarios') %>%
-#   reshape2::melt(id.vars = c('sims', 'scenario', 'simscenarios')) %>%
-#   group_by(scenario, simscenarios, variable) %>%
-#   summarize(median = median(value)) %>%
-#   reshape2::dcast(scenario + simscenarios ~ variable, value.var = 'median') %>%
-#   #calculate relative bias
-#   transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true, 
-#             phiad.bias = (phiad.obs-phiad.true)/phiad.true, 
-#             fec.bias = (fec.obs-fec.true)/fec.true)
-# 
-# #just bias
-# low.bias <- low.meds %>%
-#   dplyr::select('scenario', 'simscenarios', 'phi1.bias', 'phiad.bias', 'fec.bias') %>%
-#   reshape2::melt(id.vars = c('scenario', 'simscenarios')) %>%
-#   transform(simscenarios = factor(simscenarios, labels = c('lam = low; p=L', 'M', 'H')))
-# 
-# ggplot(low.bias, aes(x = factor(scenario), y = variable, fill = value)) +
-#   geom_tile() +
-#   xlab('Scenario') + ylab('Parameter') +
-#   facet_grid(~simscenarios, drop = T, scales = 'free_x') +
-#   scale_fill_gradient2(name = "Relative bias",
-#                       mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
-#   theme_bw() +
-#   theme(legend.position = 'top',
-#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+##Same as above but looking at the variability across data-generating true values for phiad, phi1, fec
+
+#get posterior medians; same as above except keep 'scenario' in group_by
+all.meds <- all.dat %>%
+  dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
+                            'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
+  reshape2::melt(id.vars = c('sims', 'scenario', 'simscenarios', 'lambda.scenario')) %>%
+  group_by(lambda.scenario, scenario, simscenarios, variable) %>%
+  summarize(median = median(value)) %>%
+  reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median') %>%
+  #calculate relative bias
+  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
+            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
+            fec.bias = (fec.obs-fec.true)/fec.true)
+
+#just bias; same as above, keep 'scenario'
+rel.bias <- all.meds %>%
+  dplyr::select('lambda.scenario', 'scenario', 'simscenarios', 
+                'phi1.bias', 'phiad.bias', 'fec.bias') %>%
+  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
+  transform(simscenarios = factor(simscenarios, labels = rep(c('L', 'M', 'H'),3)))
+
+#grab true values from above
+true_vals <- all.dat %>%
+  distinct(lambda.scenario, scenario, phiad.true, fec.true, phi1.true) %>%
+  transform(phiad.true = round(phiad.true,1),
+            phi1.true = round(phi1.true,1), fec.true = round(fec.true,1))
+
+#merge back onto bias df so can have something to name/view the 'scenarios' 
+rel.bias.vals <- rel.bias %>%
+  merge(true_vals, by = c('lambda.scenario', 'scenario'))
+
+#scenario number -- not a useful view, but inspired this investigation
+ggplot(rel.bias, aes(x = factor(scenario), y = variable, fill = value)) +
+  geom_tile() +
+  xlab('Scenario') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+#view by fec.true
+ggplot(rel.bias.vals, aes(x = factor(fec.true), y = variable, fill = value)) +
+  geom_tile() +
+  xlab('True fecundity') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#view by phiad.true
+ggplot(rel.bias.vals, aes(x = factor(phiad.true), y = variable, fill = value)) +
+  geom_tile() +
+  xlab('True phi.ad') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#view by phi1.true
+ggplot(rel.bias.vals, aes(x = factor(phi1.true), y = variable, fill = value)) +
+  geom_tile() +
+  xlab('True phi.1') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
 

### 2021-04-27 scripts/EuringPosterFigures.Rmd
small edits to bias plot

--- 2021-04-26 scripts/EuringPosterFigures.Rmd
@@
 library(here)
-
+library(RColorBrewer)
 ```
@@
 
+# set factor levels
+rel.bias$lambda.scenario <- factor(rel.bias$lambda.scenario, levels = c("Decreasing", "Stable", "Increasing"))
+
+# plot
 ggplot(rel.bias, aes(x = factor(simscenarios), y = variable, fill = mean.rel.bias)) +
@@
   scale_fill_gradient2(name = "Relative bias",
-                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
-  theme_bw() +
+                      mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+  theme_minimal() +
   theme(legend.position = 'top',

### 2021-04-27 scripts/EuringPosterFigures.Rmd
color palette fixed

@@
     scale_fill_gradient2(name = "Relative bias",
-                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
+                      mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2"))
   theme_bw() +
@@
     scale_fill_gradient2(name = "Relative bias",
-                      mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
+                      mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
   theme_bw() +

### 2021-04-28 scripts/EuringPosterFigures.Rmd
case study figures added, changed color palette to something bolder

--- 2021-04-27 scripts/EuringPosterFigures.Rmd
@@
 rm(list=grep('highout|medout|lowout',ls(),value=TRUE,invert=FALSE))
+
+```
+
+```{r process data}
 
@@
   transform(simscenarios = factor(simscenarios, levels = c(1:9), 
-                                  labels = rep(c('L', 'M', 'H'), 3)))
-
-# set factor levels
-rel.bias$lambda.scenario <- factor(rel.bias$lambda.scenario, levels = c("Decreasing", "Stable", "Increasing"))
+                                  labels = rep(c('L', 'M', 'H'), 3))) %>%
+  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias'),
+                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>% 
+  transform(lambda.scenario = factor(lambda.scenario, 
+                                     levels = c("Decreasing", "Stable", "Increasing")))
 
@@
 ggplot(rel.bias, aes(x = factor(simscenarios), y = variable, fill = mean.rel.bias)) +
-  geom_tile() +
+  geom_tile(color = 'grey50') +
   xlab('Detection level') + ylab('') +
@@
   scale_fill_gradient2(name = "Relative bias",
-                      mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                     mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
                       breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
@@
 rmse.vals <- all.meds %>%
-  transform(rmse.fec = (fec.obs-fec.true)^2,
-            rmse.phiad = (phiad.obs-phiad.true)^2,
-            rmse.phi1 = (phi1.obs-phi1.true)^2) %>%
-  dplyr::select(lambda.scenario, scenario, simscenarios, rmse.fec, rmse.phiad, rmse.phi1) %>%
+  transform(fec.rmse = (fec.obs-fec.true)^2,
+            phiad.rmse = (phiad.obs-phiad.true)^2,
+            phi1.rmse = (phi1.obs-phi1.true)^2) %>%
+  dplyr::select(lambda.scenario, scenario, simscenarios, fec.rmse, phiad.rmse, phi1.rmse) %>%
   reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
@@
   transform(simscenarios = factor(simscenarios, levels = c(1:9), 
-                                  labels = rep(c('L', 'M', 'H'), 3)))
+                                  labels = rep(c('L', 'M', 'H'), 3))) %>%
+  transform(variable = factor(variable, levels = c('phiad.rmse', 'phi1.rmse', 'fec.rmse'),
+                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>% 
+  transform(lambda.scenario = factor(lambda.scenario, 
+                                     levels = c("Decreasing", "Stable", "Increasing")))
 
@@
 ggplot(rmse.vals, aes(x = factor(simscenarios), y = variable, fill = rmse)) +
-  geom_tile() +
+  geom_tile(color = 'grey50') +
   xlab('Detection level') + ylab('') +
@@
 
-```{r not using but keep for now}
+```{r bias across life history scenarios}
 
-### Keep to look at difference in bias across scenarios
-# #get posterior medians
-# low.meds <- low.dat %>%
-#   dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
-#                             'sims', 'scenario', 'simscenarios') %>%
-#   reshape2::melt(id.vars = c('sims', 'scenario', 'simscenarios')) %>%
-#   group_by(scenario, simscenarios, variable) %>%
-#   summarize(median = median(value)) %>%
-#   reshape2::dcast(scenario + simscenarios ~ variable, value.var = 'median') %>%
-#   #calculate relative bias
-#   transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true, 
-#             phiad.bias = (phiad.obs-phiad.true)/phiad.true, 
-#             fec.bias = (fec.obs-fec.true)/fec.true)
-# 
-# #just bias
-# low.bias <- low.meds %>%
-#   dplyr::select('scenario', 'simscenarios', 'phi1.bias', 'phiad.bias', 'fec.bias') %>%
-#   reshape2::melt(id.vars = c('scenario', 'simscenarios')) %>%
-#   transform(simscenarios = factor(simscenarios, labels = c('lam = low; p=L', 'M', 'H')))
-# 
-# ggplot(low.bias, aes(x = factor(scenario), y = variable, fill = value)) +
-#   geom_tile() +
-#   xlab('Scenario') + ylab('Parameter') +
-#   facet_grid(~simscenarios, drop = T, scales = 'free_x') +
-#   scale_fill_gradient2(name = "Relative bias",
-#                       mid = "#FFFFFF", low = "#012345", high = "#012345", midpoint = 0) +
-#   theme_bw() +
-#   theme(legend.position = 'top',
-#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+##Same as above but looking at the variability across data-generating true values for phiad, phi1, fec
+
+#get posterior medians; same as above except keep 'scenario' in group_by
+all.meds.sc <- all.dat %>%
+  dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
+                            'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
+  reshape2::melt(id.vars = c('sims', 'scenario', 'simscenarios', 'lambda.scenario')) %>%
+  group_by(lambda.scenario, scenario, simscenarios, variable) %>%
+  summarize(median = median(value)) %>%
+  reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median') %>%
+  #calculate relative bias
+  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
+            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
+            fec.bias = (fec.obs-fec.true)/fec.true)
+
+#just bias; same as above, keep 'scenario'
+rel.bias.sc <- all.meds.sc %>%
+  dplyr::select('lambda.scenario', 'scenario', 'simscenarios', 
+                'phi1.bias', 'phiad.bias', 'fec.bias') %>%
+  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
+  transform(simscenarios = factor(simscenarios, labels = rep(c('L', 'M', 'H'),3)))
+
+#grab true values from above
+true_vals <- all.dat %>%
+  distinct(lambda.scenario, scenario, phiad.true, fec.true, phi1.true) %>%
+  transform(phiad.true = round(phiad.true,1),
+            phi1.true = round(phi1.true,1), fec.true = round(fec.true,1))
+
+#for categories: 
+fec_lims <- quantile(true_vals$fec.true, probs = c(0.33, 0.7), names = F)
+phiad_lims <- quantile(true_vals$phiad.true, probs = c(0.33, 0.7), names = F)
+phi1_lims <- quantile(true_vals$phi1.true, probs = c(0.33, 0.7), names = F)
+
+#merge back onto bias df so can have something to name/view the 'scenarios' 
+rel.bias.vals <- rel.bias.sc %>%
+  merge(true_vals, by = c('lambda.scenario', 'scenario')) %>%
+  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias'),
+                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>% 
+  transform(lambda.scenario = factor(lambda.scenario, 
+                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
+  transform(fec_cat = ifelse(fec.true < fec_lims[1], 'L', ifelse(fec.true > fec_lims[2], 'H', 'M')),
+            phiad_cat = ifelse(phiad.true < phiad_lims[1], 'L', ifelse(phiad.true > phiad_lims[2], 'H', 'M')),
+            phi1_cat = ifelse(phi1.true < phi1_lims[1], 'L', ifelse(phi1.true > phi1_lims[2], 'H', 'M'))) %>%
+  transform(fec_cat = factor(fec_cat, levels = c('L', 'M', 'H')),
+            phiad_cat = factor(phiad_cat, levels = c('L', 'M', 'H')),
+            phi1_cat = factor(phi1_cat, levels = c('L', 'M', 'H')))
+
+#scenario number -- not a very useful view, but inspired this investigation
+ggplot(rel.bias.sc, aes(x = factor(scenario), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('Scenario') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+#view by fec.true vals; 
+ggplot(rel.bias.vals, aes(x = factor(fec.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True fecundity') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#view by fec category: have to re-summarize to get mean across fec/phi ifelse categories
+fec.plot.vals <- rel.bias.vals %>%
+  group_by(lambda.scenario, simscenarios, variable, fec_cat) %>%
+  dplyr::summarize(value = mean(value), .groups = 'keep')
+
+ggplot(fec.plot.vals, aes(x = factor(fec_cat), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True fecundity') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#view by phiad.true
+ggplot(rel.bias.vals, aes(x = factor(phiad.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True phi.ad') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#view by phiad category
+phiad.plot.vals <- rel.bias.vals %>%
+  group_by(lambda.scenario, simscenarios, variable, phiad_cat) %>%
+  dplyr::summarize(value = mean(value), .groups = 'keep')
+
+ggplot(phiad.plot.vals, aes(x = factor(phiad_cat), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Adult survival') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#view by phi1.true
+ggplot(rel.bias.vals, aes(x = factor(phi1.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True phi.1') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#view by phi1 category
+phi1.plot.vals <- rel.bias.vals %>%
+  group_by(lambda.scenario, simscenarios, variable, phi1_cat) %>%
+  dplyr::summarize(value = mean(value), .groups = 'keep')
+
+ggplot(phi1.plot.vals, aes(x = factor(phi1_cat), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Juv survival') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
 
 ```
+
+### Case study
+```{r load results from case study}
+
+load(file = here::here('results', 'EuringPoster', 'processedIPMcasestudy.RData'))
+
+#only decreasing population trend for now
+# unique(row.low.cs$sims) #sims = 1:25 (sim repetitions); 
+# unique(row.low.cs$scenario) #scenarios = 1:10 (phi/fec true vals);
+# unique(row.low.cs$simscenarios) #simscenarios = 1:9 (...?)
+row.low.cs <- do.call(rbind, lapply( ls(patt='lowout'), get) )
+#remove interim rds files
+rm(list=grep('highout|medout|lowout',ls(),value=TRUE,invert=FALSE))
+
+#list of survey detection scenarios
+scenarios.cs <- readRDS(file = here::here('data', 'scenarios.rds'))[1:9,]
+
+# rename and only take columns we need -- no lambdas here
+low.cs <- row.low.cs %>%
+  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`) %>%
+  dplyr::select(!starts_with('lambda'))
+
+#check; same as above for non-case study, right?
+#label scenarios 1:25, but only 10 taken below in inner_join()
+low.params.cs <- readRDS(file = here::here('data', 'low.lam.params.rds')) %>%
+  transform(scenario = 1:25)
+
+#add in simscenario numbering to merge below
+sim.scens.cs <- scenarios.cs %>%
+  transform(simscenarios = 1:9) %>%
+  dplyr::select(-c(lambda, n.viable.combinations))
+
+#combine to get true values integrated in results; 
+#only scenarios 1:10 out of 1:25 will be taken with inner_join
+low.cs.dat <- low.cs %>%
+  inner_join(low.params.cs, by = 'scenario', suffix = c('.obs', '.true')) %>%
+  merge(sim.scens.cs) %>%
+  transform(lambda.scenario = 'Decreasing')
+
+```
+
+```{r case study bias}
+
+###case study for Euring poster:
+# only a decreasing population trend, only 'L' nest detection
+
+#get posterior medians
+cs.meds <- low.cs.dat %>%
+  dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
+                            'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
+  reshape2::melt(id.vars = c('lambda.scenario', 'sims', 'scenario', 'simscenarios')) %>%
+  group_by(lambda.scenario, scenario, simscenarios, variable) %>%
+  summarize(median = median(value), .groups = 'keep') %>%
+  reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median')
+
+# bias
+rel.bias.cs <- cs.meds %>%
+  #calculate relative bias, mean across scenarios
+  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
+            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
+            fec.bias = (fec.obs-fec.true)/fec.true) %>%
+  dplyr::select('lambda.scenario', 'scenario', 'simscenarios', 
+                'phi1.bias', 'phiad.bias', 'fec.bias') %>%
+  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
+  group_by(lambda.scenario, simscenarios, variable) %>%
+  dplyr::summarize(mean.rel.bias = mean(value), .groups = 'keep') %>%
+  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias'),
+                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>% 
+  merge(sim.scens.cs, by = 'simscenarios') %>%
+  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H')),
+            det.MR = factor(det.MR, levels = c('L', 'M', 'H')))
+
+# plot; blrgh. Why don't these results change across detection levels?!
+ggplot(rel.bias.cs, aes(x = factor(det.MR), y = variable, fill = mean.rel.bias)) +
+  geom_tile(color = 'grey50') +
+  xlab('MR detection') + ylab('') +
+  facet_grid(~det.abund, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange1", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+  theme_minimal() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+##hmmm.... adult survival on it's own scale looks like what I would expect
+ggplot(rel.bias.cs %>% filter(variable == 'Adult survival'), 
+       aes(x = factor(det.MR), y = variable, fill = mean.rel.bias)) +
+  geom_tile(color = 'grey50') +
+  xlab('MR detection') + ylab('') +
+  facet_grid(~det.abund, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+  theme_minimal() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+ggplot(rel.bias.cs %>% filter(variable == 'Juv survival'), 
+       aes(x = factor(det.MR), y = variable, fill = mean.rel.bias)) +
+  geom_tile(color = 'grey50') +
+  xlab('MR detection') + ylab('') +
+  facet_grid(~det.abund, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+  theme_minimal() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+```

### 2021-04-28 scripts/EuringPosterFigures.Rmd
case study figures added, changed color palette to something bolder

@@
 
+#filter to see that adult survival bias does in fact vary with L/M/H detection
+ggplot(phiad.plot.vals %>% filter(variable == 'Adult survival'), 
+       aes(x = factor(phiad_cat), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Adult survival') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
 #view by phi1.true

### 2021-05-04 figures/lambdaTrends.R
edits to lambda figure

@@ -61,6 +61,11 @@ for(i in 1:dim(row.high)[1]) {
 
 #saveRDS(row.low, "row.low.RDS")
 
+## load RDS
+row.low <- readRDS(file = here::here('data', 'EuringPosterReg', 'row.low.rds'))
+row.med <- readRDS(file = here::here('data', 'EuringPosterReg', 'row.med.rds'))
+row.high <- readRDS(file = here::here('data', 'EuringPosterReg', 'row.high.rds'))
+
 # MUST FIX THIS
 toplot1 <- row.low %>%
   select(contains("geomean"), scenario, sims, simscenarios) %>%

### 2021-05-04 figures/lambdaTrendsPlot.R

@@ -14,7 +14,8 @@ library(wesanderson)
 pal <- rev(wes_palette("Zissou1", 3, type = "continuous"))
 
 # AEB - idea for version two of this plot
-ggplot(toplot1, aes(x = Year, y = value)) +
+## row low
+g1 <- ggplot(toplot1, aes(x = Year, y = value)) +
  stat_pointinterval(aes(color = simscenarios, fill = simscenarios), alpha = 0.5, .width = c(0.5, 0.95)) +
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 1.00, linetype = "solid", color = "black") +
@@ -26,7 +27,7 @@ ggplot(toplot1, aes(x = Year, y = value)) +
         axis.text.x = element_blank(),
         #axis.title.x = element_blank(),
         strip.text = element_blank(),
-        legend.position = "top",
+        legend.position = "none", # turn off legend
         legend.text = element_text(size = 12),
         legend.title = element_text(size = 12),
         plot.title.position = "plot",
@@ -35,6 +36,61 @@ ggplot(toplot1, aes(x = Year, y = value)) +
   coord_cartesian(clip = "off") +
   scale_color_manual(values = pal, name = "Detection\nlevel", labels = c("Low", "Medium", "High"))  +
   scale_fill_manual(values = pal, name = "Detection\nlevel", labels = c("Low", "Medium", "High"))
+
+## row med
+g2 <- ggplot(toplot2, aes(x = Year, y = value)) +
+  stat_pointinterval(aes(color = simscenarios, fill = simscenarios), alpha = 0.5, .width = c(0.5, 0.95)) +
+  geom_hline(yintercept = 1.00, linetype = "dashed", color = "black") +
+  geom_hline(yintercept = 1.00, linetype = "solid", color = "black") +
+  facet_grid(.~simscenarios) +
+  theme_minimal() +
+  theme(panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text = element_text(size = 12),
+        axis.text.x = element_blank(),
+        #axis.title.x = element_blank(),
+        strip.text = element_blank(),
+        legend.position = "none", # turn off legend
+        legend.text = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        plot.title.position = "plot",
+        axis.title=element_text(size=12)) +
[…]

### 2021-05-04 scripts/EuringPosterFigures.Rmd
for color palette tweaking

--- 2021-04-28 scripts/EuringPosterFigures.Rmd
@@
 
-# plot
+# plot; use me - overview
 ggplot(rel.bias, aes(x = factor(simscenarios), y = variable, fill = mean.rel.bias)) +
@@
 
+#use me; just one parameter -- fec by fec
+ggplot(rel.bias.vals %>% filter(variable == 'Fecundity'), aes(x = factor(fec.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True fecundity') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5,
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
 #view by fec category: have to re-summarize to get mean across fec/phi ifelse categories
@@
 
+#use me; just one parameter -- phi ad by phi ad
+ggplot(rel.bias.vals %>% filter(variable == 'Adult survival'), aes(x = factor(phiad.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Adult survival') + ylab('Relative bias') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Adult survival",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5,  0,0.5),
+                      n.breaks = 3,
+                      labels = c('Under estimate',  "0", 'overestimate')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
 #view by phiad category
@@
     scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
+                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+  theme_bw() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+#use me; just one parameter -- phi1
+ggplot(rel.bias.vals %>% filter(variable == 'Juv survival'), aes(x = factor(phi1.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Juv Survival') + ylab('Parameter') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
                       mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
@@
 ```
+
+### Case study
+```{r load results from case study}
+
+load(file = here::here('results', 'EuringPoster', 'processedIPMcasestudy.RData'))
+
+#only decreasing population trend for now
+# unique(row.low.cs$sims) #sims = 1:25 (sim repetitions); 
+# unique(row.low.cs$scenario) #scenarios = 1:10 (phi/fec true vals);
+# unique(row.low.cs$simscenarios) #simscenarios = 1:9 (...?)
+row.low.cs <- do.call(rbind, lapply( ls(patt='lowout'), get) )
+#remove interim rds files
+rm(list=grep('highout|medout|lowout',ls(),value=TRUE,invert=FALSE))
+
+#list of survey detection scenarios
+scenarios.cs <- readRDS(file = here::here('data', 'scenarios.rds'))[1:9,]
+
+# rename and only take columns we need -- no lambdas here
+low.cs <- row.low.cs %>%
+  dplyr::rename(phi1 = `mean.phi[1]`, phiad = `mean.phi[2]`) %>%
+  dplyr::select(!starts_with('lambda'))
+
+#check; same as above for non-case study, right?
+#label scenarios 1:25, but only 10 taken below in inner_join()
+low.params.cs <- readRDS(file = here::here('data', 'low.lam.params.rds')) %>%
+  transform(scenario = 1:25)
+
+#add in simscenario numbering to merge below
+sim.scens.cs <- scenarios.cs %>%
+  transform(simscenarios = 1:9) %>%
+  dplyr::select(-c(lambda, n.viable.combinations))
+
+#combine to get true values integrated in results; 
+#only scenarios 1:10 out of 1:25 will be taken with inner_join
+low.cs.dat <- low.cs %>%
+  inner_join(low.params.cs, by = 'scenario', suffix = c('.obs', '.true')) %>%
+  merge(sim.scens.cs) %>%
+  transform(lambda.scenario = 'Decreasing')
+
+```
+
+```{r case study bias}
+
+###case study for Euring poster:
+# only a decreasing population trend, only 'L' nest detection
+
+#get posterior medians
+cs.meds <- low.cs.dat %>%
+  dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
+                            'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
+  reshape2::melt(id.vars = c('lambda.scenario', 'sims', 'scenario', 'simscenarios')) %>%
+  group_by(lambda.scenario, scenario, simscenarios, variable) %>%
+  summarize(median = median(value), .groups = 'keep') %>%
+  reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median')
+
+# bias
+rel.bias.cs <- cs.meds %>%
+  #calculate relative bias, mean across scenarios
+  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
+            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
+            fec.bias = (fec.obs-fec.true)/fec.true) %>%
+  dplyr::select('lambda.scenario', 'scenario', 'simscenarios', 
+                'phi1.bias', 'phiad.bias', 'fec.bias') %>%
+  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'simscenarios')) %>%
+  group_by(lambda.scenario, simscenarios, variable) %>%
+  dplyr::summarize(mean.rel.bias = mean(value), .groups = 'keep') %>%
+  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias'),
+                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>% 
+  merge(sim.scens.cs, by = 'simscenarios') %>%
+  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H')),
+            det.MR = factor(det.MR, levels = c('L', 'M', 'H')))
+
+# plot; blrgh. Why don't these results change across detection levels?!
+ggplot(rel.bias.cs, aes(x = factor(det.MR), y = variable, fill = mean.rel.bias)) +
+  geom_tile(color = 'grey50') +
+  xlab('MR detection') + ylab('') +
+  facet_grid(~det.abund, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange1", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+  theme_minimal() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+##hmmm.... adult survival on it's own scale looks like what I would expect
+ggplot(rel.bias.cs %>% filter(variable == 'Adult survival'), 
+       aes(x = factor(det.MR), y = variable, fill = mean.rel.bias)) +
+  geom_tile(color = 'grey50') +
+  xlab('MR detection') + ylab('') +
+  facet_grid(~det.abund, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+  theme_minimal() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+ggplot(rel.bias.cs %>% filter(variable == 'Juv survival'), 
+       aes(x = factor(det.MR), y = variable, fill = mean.rel.bias)) +
+  geom_tile(color = 'grey50') +
+  xlab('MR detection') + ylab('') +
+  facet_grid(~det.abund, drop = T, scales = 'free_x') +
+  scale_fill_gradient2(name = "Relative bias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+  theme_minimal() +
+  theme(legend.position = 'top',
+             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+
+```

### 2021-05-04 scripts/EuringPosterFigures.Rmd
created figures for bias by parameter (adult survival, juv survival, fec)

@@
 library(RColorBrewer)
+library(colorspace)
 ```
@@
   xlab('True Juv survival') + ylab('Parameter') +
-  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x',
+             labeller = ) +
     scale_fill_gradient2(name = "Relative bias",
@@
 
+```
+
+```{r phiad}
+pal1 <- diverging_hcl(n = 7, palette = "Blue-Yellow 3")
+phiad <- rel.bias.vals %>% filter(variable == 'Adult survival')
+summary(phiad$value)
+
+p1 <- ggplot(rel.bias.vals %>% filter(variable == 'Adult survival'), 
+             aes(x = factor(phiad.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Adult Survival') + ylab("") +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative\nbias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "white", low = "#9FA2FF", high = "#BAAE00", midpoint = 0,
+                      breaks = c(-0.15, -0.05, 0, 0.05), n.breaks = 4, labels = c("-0.15", "-0.05", "0", "0.05")) +
+  theme_light() +
+  theme(legend.position = 'right',
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        strip.text.x = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA),
+        axis.text.x = element_text(angle = 90, vjust = 0.5)) +
+  ggtitle("Adult Survival")
+p1
+```
+
+```{r phijuv}
+pal2 <- diverging_hcl(n = 7, palette = "Green-Orange")
+phi1 <- rel.bias.vals %>% filter(variable == 'Juv survival')
+summary(phi1$value)
+
+p2 <- ggplot(rel.bias.vals %>% filter(variable == 'Juv survival'), 
+             aes(x = factor(phi1.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Juvenile Survival') + ylab("") +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative\nbias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "white", low = "#11C638", high = "#EF9708", midpoint = 0,
+                      breaks = c(-0.2, 0, 0.2, 0.5, 0.8), n.breaks = 5, 
+                      labels = c("-0.2", "0", "0.2", "0.5", "0.8")) +
+  theme_light() +
+  theme(legend.position = 'right',
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        strip.text.x = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA),
+        axis.text.x = element_text(angle = 90, vjust = 0.5)) +
+  ggtitle("Juvenile Survival")
+p2
+```
+
+```{r fec}
+pal3 <- diverging_hcl(n = 7, palette = "Cyan-Magenta")
+fec <- rel.bias.vals %>% filter(variable == 'Fecundity')
+summary(fec$value)
+
+p3 <- ggplot(rel.bias.vals %>% filter(variable == 'Fecundity'), 
+             aes(x = factor(fec.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('True Fecundity') + ylab("") +
+  facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+    scale_fill_gradient2(name = "Relative\nbias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0,
+                      breaks = c(-0.4, -0.2, 0, 0.2), n.breaks = 4, 
+                      labels = c("-0.4", "-0.2", "0", "0.2")) +
+  theme_light() +
+  theme(legend.position = 'right',
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        strip.text.x = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA),
+        axis.text.x = element_text(angle = 90, vjust = 0.5)) +
+  ggtitle("Fecundity")
+p3
+```
+
+```{r combo}
+library(patchwork)
+plot1 <- p1 / p2 / p3
+plot1
+
+## export out
+pdf(here("figures", "params_bias.pdf"), width = 12, height = 8)
+plot_grid(p1, p2, p3, nrow = 3)
+dev.off()
 ```

### 2021-08-16 scripts/current version/4 - run models/run_scenarios.R
processed csvs

@@ -84,7 +84,11 @@ for (i in 1:nrow(scenarios)) {
   tmp <- tmp[!is.na(tmp)]
   scenarios[i, "priority"] <- length(unique(tmp))
 }
-scenarios <- scenarios %>% arrange(priority) # save in prioritized order
+scenarios <- scenarios %>% arrange(priority) %>% # save in prioritized order
+  transform(simscenarios = 1:144)
+
+write.csv(scenarios, here::here('data', 'scenario_ID.csv'), row.names = F)
+
 which.prio.1 <- which(scenarios$priority == 1)
 which.prio.2 <- which(scenarios$priority == 2)
 which.prio.3 <- which(scenarios$priority == 3)

### 2021-08-17 README.Rmd
Draft of readme for scripts and data

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Something here about background?

# Scripts, current version
## 0 - preparing scenarios
### compute_time_calc.R
Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time. 
### generate_scenarios.R
Function 'getNviable' for finding the combinations of parameters that give a population growth rate ($\lambda$) within certain bounds, using the eigenvalue from the Leslie matrix. The output from this function produced the parameter scenarios used to simulate population trajectories, observation data from the trajectories, and fit the the IPM models.

## 1 - simulating data
### IPM_sim_2.0function.R
Script contains two functions for simulating data: 'simPopTrajectory' and 'simData'. 
The 'simPopTrajectory' function simulates the true population trajectory by individuals and through age classes, starting from the stable age distribution. Individual fates are monitored throughout, to use for generating each of the IPM datasets.

The 'simData' function uses the output from the 'simPopTrajectory' function to simulate the observation process for each data type. The true individual data is broken down into independent sets and each is subject to a observation error. Mark-resight, count, and reproductive success data are output from this function.

### simulateTrajectories.R

Script that pulls from the scenarios that were created in '0 - preparing scenarios' and the functions in IPM_sim_2.0function.R to simulate datasets for model fitting.

## 2 - models

### IPMinitvalues.R
Functions for putting data in the form required by the survival model.

### IPM_marray.R
Script contains 4 IPM NIMBLE models. 1.) full IPM model ('IPMmod') where all three datasets are included, 2.) model without productivity dataset ('nonests'), 3.) model without survival model ('nomr'), and 4.) abundance only model without productivity or survival ('abundonly').

## 3 - run models
### run_scenarios_helperFns.R
Script with functions and code to run NIMBLE models. Function 'marray' transforms the survival data capture histories to m-array format. MCMC settings set in this script, with functions for each of the 4 models. Each model has a run model function that defines constants, initial values, data, parameters to monitor, and code to build/run the models in NIMBLE.

### run_scenarios.R
Script that pulls from the scripts above to run models in parallel and output results.

## 4 - process results
### 01_load_and_process_data.R
Script that checks the Gelman-Rubin convergence statistic for each output MCMC sample, then thins the chains to reduce file size and places the output into categories based on the population growth rate, $\lambda$.

### 02_compute_geom_means.R
Computes the geometric means for $\lambda$ in each MCMC output.

# Data
Contains true parameter scenarios used to simulate population trajectories and observation data, all viable parameter cominations in files '-.lam.combos.RDS' and the selected parameter combinations used to simulate data are in files '-.lam.params.RDS' (where "-" is low, med, or high). Folders labeled '-Trajectories' contain the simulated population trajectories for each of the scenarios found in '-.lam.params.RDS'. Each simulated trajectory is labeled with the scenario and the simulation number for that scenario. For example, in the highTrajectory folder: highpopTraj-1-2.RDS corresponds to the first parameter scenario in high.lam.param.RDS and is the second population trajectory simulated using those parameter values. 
File 'true.vals.csv' is a file for the true parameter values used in each scenario. 'scenario_ID.csv' tracks the combinations of detection parameters and IPM models.

### 2021-08-17 README.Rmd
fixes

@@
 ### compute_time_calc.R
-Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time. 
+Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time for each model.
+
 ### generate_scenarios.R

### 2021-08-17 README.Rmd
couple of edits

@@
 ### generate_scenarios.R
+
 Function 'getNviable' for finding the combinations of parameters that give a population growth rate ($\lambda$) within certain bounds, using the eigenvalue from the Leslie matrix.
-The output from this function produced the parameter scenarios used to simulate population trajectories, observation data from the trajectories, and fit the the IPM models.
+The output from this function produces the parameter scenarios used to simulate population trajectories, observation data from the trajectories, and fit the the IPM models.
 
@@
 Script contains two functions for simulating data: 'simPopTrajectory' and 'simData'. 
-The 'simPopTrajectory' function simulates the true population trajectory by individuals and through age classes, starting from the stable age distribution.
-Individual fates are monitored throughout, to use for generating each of the IPM datasets.
+The 'simPopTrajectory' function simulates the true population trajectory by individual and age class, starting from the stable age distribution.
+Individual fates are retained throughout total number of years.
 
 The 'simData' function uses the output from the 'simPopTrajectory' function to simulate the observation process for each data type.
-The true individual data is broken down into independent sets and each is subject to a observation error.
+The true individual data is broken down into independent sets, each is subject to observation error.
 Mark-resight, count, and reproductive success data are output from this function.
@@
 
-Script that pulls from the scenarios that were created in '0 - preparing scenarios' and the functions in IPM_sim_2.0function.R to simulate datasets for model fitting.
+Script that pulls from the scenarios that were created in '0 - preparing scenarios' and the functions in IPM_sim_2.0function.R to simulate observation data sets for model fitting.
 
@@
 Script contains 4 IPM NIMBLE models.
-1.) full IPM model ('IPMmod') where all three datasets are included, 2.) model without productivity dataset ('nonests'), 3.) model without survival model ('nomr'), and 4.) abundance only model without productivity or survival ('abundonly').
+
+ 1. full IPM model ('IPMmod') where all three datasets are included
+ 2. model without productivity dataset ('nonests')
+ 3. model without survival model ('nomr')
+ 4. abundance only model without productivity or survival ('abundonly')
 
@@
 Function 'marray' transforms the survival data capture histories to m-array format.
-MCMC settings set in this script, with functions for each of the 4 models.
+MCMC settings in this script, with functions for each of the 4 models.
 Each model has a run model function that defines constants, initial values, data, parameters to monitor, and code to build/run the models in NIMBLE.
@@
 # Data
-Contains true parameter scenarios used to simulate population trajectories and observation data, all viable parameter cominations in files '-.lam.combos.RDS' and the selected parameter combinations used to simulate data are in files '-.lam.params.RDS' (where "-" is low, med, or high).
+Contains true parameter scenarios used to simulate population trajectories and observation data, all viable parameter combinations in files '-.lam.combos.RDS' and the selected parameter combinations used to simulate data are in files '-.lam.params.RDS' *(where "-" is low, med, or high)*.
 Folders labeled '-Trajectories' contain the simulated population trajectories for each of the scenarios found in '-.lam.params.RDS'.
 Each simulated trajectory is labeled with the scenario and the simulation number for that scenario.
-For example, in the highTrajectory folder: highpopTraj-1-2.RDS corresponds to the first parameter scenario in high.lam.param.RDS and is the second population trajectory simulated using those parameter values.
-File 'true.vals.csv' is a file for the true parameter values used in each scenario. 'scenario_ID.csv' tracks the combinations of detection parameters and IPM models.
+For example, in the 'highTrajectory' folder: 'highpopTraj-1-2.RDS' corresponds to the first parameter scenario in 'high.lam.param.RDS' and is the second population trajectory simulated using those parameter values.
+
+File 'true.vals.csv' is the true parameter values used in each scenario. 'scenario_ID.csv' tracks the combinations of detection probabilities and IPM models.

### 2021-08-17 scripts/PaperFigures.Rmd
adding detection probs to figures

---
title: "EuringPosterFigures"
author: ""
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = 'asis', fig.width = 4, fig.height = 3.5)

library(tidyr)
library(dplyr)
library(cowplot)
library(ggplot2)
library(coda)
library(captioner)
library(knitr)
library(reshape2)
library(here)
library(RColorBrewer)
library(colorspace)
library(ggh4x)

rainbow2 <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")

```

```{r data load}

scenarios <- read.csv(file = here::here('data', 'scenario_ID.csv'), header = T, stringsAsFactors = F)

low <- read.csv(file = here::here('results', 'lowout.csv')) %>%
  dplyr::select(-X) %>%
  dplyr::rename(phi1 = `mean.phi.1.`, phiad = `mean.phi.2.`)
med <- read.csv(file = here::here('results', 'medout.csv'))  %>%
  dplyr::select(-X) %>%
  dplyr::rename(phi1 = `mean.phi.1.`, phiad = `mean.phi.2.`)
high <- read.csv(file = here::here('results', 'highout.csv'))  %>%
  dplyr::select(-X) %>%
  dplyr::rename(phi1 = `mean.phi.1.`, phiad = `mean.phi.2.`)

low.params <- readRDS(file = here::here('data', 'low.lam.params.rds')) %>%
  transform(scenario = 1:25)

med.params <- readRDS(file = here::here('data', 'med.lam.params.rds')) %>%
  transform(scenario = 1:25)

high.params <- readRDS(file = here::here('data', 'high.lam.params.rds')) %>%
  transform(scenario = 1:25)

#combine to get true values integrated in results
low.dat <- low %>%
  inner_join(low.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Decreasing')

med.dat <- med %>%
  inner_join(med.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Stable')

high.dat <- high %>%
  inner_join(high.params, by = 'scenario', suffix = c('.obs', '.true')) %>%
  transform(lambda.scenario = 'Increasing')

all.dat <- bind_rows(low.dat, med.dat, high.dat)

# true_vals <- all.dat %>%
#   distinct(lambda.scenario, scenario, phiad.true, fec.true, phi1.true) %>%
#   transform(phiad.true = round(phiad.true,1),
#             phi1.true = round(phi1.true,1), fec.true = round(fec.true,1))

# write.csv(true_vals, file = here::here('data', 'true.vals.csv'), row.names = F)

```

```{r bias}

#get posterior medians
# all.meds <- all.dat %>%
#   dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
#                 'p.surv', 'mean.p',
#                             'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
#   reshape2::melt(id.vars = c('lambda.scenario', 'sims', 'scenario', 'simscenarios')) %>%
#   group_by(lambda.scenario, scenario, simscenarios, variable) %>%
#   summarize(median = median(value), .groups = 'keep') %>%
#   reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median') %>%
#   merge(scenarios %>% dplyr::select(-c(n.viable.combinations, priority)), by = c('simscenarios'))
# 
# 
# write.csv(all.meds, here::here('figures', 'Processed csvs', 'all.meds.csv'), row.names = F)

all.meds <- read.csv(file = here::here('results', 'Processed csvs', 'all.meds.csv'), 
                     header = T, stringsAsFactors = F)

##bias
rel.bias <- all.meds %>%
  transform(p.surv.true = ifelse(det.abund == 'L', 0.3))
  #calculate relative bias, mean across scenarios
  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
            fec.bias = (fec.obs-fec.true)/fec.true) %>%
  dplyr::select('lambda.scenario', 'scenario',
                'phi1.bias', 'phiad.bias', 'fec.bias', 'det.MR', 'det.abund', 'det.prod') %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'det.MR', 'det.abund', 'det.prod')) %>%
  group_by(lambda.scenario, det.MR, det.abund, det.prod, variable) %>%
  dplyr::summarize(bias = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias'),
                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>%
  transform(lambda.scenario = factor(lambda.scenario,
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(det.MR = factor(det.MR, levels = c('L', 'M', 'H'))) %>%
  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H'))) %>%
  transform(det.prod = factor(det.prod, levels = c('L', 'M', 'H'))) %>%
  transform(missing.MR = ifelse(is.na(det.MR), 1, 0),
            missing.prod = ifelse(is.na(det.prod), 1, 0)) %>%
  transform(num.miss = missing.MR + missing.prod) %>%
  transform(dataset = ifelse(is.na(det.MR)&!is.na(det.prod), 'Counts+Prod', 
                             ifelse(!is.na(det.MR)&is.na(det.prod), 'Counts+MR',
                                    ifelse(is.na(det.MR)&is.na(det.prod), 'Counts', 'Full'))))

# 
# rel.bias$lambda.scenario <- factor(rel.bias$lambda.scenario, levels = c("Decreasing", "Stable", "Increasing"),
#                                 ordered = TRUE, labels = c(expression(paste("Decreasing ", lambda)), 
#                                                            expression(paste("Stable ", lambda)), 
#                                                            expression(paste("Increasing ", lambda))))

#all facets shown - visually interesting, but not using
ggplot(rel.bias, aes(x = factor(det.abund), y = variable, fill = bias)) +
  geom_tile(color = 'grey50') +
  xlab('Det.abund') + ylab('') +
  facet_grid(det.prod + num.miss ~ lambda.scenario + det.MR, 
             drop = T, scales = 'free_x', labeller = label_both) +
  scale_fill_gradient2(name = "Relative\nbias",
                     mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

#only one dataset missing
# ggplot(rel.bias %>% filter(num.miss == 1), aes(x = factor(det.abund), y = variable, fill = bias)) +
#   geom_tile(color = 'grey50') +
#   xlab('') + ylab('') +
#   facet_grid(det.MR + det.prod ~ lambda.scenario, 
#              drop = T, scales = 'free_x', labeller = label_both) +
#   scale_fill_gradient2(name = "Relative\nbias",
#                       # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
#                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0) +
#   theme(panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank())

#dotplot equivalent
# ggplot(rel.bias %>% filter(num.miss == 1), aes(x = det.abund, y = bias, 
#                                                    col = lambda.scenario, group = lambda.scenario)) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('det.abund') + ylab('Relative bias') +
#   facet_grid(variable~det.MR+det.prod, scales = 'free') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1))

###average over two layers of detection (det.MR and det.prod)
rel.bias.few <- rel.bias %>%
  group_by(variable, det.abund, lambda.scenario, dataset) %>%
  dplyr::summarize(bias = mean(bias), .groups = 'keep') %>%
  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H'))) %>%
  transform(lambda.scenario = factor(lambda.scenario, 
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(dataset = factor(dataset, levels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'),
                             labels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts')))

#use me fig 1
ggplot(rel.bias.few, aes(x = factor(det.abund), y = variable, fill = bias)) +
  geom_tile(color = 'grey50') +
  xlab('Count detection') + ylab('') +
  facet_grid(dataset ~ lambda.scenario, drop = T, scales = 'free_x', labeller = label_parsed) +
  scale_fill_gradient2(name = "Relative bias",
                     # mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0) +
                       mid = "white", low = rainbow2[2], high = rainbow2[6], midpoint = 0) +
                      # breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
 # scale_x_discrete(labels = c(expression(paste("low ", italic(p))), 
 #                             expression(paste("med ", italic(p))), 
 #                             expression(paste("high ", italic(p))))) +
  theme_light() +
  theme(legend.position = 'top',
        # axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold", color = "black"),
        strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 0, vjust = 1.5),
        # ggh4x.facet.nestline = element_line(size = 0.5), #what does this line do?
        panel.border = element_rect(color = "black", fill = NA),  
        panel.spacing.x = unit(0.75, "line"))

## dot plot
# ggplot(rel.bias.few, aes(x = det.abund, y = bias, col = factor(dataset), group = factor(dataset))) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('Count detection') + ylab('Relative bias') +
#   facet_grid(variable~lambda.scenario, scales = 'free') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[-c(1,4)], name = '')

ggplot(rel.bias.few, aes(x = det.abund, y = bias, col = factor(variable), group = factor(variable))) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('Count detection') + ylab('Relative bias') +
  facet_grid(dataset~lambda.scenario, scales = 'free_x') +
  theme_bw() +
  theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2[-c(1,4)], name = '')

```

```{r bias across life history scenarios}

##Same as above but looking at the variability across data-generating true values for phiad, phi1, fec

#get posterior medians; same as above except keep 'scenario' in group_by
# all.meds.sc <- all.dat %>%
#   dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
#                             'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
#   reshape2::melt(id.vars = c('sims', 'scenario', 'simscenarios', 'lambda.scenario')) %>%
#   group_by(lambda.scenario, scenario, simscenarios, variable) %>%
#   summarize(median = median(value), .groups = 'keep') %>%
#   reshape2::dcast(lambda.scenario + scenario + simscenarios ~ variable, value.var = 'median') %>%
#   merge(scenarios %>% dplyr::select(-c(n.viable.combinations, priority)), by = c('simscenarios')) 
# 
# write.csv(all.meds.sc, here::here('results', 'Processed csvs', 'all.meds.sc.csv'), row.names = F)

all.meds.sc <- read.csv(file = here::here('results', 'Processed csvs', 'all.meds.sc.csv'), 
                     header = T, stringsAsFactors = F)

#just bias; same as above, keep 'scenario'
rel.bias.sc <- all.meds.sc %>%
  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
            fec.bias = (fec.obs-fec.true)/fec.true) %>%
  dplyr::select('lambda.scenario', 'scenario', 
                'phi1.bias', 'phiad.bias', 'fec.bias', 'det.MR', 'det.abund', 'det.prod') %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'det.MR', 'det.abund', 'det.prod')) %>%
  transform(lambda.scenario = factor(lambda.scenario,
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(det.MR = factor(det.MR, levels = c('L', 'M', 'H'))) %>%
  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H'))) %>%
  transform(det.prod = factor(det.prod, levels = c('L', 'M', 'H'))) %>%
  transform(missing.MR = ifelse(is.na(det.MR), 1, 0),
            missing.prod = ifelse(is.na(det.prod), 1, 0)) %>%
  transform(num.miss = missing.MR + missing.prod) %>%
  transform(dataset = ifelse(is.na(det.MR)&!is.na(det.prod), 'Counts+Prod', 
                             ifelse(!is.na(det.MR)&is.na(det.prod), 'Counts+MR',
                                    ifelse(is.na(det.MR)&is.na(det.prod), 'Counts', 'Full'))))

#load saved true values
true_vals <- read.csv(file = here::here('data', 'true.vals.csv'), header = T, stringsAsFactors = F)

#for categories: 
fec_lims <- quantile(true_vals$fec.true, probs = c(0.33, 0.7), names = F)
phiad_lims <- quantile(true_vals$phiad.true, probs = c(0.33, 0.7), names = F)
phi1_lims <- quantile(true_vals$phi1.true, probs = c(0.33, 0.7), names = F)

#merge back onto bias df so can have something to name/view the 'scenarios' 
rel.bias.dem <- rel.bias.sc %>%
  merge(true_vals, by = c('lambda.scenario', 'scenario')) %>%
  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias'),
                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>% 
  transform(lambda.scenario = factor(lambda.scenario, 
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(fec_cat = ifelse(fec.true < fec_lims[1], 'L', ifelse(fec.true > fec_lims[2], 'H', 'M')),
            phiad_cat = ifelse(phiad.true < phiad_lims[1], 'L', ifelse(phiad.true > phiad_lims[2], 'H', 'M')),
            phi1_cat = ifelse(phi1.true < phi1_lims[1], 'L', ifelse(phi1.true > phi1_lims[2], 'H', 'M'))) %>%
  transform(fec_cat = factor(fec_cat, levels = c('L', 'M', 'H')),
            phiad_cat = factor(phiad_cat, levels = c('L', 'M', 'H')),
            phi1_cat = factor(phi1_cat, levels = c('L', 'M', 'H'))) %>%
  transform(lambda.scenario = factor(lambda.scenario, 
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(dataset = factor(dataset, levels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'),
                             labels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts')))

# #view by fec.true vals; 
# ggplot(rel.bias, aes(x = factor(fec.true), y = variable, fill = value)) +
#   geom_tile(color = 'grey50') +
#   xlab('True fecundity') + ylab('Parameter') +
#   facet_grid(~lambda.scenario+det.abund, drop = T, scales = 'free_x') +
#     scale_fill_gradient2(name = "Relative bias",
#                       # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
#                       mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
#                       breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
#                       labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
#   theme_bw() +
#   theme(legend.position = 'top',
#              strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
#         axis.text.x = element_text(angle = 90, vjust = 0.5))

#facet by both fec and juv true vals
plot.vals <- rel.bias.dem %>%
  group_by(variable, fec_cat, phi1_cat, dataset) %>%
  dplyr::summarize(value = mean(value), .groups = 'keep')

#use me fig 2
ggplot(plot.vals, aes(x = factor(fec_cat), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('Fecundity') + ylab('') +
  facet_grid(dataset~phi1_cat, drop = T, scales = 'free_x') +
    scale_fill_gradient2(name = "Relative bias",
                      mid = "white", low = rainbow2[2], high = rainbow2[6], midpoint = 0,
                      # breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
  theme_light() +
  theme(legend.position = 'top',
        # axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(color = "black"),
        strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 0, vjust = 1.5),
        # ggh4x.facet.nestline = element_line(size = 0.5), #what does this line do?
        panel.border = element_rect(color = "black", fill = NA),  
        panel.spacing.x = unit(0.75, "line"))

## dot plot
# ggplot(plot.vals, aes(x = fec_cat, y = value, col = factor(dataset), group = factor(dataset))) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('True fecundity') + ylab('Relative bias') +
#   facet_grid(variable~phi1_cat, scales = 'free') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[-c(1,4)], name = '')

ggplot(plot.vals, aes(x = fec_cat, y = value, col = factor(variable), group = factor(variable))) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('True fecundity') + ylab('Relative bias') +
  facet_grid(dataset~phi1_cat, scales = 'free_x') +
  theme_bw() +
  theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2[-c(1,4)], name = '')

```

```{r rmse}

#for the love of buddha, someone please check this calculation of rmse
rmse.vals <- all.meds %>%
  transform(fec.rmse = (fec.obs-fec.true)^2,
            phiad.rmse = (phiad.obs-phiad.true)^2,
            phi1.rmse = (phi1.obs-phi1.true)^2) %>%
  dplyr::select(lambda.scenario, scenario, fec.rmse, phiad.rmse, phi1.rmse, 
                det.MR, det.abund, det.prod) %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'det.MR', 'det.abund', 'det.prod')) %>%
  group_by(lambda.scenario, det.MR, det.abund, det.prod, variable) %>%
  dplyr::summarize(mean.rmse = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiad.rmse', 'phi1.rmse', 'fec.rmse'),
                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>%
  transform(lambda.scenario = factor(lambda.scenario,
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(det.MR = factor(det.MR, levels = c('L', 'M', 'H'))) %>%
  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H'))) %>%
  transform(det.prod = factor(det.prod, levels = c('L', 'M', 'H'))) %>%
  transform(missing.MR = ifelse(is.na(det.MR), 1, 0),
            missing.prod = ifelse(is.na(det.prod), 1, 0)) %>%
  transform(num.miss = missing.MR + missing.prod) %>%
  transform(dataset = ifelse(is.na(det.MR)&!is.na(det.prod), 'Counts+Prod', 
                             ifelse(!is.na(det.MR)&is.na(det.prod), 'Counts+MR',
                                    ifelse(is.na(det.MR)&is.na(det.prod), 'Counts', 'Full'))))

#all facets shown
ggplot(rmse.vals, aes(x = factor(det.abund), y = variable, fill = mean.rmse)) +
  geom_tile(color = 'grey50') +
  xlab('Det.abund') + ylab('') +
  facet_grid(det.prod + num.miss ~ lambda.scenario + det.MR, 
             drop = T, scales = 'free_x', labeller = label_both) +
  scale_fill_gradient2(name = "RMSE",
                     mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

#the dotplot version - very hard to interpret
ggplot(rmse.vals, aes(x = det.abund, y = mean.rmse, col = lambda.scenario, group = lambda.scenario)) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('det.abund') + ylab('RMSE') +
  facet_grid(variable+det.MR~det.prod+num.miss, scales = 'free') +
  theme_bw() +
  theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1))

###average over two layers of detection (det.MR and det.prod)
rmse.few <- rmse.vals %>%
  group_by(variable, det.abund, lambda.scenario, dataset) %>%
  dplyr::summarize(rmse = mean(mean.rmse), .groups = 'keep') %>%
  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H'))) %>%
  transform(lambda.scenario = factor(lambda.scenario, 
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(dataset = factor(dataset, levels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'),
                             labels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts')))

#use me fig 3
ggplot(rmse.few, aes(x = factor(det.abund), y = variable, fill = rmse)) +
  geom_tile(color = 'grey50') +
  xlab('Count detection') + ylab('') +
  facet_grid(dataset ~ lambda.scenario, drop = T, scales = 'free_x', labeller = label_parsed) +
  scale_fill_gradient2(name = "RMSE",
                       mid = "white", high = rainbow2[2], midpoint = 0) +
                      # breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
 # scale_x_discrete(labels = c(expression(paste("low ", italic(p))), 
 #                             expression(paste("med ", italic(p))), 
 #                             expression(paste("high ", italic(p))))) +
  theme_light() +
  theme(legend.position = 'top',
        # axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold", color = "black"),
        strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 0, vjust = 1.5),
        # ggh4x.facet.nestline = element_line(size = 0.5), #what does this line do?
        panel.border = element_rect(color = "black", fill = NA),  
        panel.spacing.x = unit(0.75, "line"))

## dot plot
# ggplot(rmse.few, aes(x = det.abund, y = bias, col = factor(dataset), group = factor(dataset))) +
#   geom_point() + geom_line() +
#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
#   xlab('Count detection') + ylab('Relative bias') +
#   facet_grid(variable~lambda.scenario, scales = 'free') +
#   theme_bw() +
#   theme(legend.position = 'top',
#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
#   scale_color_manual(values = rainbow2[-c(1,4)], name = '')

ggplot(rmse.few, aes(x = det.abund, y = rmse, col = factor(variable), group = factor(variable))) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('Count detection') + ylab('RMSE') +
  facet_grid(dataset~lambda.scenario, scales = 'free') +
  theme_bw() +
  theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2[-c(1,4)], name = '')

########data-generating values
rmse.vals.sc <- all.meds %>%
  transform(fec.rmse = (fec.obs-fec.true)^2,
            phiad.rmse = (phiad.obs-phiad.true)^2,
            phi1.rmse = (phi1.obs-phi1.true)^2) %>%
  dplyr::select(lambda.scenario, scenario, fec.rmse, phiad.rmse, phi1.rmse, 
                det.MR, det.abund, det.prod) %>%
  reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'det.MR', 'det.abund', 'det.prod')) %>%
  group_by(lambda.scenario, scenario, det.MR, det.abund, det.prod, variable) %>%
  dplyr::summarize(mean.rmse = mean(value), .groups = 'keep') %>%
  transform(variable = factor(variable, levels = c('phiad.rmse', 'phi1.rmse', 'fec.rmse'),
                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>%
  transform(lambda.scenario = factor(lambda.scenario,
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(det.MR = factor(det.MR, levels = c('L', 'M', 'H'))) %>%
  transform(det.abund = factor(det.abund, levels = c('L', 'M', 'H'))) %>%
  transform(det.prod = factor(det.prod, levels = c('L', 'M', 'H'))) %>%
  transform(missing.MR = ifelse(is.na(det.MR), 1, 0),
            missing.prod = ifelse(is.na(det.prod), 1, 0)) %>%
  transform(num.miss = missing.MR + missing.prod) %>%
  transform(dataset = ifelse(is.na(det.MR)&!is.na(det.prod), 'Counts+Prod', 
                             ifelse(!is.na(det.MR)&is.na(det.prod), 'Counts+MR',
                                    ifelse(is.na(det.MR)&is.na(det.prod), 'Counts', 'Full'))))

rmse.dem <- rmse.vals.sc %>%
  merge(true_vals, by = c('lambda.scenario', 'scenario')) %>%
  # transform(variable = factor(variable, levels = c('phiad.rmse', 'phi1.rmse', 'fec.rmse'),
  #                             labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>% 
  transform(lambda.scenario = factor(lambda.scenario, 
                                     levels = c("Decreasing", "Stable", "Increasing"))) %>%
  transform(fec_cat = ifelse(fec.true < fec_lims[1], 'L', ifelse(fec.true > fec_lims[2], 'H', 'M')),
            phiad_cat = ifelse(phiad.true < phiad_lims[1], 'L', ifelse(phiad.true > phiad_lims[2], 'H', 'M')),
            phi1_cat = ifelse(phi1.true < phi1_lims[1], 'L', ifelse(phi1.true > phi1_lims[2], 'H', 'M'))) %>%
  transform(fec_cat = factor(fec_cat, levels = c('L', 'M', 'H')),
            phiad_cat = factor(phiad_cat, levels = c('L', 'M', 'H')),
            phi1_cat = factor(phi1_cat, levels = c('L', 'M', 'H'))) %>%
  transform(dataset = factor(dataset, levels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'),
                             labels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts')))

#facet by both fec and juv true vals
plot.vals.rmse <- rmse.dem %>%
  group_by(variable, fec_cat, phi1_cat, dataset) %>%
  dplyr::summarize(value = mean(mean.rmse), .groups = 'keep')

#use me fig 3
ggplot(plot.vals.rmse, aes(x = factor(fec_cat), y = factor(variable), fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('True fecundity') + ylab('') +
  facet_grid(dataset~phi1_cat, drop = T, scales = 'free_x') +
    scale_fill_gradient2(name = "RMSE",
                      mid = "white", high = rainbow2[2], midpoint = 0,
                      # breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
  theme_light() +
  theme(legend.position = 'top',
        # axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(color = "black"),
        strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 0, vjust = 1.5),
        # ggh4x.facet.nestline = element_line(size = 0.5), #what does this line do?
        panel.border = element_rect(color = "black", fill = NA),  
        panel.spacing.x = unit(0.75, "line"))

## dot plot
ggplot(plot.vals.rmse, aes(x = fec_cat, y = value, col = factor(variable), group = factor(variable))) +
  geom_point() + geom_line() +
  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
  xlab('True fecundity') + ylab('RMSE') +
  facet_grid(dataset~phi1_cat, scales = 'free_x') +
  theme_bw() +
  theme(legend.position = 'top',
             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
  scale_color_manual(values = rainbow2[-c(1,4)], name = '')

```

<!-- Euring figures -->
```{r phiad}
pal1 <- diverging_hcl(n = 7, palette = "Blue-Yellow 3")
phiad <- rel.bias.vals %>% filter(variable == 'Adult survival') 
summary(phiad$value) # look at range for scale

## fix faceted labels
rel.bias.vals$lambda.scenario <- factor(rel.bias.vals$lambda.scenario, levels = c("Decreasing", "Stable", "Increasing"),
                                ordered = TRUE, labels = c(expression(paste("Decreasing ", lambda)), 
                                                           expression(paste("Stable ", lambda)), 
                                                           expression(paste("Increasing ", lambda))))
rel.bias.vals$simscenarios <- factor(rel.bias.vals$simscenarios, levels = c("L", "M", "H"),
                                ordered = TRUE, labels = c(expression(paste("low ", italic(p))), 
                                                           expression(paste("med ", italic(p))), 
                                                           expression(paste("high ", italic(p)))))

p1 <- ggplot(rel.bias.vals %>% filter(variable == 'Adult survival'),
             aes(x = factor(phiad.true), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('\nTrue Adult Survival') + ylab("") +
  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
               nest_line = TRUE) +
    scale_fill_gradient2(name = "Relative\nbias",
                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
                      # mid = "white", low = "#9FA2FF", high = "#BAAE00", midpoint = 0,
                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
                      breaks = c(-0.15, -0.05, 0, 0.05), n.breaks = 4, labels = c("-0.15", "-0.05", "0", "0.05")) +
  theme_light() +
  theme(legend.position = 'right',
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        strip.text.x = element_text(face = "bold", color = "black"),
        strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 1.5),
        ggh4x.facet.nestline = element_line(size = 0.5),
        panel.border = element_rect(color = "black", fill = NA),  
        panel.spacing.x = unit(0.75, "line")) +
        # panel.spacing = unit(1, "lines")) +
  ggtitle("Adult Survival")
p1
```

```{r phijuv}
pal2 <- diverging_hcl(n = 7, palette = "Green-Orange")
phi1 <- rel.bias.vals %>% filter(variable == 'Juv survival')
summary(phi1$value)

p2 <- ggplot(rel.bias.vals %>% filter(variable == 'Juv survival'), 
             aes(x = factor(phi1.true), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('\nTrue Juvenile Survival') + ylab("") +
  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
               nest_line = TRUE) +
  scale_fill_gradient2(name = "Relative\nbias",
                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
                      # mid = "white", low = "#11C638", high = "#EF9708", midpoint = 0,
                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
                      breaks = c(-0.2, 0, 0.2, 0.5, 0.8), n.breaks = 5, 
                      labels = c("-0.2", "0", "0.2", "0.5", "0.8")) +
  theme_light() +
 theme(legend.position = 'right',
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        strip.text.x = element_text(face = "bold", color = "black"),
        strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 1.5),
        ggh4x.facet.nestline = element_line(size = 0.5),
        panel.border = element_rect(color = "black", fill = NA),  
        panel.spacing.x = unit(0.75, "line")) +
        # panel.spacing = unit(1, "lines")) +
  ggtitle("Juvenile Survival")
p2
```

```{r fec}
pal3 <- diverging_hcl(n = 7, palette = "Cyan-Magenta")
fec <- rel.bias.vals %>% filter(variable == 'Fecundity')
summary(fec$value)

p3 <- ggplot(rel.bias.vals %>% filter(variable == 'Fecundity'), 
             aes(x = factor(fec.true), y = variable, fill = value)) +
  geom_tile(color = 'grey50') +
  xlab('\nTrue Fecundity') + ylab("") +
  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
               nest_line = TRUE) +
  scale_fill_gradient2(name = "Relative\nbias",
                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
                      # mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0,
                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
                      breaks = c(-0.4, -0.2, 0, 0.2), n.breaks = 4, 
                      labels = c("-0.4", "-0.2", "0", "0.2")) +
  theme_light() +
  theme(legend.position = 'right',
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
        strip.text.x = element_text(face = "bold", color = "black"),
        strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 1.5),
        ggh4x.facet.nestline = element_line(size = 0.5),
        panel.border = element_rect(color = "black", fill = NA),  
        panel.spacing.x = unit(0.75, "line")) +
        # panel.spacing = unit(1, "lines")) +
  ggtitle("Fecundity")
p3
```

```{r combo}
## export out
pdf(here("figures", "params_bias.pdf"), width = 12, height = 8)
plot_grid(p3, p2, p1, nrow = 3)
dev.off()

## try with overview plot
all <- ov + (p3/p2/p1)
ggsave(filename = here("figures", "all_bias.pdf"), plot = all, width = 22, height = 12)
#use # for spaces with missing plots
# layout_string <- 'A#BB
#                   ##CC
#                   ##DD'
# junk <- ov + (p3/p2/p1) + plot_layout(design=layout_string,
#                               widths=c(3,1,5,6),
#                               heights=c(5,3))
```

### 2021-08-17 figures/lambdaTrendsPlot.R
lambda trends

@@ -91,7 +91,8 @@ toplot <- bind_rows(toplot1, toplot2, toplot3) %>%
   mutate(Quantile = str_remove(Quantile, "\\%")) %>% 
   mutate(Quantile = paste("X", Quantile, sep = "")) %>% 
   reshape2::dcast(dataset + Year +  det.MR + det.prod + det.abund   + lambda ~ Quantile, value.var = "value") %>% 
-  filter(Year %in% c(1:5, 10)) %>% 
+  mutate(Year = Year + 1) %>% 
+  filter(Year %in% c(2:6, 15)) %>% 
   mutate(Year = factor(Year)) %>% 
   mutate(det.abund = factor(det.abund, levels = c("L", "M", "H"))) %>% 
   mutate(det.prod = factor(det.prod, levels = c("L", "M", "H"))) %>% 
@@ -99,7 +100,11 @@ toplot <- bind_rows(toplot1, toplot2, toplot3) %>%
   transform(lambda = factor(lambda, levels = c("L", "M", "H"), 
                             labels = c("Decreasing", "Stable", "Increasing"))) %>%
   transform(dataset = factor(dataset, levels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'),
-                             labels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts')))
+                             labels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'))) %>% 
+  mutate(intercept = case_when(
+    lambda == "Decreasing" ~ 0.95, 
+    lambda == "Stable" ~ 1,
+    lambda == "Increasing" ~ 1.05))
 
 
 pdf(here("figures", "lambdaTrends.pdf"), width = 12, height = 8)
@@ -108,8 +113,8 @@ ggplot(toplot) +
   geom_point(aes(x = Year, y = X50, col = det.abund, group = det.abund), position = position_dodge(width = 0.5)) +
   #geom_linerange(aes(ymin = X2.5, ymax = X97.5, x = Year), position = position_dodge(width = 0.5)) +
   #geom_hline(aes(yintercept = 1.0), linetype = 'dotted') +
-  xlab('Year') + ylab('Lambda') + 
-  ylim(0.87, 1.14) +
+  xlab('Years') + ylab('Lambda') + 
+  #ylim(0.87, 1.14) +
   facet_grid(dataset ~ lambda, scales = 'free') +
   theme_bw() +
[…]

### 2021-08-17 ARCHIVE/documents/ARCHIVE/DraftReport.Rmd
repo reorganization

--- 2020-12-14 documents/DraftReport.Rmd
@@
   merge(true_vals %>% dplyr::select(scenario, variable, Truth), by = c('scenario', 'variable')) %>%
-  transform(rmse.temp = (value-Truth)^2) %>%
+  transform(rmse.temp = (value-Truth)^2) %>% #simulation level
   group_by(Data, scenario, variable) %>%
-  dplyr::summarize(rmse = sqrt(mean(rmse.temp)))
+  dplyr::summarize(rmse = sqrt(mean(rmse.temp))) #scenario level
 

### 2021-08-17 README.Rmd
repo reorganization

@@
 
-Something here about background?
+Understanding the processes that control population dynamics in wild populations is a critical component of conservation and management but is frequently limited by missing demographic information, uncertainty, and temporally or spatially misaligned datasets (Schaub & Abadi 2011, Zipkin & Saunders 2017).
+Integrated population models (IPMs; Besbeas et al.
+2002, Brooks et al.
+2004) are an increasingly popular tool in ecology that can help overcome these limitations by combining disparate datasets in a single, unified analysis (Schaub & Abadi 2011).
+This approach can provide several advantages, such as improved precision (Schaub et al.
+2007, Tavecchia 2009, Abadi et al.
+2010) and the estimation of parameters that would otherwise be unidentifiable (Besbeas 2005, Schaub et al.
+2007, Veran & Lebreton 2008), which is often the case for key demographic rates such as recruitment or immigration.
+These advances have practical benefits for resource managers that must make decisions based on limited available information.
+As IPMs become more widely used, however, it is critical that they are robustly evaluated.
+In particular, the limitations of the IPM framework are still underexplored, particularly in situations where available data are sparse, of poor quality, or when data to inform specific parameters are entirely absent.
+
+Through a simulation analysis, we will identify monitoring programs and demographic scenarios where integrated population models may not lead to improved precision or return unbiased estimates of parameters that would otherwise be unidentifiable when certain datasets are excluded.
+We apply this simulation study to to a passerine bird life history.
+We assume all model assumptions are met and compare parameter estimation for a traditional IPM with data informing abundance (count data), survival (mark-recapture data), and reproductive output (nest monitoring data) to situations when either of the latter two datasets are omitted.
+In doing so, we will refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool or might not be necessary or warrant the collection of additional data.
+Improving our understanding of the strengths and weaknesses of the IPM framework can provide helpful insights for the design and implementation of field survey programs so as to most efficiently make use of limited resources to inform management efforts aimed at monitoring or recovering species that are deemed ecosystem indicators or depleted and in need of conservation intervention.
 
 # Scripts, current version
+
 ## 0 - preparing scenarios
+
 ### compute_time_calc.R
@@
 ## 1 - simulating data
+
 ### IPM_sim_2.0function.R
+
 Script contains two functions for simulating data: 'simPopTrajectory' and 'simData'.
+
 The 'simPopTrajectory' function simulates the true population trajectory by individual and age class, starting from the stable age distribution.
@@
 ## 3 - run models
+
 ### run_scenarios_helperFns.R
+
 Script with functions and code to run NIMBLE models.
@@
 ### run_scenarios.R
-Script that pulls from the scripts above to run models in parallel and output results.
+
+Main workhorse script that pulls from the scripts above to run models in parallel and output results.
 
 ## 4 - process results
+
 ### 01_load_and_process_data.R
+
 Script that checks the Gelman-Rubin convergence statistic for each output MCMC sample, then thins the chains to reduce file size and places the output into categories based on the population growth rate, $\lambda$.
@@
 ### 02_compute_geom_means.R
+
 Computes the geometric means for $\lambda$ in each MCMC output.
@@
 # Data
+
 Contains true parameter scenarios used to simulate population trajectories and observation data, all viable parameter combinations in files '-.lam.combos.RDS' and the selected parameter combinations used to simulate data are in files '-.lam.params.RDS' *(where "-" is low, med, or high)*.

### 2021-08-17 figures/ARCHIVE/EuringPosterFigures.Rmd
repo reorganization

--- 2021-05-04 scripts/EuringPosterFigures.Rmd
@@
 library(RColorBrewer)
+library(colorspace)
+library(ggh4x)
 ```
@@
 # plot; use me - overview
-ggplot(rel.bias, aes(x = factor(simscenarios), y = variable, fill = mean.rel.bias)) +
+## change factor levels
+rel.bias$variable <- factor(rel.bias$variable, levels = c("Adult survival", "Juv survival", "Fecundity"),
+                            ordered = TRUE, labels = c("Adult Survival", "Juvenile Survival", "Fecundity"))
+
+rel.bias$lambda.scenario <- factor(rel.bias$lambda.scenario, levels = c("Decreasing", "Stable", "Increasing"),
+                                ordered = TRUE, labels = c(expression(paste("Decreasing ", lambda)), 
+                                                           expression(paste("Stable ", lambda)), 
+                                                           expression(paste("Increasing ", lambda))))
+# rel.bias$simscenarios <- factor(rel.bias$simscenarios, levels = c("L", "M", "H"),
+#                                 ordered = TRUE, labels = c(expression(paste("low ", italic(p))), 
+#                                                            expression(paste("med ", italic(p))), 
+#                                                            expression(paste("high ", italic(p)))))
+
+## color palette
+pal <- diverging_hcl(n = 7, palette = "Cyan-Magenta")
+
+## plotting
+ov <- ggplot(rel.bias, aes(x = factor(simscenarios), y = variable, fill = mean.rel.bias)) +
   geom_tile(color = 'grey50') +
-  xlab('Detection level') + ylab('') +
-  facet_grid(~lambda.scenario, drop = T, scales = 'free_x') +
-  scale_fill_gradient2(name = "Relative bias",
+  xlab('') + ylab('') +
+  facet_grid(~lambda.scenario, drop = T, scales = 'free_x', labeller = label_parsed) +
+  scale_fill_gradient2(name = "Relative\nbias",
                       # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
-                     mid = "#f7f7f7", low = "royalblue4", high = "orange2", midpoint = 0,
+                     mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0,
                       breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
-  theme_minimal() +
+ scale_x_discrete(labels = c(expression(paste("low ", italic(p))), 
+                             expression(paste("med ", italic(p))), 
+                             expression(paste("high ", italic(p))))) +
+  theme_light() +
   theme(legend.position = 'top',
-             strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"))
+        #axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        strip.text.x = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA, color = "black"),
+        axis.text.x = element_text(angle = 0, vjust = 1.5),
+         panel.border = element_rect(color = "black", fill = NA),  
+        panel.spacing.x = unit(0.75, "line")) +
+        # ggh4x.facet.nestline = element_line(size = 0.5)) +
+        # panel.spacing = unit(1, "lines")) +
+  ggtitle("") 
+ov
 
+## export out
+ggsave(filename = here("figures", "overview_params_bias.pdf"), plot = ov)
 ```
@@
 
+```
+
+```{r phiad}
+pal1 <- diverging_hcl(n = 7, palette = "Blue-Yellow 3")
+phiad <- rel.bias.vals %>% filter(variable == 'Adult survival') 
+summary(phiad$value) # look at range for scale
+
+## fix faceted labels
+rel.bias.vals$lambda.scenario <- factor(rel.bias.vals$lambda.scenario, levels = c("Decreasing", "Stable", "Increasing"),
+                                ordered = TRUE, labels = c(expression(paste("Decreasing ", lambda)), 
+                                                           expression(paste("Stable ", lambda)), 
+                                                           expression(paste("Increasing ", lambda))))
+rel.bias.vals$simscenarios <- factor(rel.bias.vals$simscenarios, levels = c("L", "M", "H"),
+                                ordered = TRUE, labels = c(expression(paste("low ", italic(p))), 
+                                                           expression(paste("med ", italic(p))), 
+                                                           expression(paste("high ", italic(p)))))
+
+p1 <- ggplot(rel.bias.vals %>% filter(variable == 'Adult survival'),
+             aes(x = factor(phiad.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('\nTrue Adult Survival') + ylab("") +
+  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
+               nest_line = TRUE) +
+    scale_fill_gradient2(name = "Relative\nbias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      # mid = "white", low = "#9FA2FF", high = "#BAAE00", midpoint = 0,
+                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
+                      breaks = c(-0.15, -0.05, 0, 0.05), n.breaks = 4, labels = c("-0.15", "-0.05", "0", "0.05")) +
+  theme_light() +
+  theme(legend.position = 'right',
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        strip.text.x = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA, color = "black"),
+        axis.text.x = element_text(angle = 90, vjust = 1.5),
+        ggh4x.facet.nestline = element_line(size = 0.5),
+        panel.border = element_rect(color = "black", fill = NA),  
+        panel.spacing.x = unit(0.75, "line")) +
+        # panel.spacing = unit(1, "lines")) +
+  ggtitle("Adult Survival")
+p1
+```
+
+```{r phijuv}
+pal2 <- diverging_hcl(n = 7, palette = "Green-Orange")
+phi1 <- rel.bias.vals %>% filter(variable == 'Juv survival')
+summary(phi1$value)
+
+p2 <- ggplot(rel.bias.vals %>% filter(variable == 'Juv survival'), 
+             aes(x = factor(phi1.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('\nTrue Juvenile Survival') + ylab("") +
+  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
+               nest_line = TRUE) +
+  scale_fill_gradient2(name = "Relative\nbias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      # mid = "white", low = "#11C638", high = "#EF9708", midpoint = 0,
+                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
+                      breaks = c(-0.2, 0, 0.2, 0.5, 0.8), n.breaks = 5, 
+                      labels = c("-0.2", "0", "0.2", "0.5", "0.8")) +
+  theme_light() +
+ theme(legend.position = 'right',
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        strip.text.x = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA, color = "black"),
+        axis.text.x = element_text(angle = 90, vjust = 1.5),
+        ggh4x.facet.nestline = element_line(size = 0.5),
+        panel.border = element_rect(color = "black", fill = NA),  
+        panel.spacing.x = unit(0.75, "line")) +
+        # panel.spacing = unit(1, "lines")) +
+  ggtitle("Juvenile Survival")
+p2
+```
+
+```{r fec}
+pal3 <- diverging_hcl(n = 7, palette = "Cyan-Magenta")
+fec <- rel.bias.vals %>% filter(variable == 'Fecundity')
+summary(fec$value)
+
+p3 <- ggplot(rel.bias.vals %>% filter(variable == 'Fecundity'), 
+             aes(x = factor(fec.true), y = variable, fill = value)) +
+  geom_tile(color = 'grey50') +
+  xlab('\nTrue Fecundity') + ylab("") +
+  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
+  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
+               nest_line = TRUE) +
+  scale_fill_gradient2(name = "Relative\nbias",
+                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
+                      # mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0,
+                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
+                      breaks = c(-0.4, -0.2, 0, 0.2), n.breaks = 4, 
+                      labels = c("-0.4", "-0.2", "0", "0.2")) +
+  theme_light() +
+  theme(legend.position = 'right',
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
+        strip.text.x = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA, color = "black"),
+        axis.text.x = element_text(angle = 90, vjust = 1.5),
+        ggh4x.facet.nestline = element_line(size = 0.5),
+        panel.border = element_rect(color = "black", fill = NA),  
+        panel.spacing.x = unit(0.75, "line")) +
+        # panel.spacing = unit(1, "lines")) +
+  ggtitle("Fecundity")
+p3
+```
+
+```{r combo}
+## export out
+pdf(here("figures", "params_bias.pdf"), width = 12, height = 8)
+plot_grid(p3, p2, p1, nrow = 3)
+dev.off()
+
+## try with overview plot
+all <- ov + (p3/p2/p1)
+ggsave(filename = here("figures", "all_bias.pdf"), plot = all, width = 22, height = 12)
+#use # for spaces with missing plots
+# layout_string <- 'A#BB
+#                   ##CC
+#                   ##DD'
+# junk <- ov + (p3/p2/p1) + plot_layout(design=layout_string,
+#                               widths=c(3,1,5,6),
+#                               heights=c(5,3))
 ```

### 2021-08-17 figures/PaperFigures.Rmd
repo reorganization

--- 2021-08-17 scripts/PaperFigures.Rmd
@@
 #   dplyr::select('phi1.obs', 'phi1.true', 'phiad.obs', 'phiad.true', 'fec.obs', 'fec.true',
-#                 'p.surv', 'mean.p',
 #                             'sims', 'scenario', 'simscenarios', 'lambda.scenario') %>%
@@
 # 
-# write.csv(all.meds, here::here('figures', 'Processed csvs', 'all.meds.csv'), row.names = F)
+# write.csv(all.meds, here::here('results', 'Processed csvs', 'all.meds.csv'), row.names = F)
 
@@
 
-##bias
+# #just bias
 rel.bias <- all.meds %>%
-  transform(p.surv.true = ifelse(det.abund == 'L', 0.3))
   #calculate relative bias, mean across scenarios

### 2021-08-17 figures/PaperFigures.Rmd

--- 2021-08-17 scripts/PaperFigures.Rmd
@@
 
-all.meds <- read.csv(file = here::here('results', 'Processed csvs', 'all.meds.csv'), 
+all.meds <- read.csv(file = here::here('figures', 'Processed csvs', 'all.meds.csv'), 
                      header = T, stringsAsFactors = F)
@@
 rel.bias <- all.meds %>%
-  transform(p.surv.true = ifelse(det.abund == 'L', 0.3))
+  transform(p.surv.true = ifelse(det.abund == 'L', 0.3, 
+            ifelse(det.abund == 'M', 0.5, 
+                   ifelse(det.abund == 'H', 0.8, NA)))) %>%
+  transform(mean.p.true = ifelse(det.MR == 'L', 0.3, 
+            ifelse(det.MR == 'M', 0.5, 
+                   ifelse(det.MR == 'H', 0.8, NA)))) %>%
   #calculate relative bias, mean across scenarios
@@
             phiad.bias = (phiad.obs-phiad.true)/phiad.true,
-            fec.bias = (fec.obs-fec.true)/fec.true) %>%
+            fec.bias = (fec.obs-fec.true)/fec.true,
+            p.surv.bias = (p.surv-p.surv.true)/p.surv.true,
+            mean.p.bias = (mean.p-mean.p.true)/mean.p.true) %>%
   dplyr::select('lambda.scenario', 'scenario',
-                'phi1.bias', 'phiad.bias', 'fec.bias', 'det.MR', 'det.abund', 'det.prod') %>%
+                'phi1.bias', 'phiad.bias', 'fec.bias', 'det.MR', 'det.abund', 'det.prod', 'mean.p.bias', 'p.surv.bias') %>%
   reshape2::melt(id.vars = c('lambda.scenario', 'scenario', 'det.MR', 'det.abund', 'det.prod')) %>%
@@
   dplyr::summarize(bias = mean(value), .groups = 'keep') %>%
-  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias'),
-                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>%
+  transform(variable = factor(variable, levels = c('phiad.bias', 'phi1.bias', 'fec.bias', 'p.surv.bias', 'mean.p.bias'),
+                              labels = c('Adult survival', 'Juv survival', 'Fecundity', 'Count detection', 'Resight detection'))) %>%
   transform(lambda.scenario = factor(lambda.scenario,
@@
 ###average over two layers of detection (det.MR and det.prod)
+obs.pars <- c('Resight detection', 'Count detection')
 rel.bias.few <- rel.bias %>%
@@
 
-#use me fig 1
-ggplot(rel.bias.few, aes(x = factor(det.abund), y = variable, fill = bias)) +
-  geom_tile(color = 'grey50') +
-  xlab('Count detection') + ylab('') +
-  facet_grid(dataset ~ lambda.scenario, drop = T, scales = 'free_x', labeller = label_parsed) +
-  scale_fill_gradient2(name = "Relative bias",
-                     # mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0) +
-                       mid = "white", low = rainbow2[2], high = rainbow2[6], midpoint = 0) +
-                      # breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
- # scale_x_discrete(labels = c(expression(paste("low ", italic(p))), 
- #                             expression(paste("med ", italic(p))), 
- #                             expression(paste("high ", italic(p))))) +
-  theme_light() +
-  theme(legend.position = 'top',
-        # axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        strip.text = element_text(face = "bold", color = "black"),
-        strip.background = element_rect(fill = NA, color = "black"),
-        axis.text.x = element_text(angle = 0, vjust = 1.5),
-        # ggh4x.facet.nestline = element_line(size = 0.5), #what does this line do?
-        panel.border = element_rect(color = "black", fill = NA),  
-        panel.spacing.x = unit(0.75, "line"))
+#heat tiles
+# ggplot(rel.bias.few %>% filter(variable %nin% obs.pars), 
+#        aes(x = factor(det.abund), y = variable, fill = bias)) +
+#   geom_tile(color = 'grey50') +
+#   xlab('Count detection') + ylab('') +
+#   facet_grid(dataset ~ lambda.scenario, drop = T, scales = 'free_x', labeller = label_parsed) +
+#   scale_fill_gradient2(name = "Relative bias",
+#                      # mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0) +
+#                        mid = "white", low = rainbow2[2], high = rainbow2[6], midpoint = 0) +
+#                       # breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+#  # scale_x_discrete(labels = c(expression(paste("low ", italic(p))), 
+#  #                             expression(paste("med ", italic(p))), 
+#  #                             expression(paste("high ", italic(p))))) +
+#   theme_light() +
+#   theme(legend.position = 'top',
+#         # axis.text.y = element_blank(),
+#         axis.ticks.y = element_blank(),
+#         axis.ticks.x = element_blank(),
+#         panel.grid.major = element_blank(),
+#         panel.grid.minor = element_blank(),
+#         strip.text = element_text(face = "bold", color = "black"),
+#         strip.background = element_rect(fill = NA, color = "black"),
+#         axis.text.x = element_text(angle = 0, vjust = 1.5),
+#         # ggh4x.facet.nestline = element_line(size = 0.5), #what does this line do?
+#         panel.border = element_rect(color = "black", fill = NA),  
+#         panel.spacing.x = unit(0.75, "line"))
 
@@
 
-ggplot(rel.bias.few, aes(x = det.abund, y = bias, col = factor(variable), group = factor(variable))) +
+#use me fig 1a
+pdf(here("figures", "bias.lambda.ecol.pdf"), width = 8, height = 12)
+ggplot(rel.bias.few  %>% filter(variable %nin% obs.pars), 
+       aes(x = det.abund, y = bias, col = factor(variable), group = factor(variable))) +
   geom_point() + geom_line() +
@@
   scale_color_manual(values = rainbow2[-c(1,4)], name = '')
+dev.off()
+
+#use me fig 1b
+pdf(here("figures", "bias.lambda.obs.pdf"), width = 8, height = 12)
+ggplot(rel.bias.few  %>% filter(variable %in% obs.pars), 
+       aes(x = det.abund, y = bias, col = factor(variable), group = factor(variable))) +
+  geom_point() + geom_line() +
+  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
+  xlab('Count detection') + ylab('Relative bias') +
+  facet_grid(dataset~lambda.scenario, scales = 'free_x') +
+  theme_bw() +
+  theme(legend.position = 'top',
+             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
+  scale_color_manual(values = rainbow2[-c(1,4)], name = '')
+dev.off()
 
@@
 ## dot plot
-# ggplot(plot.vals, aes(x = fec_cat, y = value, col = factor(dataset), group = factor(dataset))) +
-#   geom_point() + geom_line() +
-#   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
-#   xlab('True fecundity') + ylab('Relative bias') +
-#   facet_grid(variable~phi1_cat, scales = 'free') +
-#   theme_bw() +
-#   theme(legend.position = 'top',
-#              plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
-#   scale_color_manual(values = rainbow2[-c(1,4)], name = '')
-
+#use me figure 2
+pdf(here("figures", "bias.dem.ecol.pdf"), width = 8, height = 12)
 ggplot(plot.vals, aes(x = fec_cat, y = value, col = factor(variable), group = factor(variable))) +
@@
   scale_color_manual(values = rainbow2[-c(1,4)], name = '')
+dev.off()
 
@@
 rmse.vals <- all.meds %>%
+  transform(p.surv.true = ifelse(det.abund == 'L', 0.3, 
+            ifelse(det.abund == 'M', 0.5, 
+                   ifelse(det.abund == 'H', 0.8, NA)))) %>%
+  transform(mean.p.true = ifelse(det.MR == 'L', 0.3, 
+            ifelse(det.MR == 'M', 0.5, 
+                   ifelse(det.MR == 'H', 0.8, NA)))) %>%
   transform(fec.rmse = (fec.obs-fec.true)^2,
             phiad.rmse = (phiad.obs-phiad.true)^2,
-            phi1.rmse = (phi1.obs-phi1.true)^2) %>%
-  dplyr::select(lambda.scenario, scenario, fec.rmse, phiad.rmse, phi1.rmse, 
+            phi1.rmse = (phi1.obs-phi1.true)^2,
+            mean.p.rmse = (mean.p-mean.p.true)^2,
+            p.surv.rmse = (p.surv-p.surv.true)^2) %>%
+  dplyr::select(lambda.scenario, scenario, fec.rmse, phiad.rmse, phi1.rmse, p.surv.rmse, mean.p.rmse,
                 det.MR, det.abund, det.prod) %>%
@@
   dplyr::summarize(mean.rmse = mean(value), .groups = 'keep') %>%
-  transform(variable = factor(variable, levels = c('phiad.rmse', 'phi1.rmse', 'fec.rmse'),
-                              labels = c('Adult survival', 'Juv survival', 'Fecundity'))) %>%
+  transform(variable = factor(variable, levels = c('phiad.rmse', 'phi1.rmse', 'fec.rmse', 
+                                                   'p.surv.rmse', 'mean.p.rmse'),
+                              labels = c('Adult survival', 'Juv survival', 'Fecundity',
+                                         'Count detection', 'Resight detection'))) %>%
   transform(lambda.scenario = factor(lambda.scenario,
@@
 #use me fig 3
-ggplot(rmse.few, aes(x = factor(det.abund), y = variable, fill = rmse)) +
+pdf(here("figures", "rmse.lambda.ecol.pdf"), width = 8, height = 12)
+ggplot(rmse.few %>% filter(variable %nin% obs.pars), aes(x = factor(det.abund), y = variable, fill = rmse)) +
   geom_tile(color = 'grey50') +
@@
         panel.spacing.x = unit(0.75, "line"))
+dev.off()
+
+pdf(here("figures", "rmse.lambda.obs.pdf"), width = 8, height = 12)
+ggplot(rmse.few %>% filter(variable %in% obs.pars), aes(x = factor(det.abund), y = variable, fill = rmse)) +
+  geom_tile(color = 'grey50') +
+  xlab('Count detection') + ylab('') +
+  facet_grid(dataset ~ lambda.scenario, drop = T, scales = 'free_x', labeller = label_parsed) +
+  scale_fill_gradient2(name = "RMSE",
+                       mid = "white", high = rainbow2[2], midpoint = 0) +
+                      # breaks = c(-0.2, 0, 0.2), n.breaks = 3, labels = c("-0.2", "0", "0.2")) +
+ # scale_x_discrete(labels = c(expression(paste("low ", italic(p))), 
+ #                             expression(paste("med ", italic(p))), 
+ #                             expression(paste("high ", italic(p))))) +
+  theme_light() +
+  theme(legend.position = 'top',
+        # axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        strip.text = element_text(face = "bold", color = "black"),
+        strip.background = element_rect(fill = NA, color = "black"),
+        axis.text.x = element_text(angle = 0, vjust = 1.5),
+        # ggh4x.facet.nestline = element_line(size = 0.5), #what does this line do?
+        panel.border = element_rect(color = "black", fill = NA),  
+        panel.spacing.x = unit(0.75, "line"))
+dev.off()
 
@@
 
-ggplot(rmse.few, aes(x = det.abund, y = rmse, col = factor(variable), group = factor(variable))) +
+ggplot(rmse.few %>% filter(variable %nin% obs.pars), aes(x = det.abund, y = rmse, col = factor(variable), group = factor(variable))) +
+  geom_point() + geom_line() +
+  geom_hline(aes(yintercept = 0), linetype = 'dotted') +
+  xlab('Count detection') + ylab('RMSE') +
+  facet_grid(dataset~lambda.scenario, scales = 'free') +
+  theme_bw() +
+  theme(legend.position = 'top',
+             plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
+  scale_color_manual(values = rainbow2[-c(1,4)], name = '')
+
+ggplot(rmse.few %>% filter(variable %in% obs.pars), aes(x = det.abund, y = rmse, col = factor(variable), group = factor(variable))) +
   geom_point() + geom_line() +
@@
 #use me fig 3
+pdf(here("figures", "rmse.dem.ecol.pdf"), width = 8, height = 12)
 ggplot(plot.vals.rmse, aes(x = factor(fec_cat), y = factor(variable), fill = value)) +
@@
         panel.spacing.x = unit(0.75, "line"))
+dev.off()
 

### 2021-08-27 figures/PaperFigures.Rmd
Figure tidying post lab meeting code review.

@@
 library(ggh4x)
+library(Hmisc)
 
@@
   theme(legend.position = 'top',
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 9),
         # axis.text.y = element_blank(),
@@
         # axis.text.y = element_blank(),
+        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
+        legend.title = element_text(size = 10, vjust = 0.75),
         axis.ticks.y = element_blank(),
@@
     scale_fill_gradient2(name = "RMSE",
-                      mid = "white", high = rainbow2[2], midpoint = 0,
-                      # breaks = c(-0.5, -0.25, 0, 0.25, 0.5), n.breaks = 5, 
-                      labels = c('0.5', "-0.25", "0", "0.25", '0.5')) +
+                      mid = "white", high = rainbow2[2]) +
   theme_light() +
@@
         # axis.text.y = element_blank(),
+        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
+        legend.title = element_text(size = 11, vjust = 0.75),
         axis.ticks.y = element_blank(),
@@
   theme(legend.position = 'right',
+        legend.text = element_text(size = 8, angle = 90),
         axis.text.y = element_blank(),

### 2021-08-27 scripts/Figures.Rmd
Figure tidying post lab meeting code review.

--- 2021-08-17 figures/ARCHIVE/EuringPosterFigures.Rmd
@@
 
-load(file = here::here('results', 'EuringPoster', 'processedIPMOutput.RData'))
-
-row.low <- do.call(rbind, lapply( ls(patt='lowout'), get) )
-row.med <- do.call(rbind, lapply( ls(patt='medout'), get) )
-row.high <- do.call(rbind, lapply( ls(patt='highout'), get) )
-rm(list=grep('highout|medout|lowout',ls(),value=TRUE,invert=FALSE))
+row.low <- read.csv(file = here::here('results', 'lowout.csv'))
+row.med <- read.csv(file = here::here('results', 'medout.csv'))
+row.high <- read.csv(file = here::here('results', 'highout.csv'))
 

### 2021-09-03 figures/PaperFigures.Rmd
tidied figure facet labels

@@
 
-all.meds.sc <- read.csv(file = here::here('results', 'Processed csvs', 'all.meds.sc.csv'), 
+all.meds.sc <- read.csv(file = here::here('figures', 'Processed csvs', 'all.meds.sc.csv'), 
                      header = T, stringsAsFactors = F)
@@
             phiad_cat = factor(phiad_cat, levels = c('L', 'M', 'H')),
-            phi1_cat = factor(phi1_cat, levels = c('L', 'M', 'H'))) %>%
+            phi1_cat = factor(phi1_cat, levels = c('L', 'M', 'H'),
+                              labels = c('True juvenile survival: L', 
+                                         'True juvenile survival: M', 
+                                         'True juvenile survival: H'))) %>%
   transform(lambda.scenario = factor(lambda.scenario, 
@@
 
-#use me fig 2
+#heat map
 ggplot(plot.vals, aes(x = factor(fec_cat), y = variable, fill = value)) +
@@
             phiad_cat = factor(phiad_cat, levels = c('L', 'M', 'H')),
-            phi1_cat = factor(phi1_cat, levels = c('L', 'M', 'H'))) %>%
+            phi1_cat = factor(phi1_cat, levels = c('L', 'M', 'H'),
+                              labels = c('True juvenile survival: L', 
+                                                   'True juvenile survival: M', 
+                                                   'True juvenile survival: H'))) %>%
   transform(dataset = factor(dataset, levels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'),

### 2021-10-12 figures/lambdaTrendsPlot.R
abundance trend figure update

@@ -92,7 +92,7 @@ toplot <- bind_rows(toplot1, toplot2, toplot3) %>%
   mutate(Quantile = paste("X", Quantile, sep = "")) %>% 
   reshape2::dcast(dataset + Year +  det.MR + det.prod + det.abund   + lambda ~ Quantile, value.var = "value") %>% 
   mutate(Year = Year + 1) %>% 
-  filter(Year %in% c(2:6, 15)) %>% 
+  filter(Year %in% c(15)) %>% 
   mutate(Year = factor(Year)) %>% 
   mutate(det.abund = factor(det.abund, levels = c("L", "M", "H"))) %>% 
   mutate(det.prod = factor(det.prod, levels = c("L", "M", "H"))) %>% 
@@ -107,28 +107,21 @@ toplot <- bind_rows(toplot1, toplot2, toplot3) %>%
     lambda == "Increasing" ~ 1.05))
 
 
-pdf(here("figures", "lambdaTrends.pdf"), width = 12, height = 8)
+pdf(here("figures", "lambdaTrends.pdf"), width = 8, height = 8)
 rainbow2 <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")
 ggplot(toplot) +
   geom_point(aes(x = Year, y = X50, col = det.abund, group = det.abund), position = position_dodge(width = 0.5)) +
-  #geom_linerange(aes(ymin = X2.5, ymax = X97.5, x = Year), position = position_dodge(width = 0.5)) +
-  #geom_hline(aes(yintercept = 1.0), linetype = 'dotted') +
-  xlab('Years') + ylab('Lambda') + 
+  geom_linerange(aes(x = Year, ymin = X2.5, ymax = X97.5, col = det.abund, group = det.abund), position = position_dodge(width = 0.5)) +
+  geom_hline(aes(yintercept = intercept), linetype = 'dotted') +
+  geom_hline(aes(yintercept = 1.0), linetype = 'solid') +
+  xlab('Years') + 
+  ylab('Lambda') + 
   #ylim(0.87, 1.14) +
   facet_grid(dataset ~ lambda, scales = 'free') +
   theme_bw() +
   theme(legend.position = 'top',
         plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
-  scale_color_manual(values = rainbow2[-c(1,4)], name = 'Abundance detection level') + 
-  geom_rect(aes(ymin=-Inf,
[…]

### 2021-10-12 figures/ARCHIVE/figure_exploring.R
big repo reorg for publication version

@@ -1,190 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(cowplot)
-library(ggplot2)
-library(coda)
-library(captioner)
-library(knitr)
-library(reshape2)
-library(here)
-
-rm(list=ls())
-
-#true vals
-truth <- readRDS(here::here('data', 'ARCHIVE', 'scenarios.Rdata')) 
-colnames(truth) <- gsub(' ', '', colnames(truth))
-true_vals <- truth %>%
-  dplyr::select(ScenarioNumber, MRdetection, Abunddetection, AdultSurv, JuvSurv, Dailynestsurvival, Fec) %>%
-  dplyr::rename(scenario = ScenarioNumber) %>%
-  reshape2::melt(id.vars = c('scenario'), value.name = 'Truth') %>%
-  transform(variable = factor(variable,
-                              levels = c('JuvSurv', 'AdultSurv', 'Fec', 'Dailynestsurvival', 'MRdetection', 'Abunddetection'),
-                              labels = c('Juvenile survival', 'Adult survival', 'Fecundity', 'Daily nest survival', 
-                                         'MR detection', 'Survey detection'))) %>%
-  transform(Data = ifelse(scenario < 6, 'All', 
-                          ifelse(scenario < 11 & scenario > 5, 'Counts + Productivity', 'Counts + MR'))) %>%
-  transform(Truth = ifelse(Data == 'Counts + Productivity' & variable == 'MR detection', NA, Truth))
-
-# ~~~~~~~~~~~~~~~~~~~~~~~~~~ #
-
-### assessing convergence
-
-# test <-readRDS(file = here::here('data', paste0('out_3-1.Rdata')))[3500:4000,]
-
-
-sc <- c(1:5, 11:15)
-# sc <- c(1:5)
-nsim <- 25
-pars <- c('fec', 'lambdaf', 'mean.p', 'mean.phi[1]', 'mean.phi[2]', 'p.surv', 'phi.nest')
-
-out_meds1 <- data.frame() #storage
-for (s in sc) {
-  for (i in 1:nsim) { 
-    chains <- as.mcmc.list(list(
-      as.mcmc(readRDS(file = here::here('data', 'ARCHIVE', paste0('out', s, '_', i, '-', 1, '.Rdata')))[3500:4000,]), 
[…]

### 2021-10-12 scripts/ARCHIVE/3 - test models/testAllMods.R

@@ -1,198 +1,0 @@
-# aeb
-# march 2, 2020
-# example post processing
-library(tidyverse)
-library(here)
-library(nimble)
-library(foreach)
-library(doParallel)
-library(beepr)
-
-# load data
-scenarios <- readRDS(here("data", "scenarios.RDS"))
-low.lam.combos <- readRDS(here("data", "low.lam.combos.RDS"))
-med.lam.combos <- readRDS(here("data", "med.lam.combos.RDS"))
-high.lam.combos <- readRDS(here("data", "high.lam.combos.RDS"))
-
-# functions
-source(here("scripts", "current version",
-            "1 - simulating data", "IPM_sim_2.0function.R"))
-source(here("scripts", "current version",
-            "2 - models", "IPM_marray.R"))
-source(here("scripts", "current version",
-            "4 - run models", "run_scenarios_helperFns.R"))
-
-low.comb <- low.lam.combos[sample(1:5000, 1), 1:3]
-med.comb <- med.lam.combos[sample(1:5000, 1), 1:3]
-high.comb <- high.lam.combos[sample(1:5000, 1), 1:3]
-
-lowpopTraj <- simPopTrajectory(n.years=15,
-                               n.data.types=c(0.25,0.25,0.25),
-                               age.init=c(150,150),
-                               phi.1=as.numeric(low.comb[2]),
-                               phi.ad=as.numeric(low.comb[3]),
-                               f=as.numeric(low.comb[1]))
-
-medpopTraj <- simPopTrajectory(n.years=15,
-                               n.data.types=c(0.25,0.25,0.25),
-                               age.init=c(150,150),
-                               phi.1=as.numeric(med.comb[2]),
-                               phi.ad=as.numeric(med.comb[3]),
-                               f=as.numeric(med.comb[1]))
-
-highpopTraj <- simPopTrajectory(n.years=15,
-                                n.data.types=c(0.25,0.25,0.25),
-                                age.init=c(150,150),
-                                phi.1=as.numeric(high.comb[2]),
-                                phi.ad=as.numeric(high.comb[3]),
-                                f=as.numeric(high.comb[1]))
-
-# simulate data
-detect.l <- 0.3
-detect.m <- 0.5
-detect.h <- 0.8
-
-detect <- c(detect.l, detect.m, detect.h)
-
[…]

### 2021-10-12 scripts/ARCHIVE/3 - test models/testFigure.R

@@ -1,275 +1,0 @@
-#### Processing #######
-
-library(ggplot2)
-library(gtable)
-library(RColorBrewer)
-library(wesanderson)
-
-pal <- rev(wes_palette("Zissou1", 3, type = "continuous"))
-
-# 9 panel plot
-  # rows - low, med, high lambda
-  # columns - low, med, high detection
-
-# row 1
-# declining population
-
-row1 <- bind_rows(as.data.frame(as.matrix(`highout-1-1`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-1-2`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-1-3`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-1-4`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-2-1`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-2-2`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-2-3`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-2-4`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-3-1`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-3-2`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-3-3`)[1:1000, ]),
-                  as.data.frame(as.matrix(`highout-3-4`)[1:1000, ])
-                  ) %>%
-  select(contains("lambda"))
-
-for (i in 1:14) {
-  row1 <- row1 %>%
-    mutate("geomean.{i}" :=  NA_real_)
-}
-
-for(i in 1:dim(row1)[1]) {
-  for(j in 1:((ncol(row1) - 2 )/2)) {
-    print(paste("row", i, "col", j))
-    row1[i, (ncol(row1)/2 + j)] <- exp(mean(unlist(log(row1[i, 1:j]))))
-  }
-}
-
-row1 <- cbind(row1,
-              model = rep(rep(1:4, each = 1000), times = 3),
-              detection = rep(1:3, each = 1000*4)) %>%
-  as.data.frame()
-
-toplot1 <- row1 %>%
-  select(contains("geomean"), model, detection) %>%
-  #group_by(model, detection)
-  pivot_longer(cols = starts_with("geomean"), names_to = "Year") %>%
-  filter(!is.na(value)) %>%
[…]

### 2021-10-12 scripts/ARCHIVE/ExampleUpdateNimbleConverge.R

@@ -1,162 +1,0 @@
-#example for how to check for convergence and update model runs
-#using an occupancy model and a bit of clunky coding!
-
-library(nimble)
-################simulate data
-#simulate some occupancy data for a super simple occupany model
-#50 sites, 3 surveys
-true_z<-rbinom(50,1,0.8) #true state of each site
-obs_y<-matrix(nrow=50,ncol=3)
-for(i in 1:50){
-  #observed state of each site at each survey
-  #0.6 is the probability of detecting whatever species during a survey
-  obs_y[i,]<-rbinom(3,1, true_z[i]*0.6)
-}
-
-#make the observation data
-dat_sim<-list(y=obs_y)
-
-############nimble stuff
-#nimble model
-mod<-nimbleCode({
-  p~dunif(0,1)
-  phi~dunif(0,1)
-  for(i in 1:N){
-    z[i]~dbern(phi)
-  }
-  for(i in 1:N){
-    for(j in 1:J){
-      zp[i,j]<-z[i]*p
-      y[i,j]~dbern(zp[i,j])
-    }
-  }
-})
-
-#function to grab initial values
-initsfn<-function(){
-  p<-runif(1,0,1)
-  phi<-runif(1,0,1)
-  return(list(p=p,phi=phi))
-}
-
-##################################
-#Here is a clunky version
-#these commanands are for all chains
-constants<-list(N=50, J=3)
-Rmodel<-nimbleModel(code=mod, constants=constants)
-conf<-configureMCMC(Rmodel)
-conf$addMonitors(c("phi","p"))
-
-#then need to set it up to run different chains with different initial values
-#so one for each chain
-Rmcmc1<-buildMCMC(conf)
-Rmcmc2<-buildMCMC(conf)
-
-Cmodel1<-compileNimble(Rmodel)
-Cmodel2<-compileNimble(Rmodel)
-
-Cmodel1$setData(dat_sim)
-Cmodel2$setData(dat_sim)
-
-Cmodel1$setInits(inits=initsfn())
-Cmodel2$setInits(inits=initsfn())
-
-Cmcmc1<-compileNimble(Rmcmc1, project=Cmodel1)
-#run the chain
-Cmcmc1$run(thin=10, reset=T, niter=1000,nburnin=500)
-#save output for this chain
-x11<-(as.data.frame(as.matrix(Cmcmc1$mvSamples)))
-
-#same as above
[…]

### 2021-10-12 scripts/ARCHIVE/IPMNimble.R

@@ -1,225 +1,0 @@
-# TODO
-# abby put nest stuff in here
-
-#IPM model for sim data
-library(here)
-#source(here("scripts", "dataSims","BPA_IPMsimData.R")) #full on BPA WinBUGS code
-source(here("scripts", "dataSims","IPMsimData_nonest.R")) #changed the MR data
-
-# 11.3. Example of a simple IPM (counts, capture-recapture, reproduction)
-# 11.3.1. Load data
-# Population counts (from years 1 to n.years)
-y <- df$SUR
-n.sam <- df$n.sam
-
-# Capture-recapture data (in m-array format, from years 1 to n.years)
-m <- df$ch
-first <- df$first
-age_ch <- df$age_ch
-
-#Productivity/nest success not in the data sim right now
-#these will need to be changed once we have the nest success data
-# Productivity data (from years 1 to n.years-1)
-#J<-nestlings#df$nestlings
-#R<-R#df$R
- 
-H <- df$H
-Fledged <- df$Fledged
-first.nest <- df$first.nest
-last.nest <- df$last.nest
-max.nest.age <- df$max.nest.age
-n.nests <- df$n.nests
-n.succ.nests <- df$n.succ.nests
-
-library("nimble")
-IPMmod<-nimbleCode({
-  #-------------------------------------------------
-  #  Integrated population model
-  #  - Age structured model with 2 age classes: 
-  #		1-year old and adult (at least 2 years old)
-  #  - Age at first breeding = 1 year
-  #  - Prebreeding census, female-based
-  #  - All vital rates assumed to be constant
-  #-------------------------------------------------
-  
-  #population projection part
-  # Initial population sizes
-  N1[1]<-n1.start
-  Nad[1]<-nad.start
-  # Population growth rate
-  for (t in 1:(nyears-1)){
-    lambda[t] <- Ntot[t+1] / Ntot[t]
-  }
-  #prior for survey detectin probability
-  p.surv~dunif(0,1)
-  #  productivity
-  for (t in 1:(nyears-1)){
-    f[t] <- fec
-  }
-  #mean.fec ~ dunif(0, 20)
-  
-  # 3.1. Likelihood for population population count data (state-space model)
-  # 3.1.1 System process
-  for (t in 2:nyears){
[…]

### 2021-10-12 scripts/ARCHIVE/IPMNimble_markedasYoYandAD.R

@@ -1,278 +1,0 @@
-# 11.3. Example of a simple IPM (counts, capture-recapture, reproduction)
-# 11.3.1. Load data
-# Population counts (from years 1 to n.years)
-
-###### FULL IPM ######
-
-library("nimble")
-IPMmod<-nimbleCode({
-  #-------------------------------------------------
-  #  Integrated population model
-  #  - Age structured model with 2 age classes: 
-  #		1-year old and adult (at least 2 years old)
-  #  - Age at first breeding = 1 year
-  #  - Prebreeding census, female-based
-  #  - All vital rates assumed to be constant
-  #-------------------------------------------------
-  
-  #population projection part
-  # Initial population sizes
-  n1.start ~ dunif(0, 1000)
-  nad.start ~ dunif(0, 1000)
-  N1[1] <- n1.start
-  Nad[1] <- nad.start
-  # Population growth rate
-  for (t in 1:(nyears-1)){
-    lambda[t] <- Ntot[t+1] / Ntot[t]
-  }
-  #prior for survey detectin probability
-  p.surv~dunif(0,1)
-  #  productivity
-  for (t in 1:(nyears-1)){
-    f[t] <- fec
-  }
-  #mean.fec ~ dunif(0, 20)
-
-  # 3.1. Likelihood for population population count data (state-space model)
-  # 3.1.1 System process
-  for (t in 2:nyears){
-    #I think this was causing the problem, it was a little mis-specified
-    #Double check that this is right, but it doesnt cause an error in
-    #Slice samplers this way
-    #mean1[t]<-((f[t-1]*N1[t-1])+(f[t-1]*Nad[t-1]))*mean.phi[1]
-    #for reference, this is how it was:
-    #mean1[t] <- f[t-1] * mean.phi[1] * Ntot[t-1]
-    N1[t] ~ dpois(((f[t-1]*N1[t-1])+(f[t-1]*Nad[t-1]))*mean.phi[1])
-    #N1[t] ~ dpois(mean1[t])
-    #Also changed that the total is the sum of N1 and Nad, it was just Ntot before
-    Nad[t] ~ dbin(mean.phi[2],(Ntot[t-1]))
-  }
-  for (t in 1:nyears){
-    Ntot[t] <- round(Nad[t] + N1[t])
-  }
-
-  # 3.1.2 Observation process
-  for(n in 1:n.sam){
-    for (t in 1:nyears){
[…]

### 2021-10-12 scripts/ARCHIVE/IPMSimData.R

@@ -1,458 +1,0 @@
-#HAS - need to comment throughout, but got the MR data (I think) how it should be
-
-
-#altered IPM sim with placeholder for nest success
-#population projections and count code is from BPA WinBUGS appendix 2
-#code adapted from WinBugsBPA Appendix 2, code to simulate IPM
-# link to code is here: https://www.vogelwarte.ch/assets/files/publications/BPA/Web%20appendix%202.txt
-
-library(nimble)
-
-n.years=10; n.data=c(50,50,100);init.age = c(100,100); 
-phi.1=0.3; phi.ad=0.77;p.1=0.98; p.ad=0.3;
-p.sur=0.3
-
-n.initiation.dates <- 31
-
-max.nest.age <- 30
-first.initiation.date <- 1
-last.fledge.date <- n.initiation.dates + max.nest.age
-season.length <- last.fledge.date - first.initiation.date + 1 + 2
-
-# mean clutch size
-mean.clutch.size <- 2.5
-
-# daily nest survival
-phi.nest <- 0.975
-
-prop.nests.found <- 0.8
-
-n.sam<-3
-
-visit.interval <- 3
-
-simIPMdata<-function(n.years, n.data, init.age, phi.1, phi.ad, p.1,p.ad,p.sur,
- n.initiation.dates, max.nest.age, first.initiation.date, last.fledge.date, 
- season.length, mean.clutch.size, phi.nest, prop.nests.found, visit.interval, n.sam){
-  
-  ti<-n.years
-  ni<-n.years-1
-  nd<-n.data
-  Ni<-init.age
-  phi1<-phi.1
-  phi2<-phi.ad
-  pjuv<-p.1
-  psur<-p.sur
-  padults<-p.ad
-  n.initiation.dates <- n.initiation.dates
-  max.nest.age <- max.nest.age
-  first.initiation.date <- first.initiation.date
-  last.fledge.date <- n.initiation.dates + max.nest.age
-  season.length <- last.fledge.date - first.initiation.date + 1 + 2
-  mean.clutch.size <- mean.clutch.size
-  phi.nest <- phi.nest
-  prop.nests.found <- prop.nests.found
-  visit.interval <- visit.interval
-  n.sam <- n.sam
-  
-  # DERIVE TRUE FECUNDITY
[…]

### 2021-10-12 scripts/ARCHIVE/LambdaCombos.R

@@ -1,128 +1,0 @@
-
-#Here is a little code to pick scenarios of parameter values
-# that align with lambda
-
-getNviable <- function() {
-  # TODO
-  # 1 - figure out how many viable combos there are - then add them to the scenarios table
-  # 2 - what ARE the viable combinations
-  
-  # NOW THE EXCITING PART - how many valid combos
-  #creat a matrix for all possible combinations of these 3
-  coms <- expand.grid(fec = fec, phi1 = surv, phiad = surv)
-  #output vector
-  lams<-numeric(length(coms[,1]))
-  #leslie matrix for each
-  mats<-array(dim=c(length(coms[,1]), 2,2))
-  
-  for(i in 1:length(coms[,1])){ #loop over combinations
-    #leslie matrix for each combination
-    mats[i,1,]<-c(coms[i,"phi1"]*coms[i,"fec"],coms[i,"phi1"]*coms[i,"fec"])
-    mats[i,2,]<-c(coms[i,"phiad"], coms[i,"phiad"])
-    #Get the eigenvalue for each leslie matrix, which is the lambda
-    lams[i]<-eigen(mats[i,,])$values[1]
-  }
-  
-  return(cbind(coms, lams))
-}  
-
-#saveRDS(scenarios, scenarios.RDS)
-
-      # #look at how many are within these arbitrary bounds
-      # length(which(lams>=0.95 & lams<=1.05))
-      # #put those that we care about in a matrix and pull out the values for the parameters
-      # intv<-which(lams>=0.95 & lams<=1.05)
-      # scenmat<-matrix(nrow=length(intv), ncol=3)
-      # for(i in 1:length(scenmat[,1])){
-      #   scenmat[i,1]<-phi1[coms[intv[i],1]]
-      #   scenmat[i,2]<-fec[coms[intv[i],2]]
-      #   scenmat[i,3]<-phiad[coms[intv[i],3]]
-
-#HANNAH STUFF ########
-
-#make a sequence of adult and juvenile survival
-phiad<-phi1<-seq(0.05,0.95,by=0.05)
-#make the same length for fecundity
-fec<-seq(0.25,4, length.out=length(phiad))/2
-#creat a matrix for all possible combinations of these 3
-coms<-permutations(length(phiad),3,replace=T)
-#output vector
-lams<-numeric(length(coms[,1]))
-#leslie matrix for each
[…]

### 2021-10-12 scripts/ARCHIVE/SJC code from SEFS590W2020_cjs.R

@@ -1,324 +1,0 @@
-#original code from SEFS 590 winter 2020 class for reference
-#########
-
-#SEFS 590 - University of Washington
-#Week 5 - Cormack-Jolly-Seber Model
-#Simulate a CJS data set
-#fit in optim and using 3 different methods in JAGS 
-#(multinomial on encounter histories, state-space, multinomial on m-array) 
-#Code updated 1 February 2020
-#Code written by Sarah Converse, sconver@uw.edu 
-#CJS simulation code and m-array analysis code adapted from Kery and Schaub
-
-#############################################################################
-
-
-#############################################################################
-
-#Simulate CJS data 
-n.occ <- 3 #number of occasions
-marked <- rep(50,n.occ) #number of inds marked per occasion
-#matrix of values for phi and p (occasion - 1)*inds
-PHI <- matrix(c(0.4,0.4), ncol = n.occ-1, nrow = sum(marked),byrow=TRUE)
-P <- matrix(c(0.7,0.7), ncol = n.occ-1, nrow = sum(marked),byrow=TRUE)
-
-#simulate CJS data 
-CJS.sim <- function(marked,P,PHI){
-  n.occ <- dim(PHI)[2]+1 #number of occasions
-  EH <- matrix(0, ncol = n.occ, nrow = sum(marked)) #assign matrix
-  first <- rep(1:length(marked),marked[1:length(marked)]) #get first encounter occasion for each ind
-  for(i in 1:sum(marked)){ #loop over inds
-    EH[i,first[i]] <- 1 #condition on first capture
-    if(first[i]==n.occ) next #if first capture occasion for ind is last occasion, next
-    for(t in (first[i]+1):n.occ){ #loop over time 
-      surv <- rbinom(1,1,PHI[i,t-1]) #Bernoulli survival trial
-      if(surv == 0) break #if didn't survive, done
-      cap <- rbinom(1, 1, P[i,t-1]) #if survived, Bernoulli capture trial
-      if (cap==1) EH[i,t] <- 1 #fill in the EH
-    }
-  }
-  return(EH)
-}
-#run the simulation to get encounter histories 
-EH <- CJS.sim(marked,P,PHI)
-
[…]

### 2021-10-12 scripts/ARCHIVE/check-trace.R

@@ -1,33 +1,0 @@
-# example code for checking traces
-library(here)
-aaa <- readRDS(here("data", "out1_1-1.Rdata"))
-bbb <- readRDS(here("data", "out1_1-2.Rdata"))
-ccc <- readRDS(here("data", "out1_1-3.Rdata"))
-for (i in 1:18) {
-  plot(aaa[, i], type = "l", col = 1)
-  lines(bbb[, i], col = 2)
-  lines(ccc[, i], col = 3)
-}
-colnames(aaa)
-
-aaa <- readRDS(here("data", "out6_1-1.Rdata"))
-bbb <- readRDS(here("data", "out6_1-2.Rdata"))
-ccc <- readRDS(here("data", "out6_1-3.Rdata"))
-for (i in 1:18) {
-  plot(aaa[, i], type = "l", col = 1)
-  lines(bbb[, i], col = 2)
-  lines(ccc[, i], col = 3)
-}
-colnames(aaa)
-
-# AEB notes
-# this is not good
-aaa <- readRDS(here("data", "out11_1-1.Rdata"))
-bbb <- readRDS(here("data", "out11_1-2.Rdata"))
-ccc <- readRDS(here("data", "out11_1-3.Rdata"))
-for (i in 1:18) {
-  plot(aaa[, i], type = "l", col = 1)
-  lines(bbb[, i], col = 2)
-  lines(ccc[, i], col = 3)
-}
-colnames(aaa)

### 2021-10-12 scripts/ARCHIVE/testAllMods.R

@@ -1,163 +1,0 @@
-library(here)
-library(nimble)
-
-# load data
-scenarios <- readRDS(here("data", "scenarios.RDS"))
-low.lam.combos <- readRDS(here("data", "low.lam.combos.RDS"))
-med.lam.combos <- readRDS(here("data", "med.lam.combos.RDS"))
-high.lam.combos <- readRDS(here("data", "high.lam.combos.RDS"))
-
-# functions
-source(here("scripts", "current version",
-            "1 - simulating data", "IPM_sim_2.0function.R"))
-source(here("scripts", "current version",
-            "2 - models", "IPM_marray.R"))
-source(here("scripts", "current version",
-            "4 - run models", "run_scenarios_helperFns.R"))
-
-low.comb <- low.lam.combos[sample(1:5000, 1), 1:3]
-med.comb <- med.lam.combos[sample(1:5000, 1), 1:3]
-high.comb <- high.lam.combos[sample(1:5000, 1), 1:3]
-
-lowpopTraj <- simPopTrajectory(n.years=15,
-                               n.data.types=c(0.25,0.25,0.25),
-                               age.init=c(150,150),
-                               phi.1=as.numeric(low.comb[2]),
-                               phi.ad=as.numeric(low.comb[3]),
-                               f=as.numeric(low.comb[1]))
-
-medpopTraj <- simPopTrajectory(n.years=15,
-                               n.data.types=c(0.25,0.25,0.25),
-                               age.init=c(150,150),
-                               phi.1=as.numeric(med.comb[2]),
-                               phi.ad=as.numeric(med.comb[3]),
-                               f=as.numeric(med.comb[1]))
-
-highpopTraj <- simPopTrajectory(n.years=15,
-                                n.data.types=c(0.25,0.25,0.25),
-                                age.init=c(150,150),
-                                phi.1=as.numeric(high.comb[2]),
-                                phi.ad=as.numeric(high.comb[3]),
-                                f=as.numeric(high.comb[1]))
-
-# simulate data
-detect.l <- 0.3
-detect.m <- 0.5
-detect.h <- 0.8
-
-detect <- c(detect.l, detect.m, detect.h)
-
-nb <- 100000#0 #burn-in
-ni <- nb + nb #total iterations
-nt <- 10  #thin
-nc <- 3  #chains
-
-# TODO #####
-# Fix this section
-assign("test8", outIPM)
-
[…]

### 2021-10-12 scripts/ARCHIVE/test_allMods_lowTraj.R

@@ -1,503 +1,0 @@
-library(here)
-library(nimble)
-
-# load data
-scenarios <- readRDS(here("data", "scenarios.RDS"))
-low.lam.combos <- readRDS(here("data", "low.lam.combos.RDS"))
-med.lam.combos <- readRDS(here("data", "med.lam.combos.RDS"))
-high.lam.combos <- readRDS(here("data", "high.lam.combos.RDS"))
-
-# functions
-source(here("scripts", "current version",
-            "1 - simulating data", "IPM_sim_2.0function.R"))
-source(here("scripts", "current version",
-            "2 - models", "IPM_marray.R"))
-source(here("scripts", "current version",
-            "4 - run models", "run_scenarios_helperFns.R"))
-
-low.comb <- low.lam.combos[sample(1:5000, 1), 1:3]
-
-lowpopTraj <- simPopTrajectory(n.years=15,
-                               n.data.types=c(0.25,0.25,0.25),
-                               age.init=c(150,150),
-                               phi.1=as.numeric(low.comb[2]),
-                               phi.ad=as.numeric(low.comb[3]),
-                               f=as.numeric(low.comb[1]))
-
-# simulate data
-detect.l <- 0.3
-detect.m <- 0.5
-detect.h <- 0.8
-
-detect <- c(detect.l, detect.m, detect.h)
-
-nb <- 100000#0 #burn-in
-ni <- nb + nb #total iterations
-nt <- 10  #thin
-nc <- 3  #chains
-
-# TODO #####
-# Fix this section
-assign("test8", outIPM)
-
-saveRDS(test7, "test7.RDS")
-saveRDS(test8, "test8.RDS")
-#######
-
-for (d in 1:3) {
-  lowpopDat <- simData (indfates = lowpopTraj$indfates,
-                        n.years = 15,
-                        n.data.types = c(0.25,0.25,0.25),
-                        ADonly = T,
-                        p.1 = detect[d],
-                        p.ad = detect[d],
-                        p.count = detect[d],
-                        p.prod = detect[d],
-                        BinMod = T,
-                        n.sam = 3,
-                        sig = 0,
-                        productivity = T)
-  medpopDat <- simData (indfates = lowpopTraj$indfates,
[…]

### 2021-10-12 scripts/ARCHIVE/test_p_surv.R

@@ -1,226 +1,0 @@
-# todo
-
-# simulate trajectory
-
-# load data
-scenarios <- readRDS(here("data", "scenarios.RDS"))
-low.lam.combos <- readRDS(here("data", "low.lam.combos.RDS"))
-med.lam.combos <- readRDS(here("data", "med.lam.combos.RDS"))
-high.lam.combos <- readRDS(here("data", "high.lam.combos.RDS"))
-
-# functions
-source(here("scripts", "current version",
-            "1 - simulating data", "IPM_sim_2.0function.R"))
-source(here("scripts", "current version",
-            "2 - models", "IPM_marray.R"))
-source(here("scripts", "current version",
-            "4 - run models", "run_scenarios_helperFns.R"))
-
-
- #low.comb <- low.lam.combos[sample(1:5000, 1), 1:3]
- #med.comb <- med.lam.combos[sample(1:5000, 1), 1:3]
- #high.comb <- high.lam.combos[sample(1:5000, 1), 1:3]
-# 
-# # thought
-# # make adult survival be higher?? force it
-# 
-# lowpopTraj <- simPopTrajectory(n.years=15,
-#                                n.data.types=c(0.25,0.25,0.25),
-#                                age.init=c(150,150),
-#                                phi.1=as.numeric(low.comb[2]),
-#                                phi.ad=as.numeric(low.comb[3]),
-#                                f=as.numeric(low.comb[1]))
-# # returns indfates and n
-# 
-# medpopTraj <- simPopTrajectory(n.years=15,
-#                                n.data.types=c(0.25,0.25,0.25),
-#                                age.init=c(150,150),
-#                                phi.1=as.numeric(med.comb[2]),
-#                                phi.ad=as.numeric(med.comb[3]),
-#                                f=as.numeric(med.comb[1]))
-# 
-# highpopTraj <- simPopTrajectory(n.years=15,
-#                                 n.data.types=c(0.25,0.25,0.25),
-#                                 age.init=c(150,150),
-#                                 phi.1=as.numeric(high.comb[2]),
-#                                 phi.ad=as.numeric(high.comb[3]),
-#                                 f=as.numeric(high.comb[1]))
[…]

### 2021-10-12 scripts/ARCHIVE/testmodels.R

@@ -1,202 +1,0 @@
-# todo
-
-# simulate trajectory
-
-# load data
-scenarios <- readRDS(here("data", "scenarios.RDS"))
-low.lam.combos <- readRDS(here("data", "low.lam.combos.RDS"))
-med.lam.combos <- readRDS(here("data", "med.lam.combos.RDS"))
-high.lam.combos <- readRDS(here("data", "high.lam.combos.RDS"))
-
-# functions
-#source(here("scripts", "in progress scripts", "simHelperFns.R"))
-source(here("scripts", "in progress scripts", "IPM_sim_2.0function.R"))
-source(here("scripts", "in progress scripts", "compute_time_calc.R"))
-
-low.comb <- low.lam.combos[sample(1:5000, 1), 1:3]
-med.comb <- med.lam.combos[sample(1:5000, 1), 1:3]
-high.comb <- high.lam.combos[sample(1:5000, 1), 1:3]
-
-# thought
-# make adult survival be higher?? force it
-
-lowpopTraj <- simPopTrajectory(n.years=15, 
-                               n.data.types=c(0.25,0.25,0.25),
-                               age.init=c(150,150), 
-                               phi.1=as.numeric(low.comb[2]), 
-                               phi.ad=as.numeric(low.comb[3]), 
-                               f=as.numeric(low.comb[1]))
-# returns indfates and n
-
-medpopTraj <- simPopTrajectory(n.years=15, 
-                               n.data.types=c(0.25,0.25,0.25),
-                               age.init=c(150,150), 
-                               phi.1=as.numeric(med.comb[2]), 
-                               phi.ad=as.numeric(med.comb[3]), 
-                               f=as.numeric(med.comb[1]))
-
-highpopTraj <- simPopTrajectory(n.years=15, 
-                                n.data.types=c(0.25,0.25,0.25),
-                                age.init=c(150,150), 
-                                phi.1=as.numeric(high.comb[2]), 
-                                phi.ad=as.numeric(high.comb[3]), 
-                                f=as.numeric(high.comb[1]))
-
-
-# simulate data
-
-detect.l <- 0.3
-detect.m <- 0.5
-detect.h <- 0.8
-
-# next steps
-# abby needs to make the repository more readable
-
-lowpopDat <- simData (indfates = lowpopTraj$indfates, 
-                      n.years = 15, 
-                      n.data.types = c(0.25,0.25,0.25), 
-                      ADonly = T, 
[…]

### 2021-10-12 README.Rmd
more reorg

--- 2021-08-17 README.Rmd
@@
 
-Something here about background?
+## TITLE WITH LINK TO ARTICLE WEBPAGE
 
-# Scripts, current version
-## 0 - preparing scenarios
-### compute_time_calc.R
+AUTHORS
+
+JOURNAL NAME
+
+DOI:
+
+Please contact the corresponding author for questions about the code or data: CORRESPONDING AUTHOR NAME AND EMAIL
+
+## Citation
+
+CITATION
+
+## Abstract
+
+Integrated population models (IPMs) are an increasingly popular tool for population modeling in ecology.
+By combining datasets in a unified analysis, this framework links demographic rates with population dynamics and can thereby reduce bias, improve precision, and estimate parameters that would otherwise be unidentifiable using traditional approaches.
+However, the performance of IPMs in different situations remains underexplored and the conditions under which IPMs produce more precise estimates compared with those produced in singular analyses are poorly understood.
+In this study, we assessed the performance of IPMs across a range of life history parameters, population trajectories, and data availability scenarios using simulated count, mark-resight, and productivity data that could be typical for a passerine monitoring program.
+We examined relative bias and root mean squared error across model frameworks when all datasets were included compared with when one or two datasets were omitted and when the included datasets were of poorer quality (i.e., lower detection probabilities).
+In general, parameter estimates showed higher bias when the datasets that more directly informed those parameters were omitted or when the true data generating parameters were representative of more extreme life history traits.
+Higher detection rates alone did little to reduce bias in demographic rate parameters.
+However, having higher detection rates in count surveys was more effective at reducing uncertainty in abundance trends compared with using more of the available datasets.
+Our results suggest that the unique context of a given monitoring program and study species will largely determine whether the collection of additional data versus improving the quality (i.e., detection rates or sample sizes) of existing data streams more effectively improves model estimates.
+Practitioners should carefully consider the best ways of answering the specific ecological or management questions of interest (i.e., estimating demography versus detecting trends) given the life history of the species and the costs of additional data collection.
+This work provides a foundation from which to further explore the performance of IPM frameworks and is useful when designing monitoring programs and considering the trade-offs between the costs and benefits of collecting additional data when monitoring resources are scarce.
+
+## Repository description
+
+### Data
+
+Contains true parameter combinations used to simulate population trajectories and observation data.
+All viable parameter combinations are in files `-.lam.combos.RDS` and the selected parameter combinations used to simulate data are in files `-.lam.params.RDS`, where "-" is "low", "med", or "high".
+Folders labeled `-Trajectories` contain the simulated population trajectories for each of the scenarios found in `-.lam.params.RDS`.
+Each simulated trajectory is labeled with the parameter combination number and the simulation number for that scenario.
+For example, in the `highTrajectory` folder: `highpopTraj-1-2.RDS` corresponds to the first parameter combination in `high.lam.param.RDS` and is the second population trajectory simulated using those parameter values.
+
+File `true.vals.csv` is the true parameter values used in each scenario. `scenario_ID.csv` and `scenarios.RDS` track the combinations of detection probabilities and IPM models to be run.
+
+### Scripts
+
+#### 1 - Simulations
+
+##### 0 - preparing scenarios
+
+###### compute_time_calc.R
 Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time for each model.
 
-### generate_scenarios.R
+###### generate_scenarios.R
 
-Function 'getNviable' for finding the combinations of parameters that give a population growth rate ($\lambda$) within certain bounds, using the eigenvalue from the Leslie matrix.
+Function `getNviable` for finding the combinations of parameters that give a population growth rate ($\lambda$) within certain bounds, using the eigenvalue from the Leslie matrix.
 The output from this function produces the parameter scenarios used to simulate population trajectories, observation data from the trajectories, and fit the the IPM models.
 
-## 1 - simulating data
-### IPM_sim_2.0function.R
-Script contains two functions for simulating data: 'simPopTrajectory' and 'simData'. 
-The 'simPopTrajectory' function simulates the true population trajectory by individual and age class, starting from the stable age distribution.
+##### 1 - simulating data
+
+###### IPM_sim_2.0function.R
+
+Script contains two functions for simulating data: `simPopTrajectory` and `simData`.
+
+The `simPopTrajectory` function simulates the true population trajectory by individual and age class, starting from the stable age distribution.
 Individual fates are retained throughout total number of years.
 
-The 'simData' function uses the output from the 'simPopTrajectory' function to simulate the observation process for each data type.
-The true individual data is broken down into independent sets, each is subject to observation error.
+The `simData` function uses the output from the `simPopTrajectory` function to simulate the observation process for each data type.
+The true individual data is broken down into independent sets and each is subject to observation error.
 Mark-resight, count, and reproductive success data are output from this function.
 
-### simulateTrajectories.R
+###### simulateTrajectories.R
 
-Script that pulls from the scenarios that were created in '0 - preparing scenarios' and the functions in IPM_sim_2.0function.R to simulate observation data sets for model fitting.
+Script that pulls from the scenarios that were created in `0 - preparing scenarios` and the functions in `IPM_sim_2.0function.R` to simulate observation data sets for model fitting.
 
-## 2 - models
+##### 2 - models
 
-### IPMinitvalues.R
+###### IPMinitvalues.R
 Functions for putting data in the form required by the survival model.
 
-### IPM_marray.R
-Script contains 4 IPM NIMBLE models.
+###### IPM_marray.R
+Script contains 4 NIMBLE models.
 
- 1. full IPM model ('IPMmod') where all three datasets are included
- 2. model without productivity dataset ('nonests')
- 3. model without survival model ('nomr')
- 4. abundance only model without productivity or survival ('abundonly')
+ 1. full IPM model (`IPMmod`) where all three datasets are included
+ 2. model without productivity dataset (`nonests`)
+ 3. model without survival model (`nomr`)
+ 4. abundance only model without productivity or survival (`abundonly`)
 
-## 3 - run models
-### run_scenarios_helperFns.R
+##### 3 - run models
+
+###### run_scenarios_helperFns.R
+
 Script with functions and code to run NIMBLE models.
-Function 'marray' transforms the survival data capture histories to m-array format.
+Function `marray` transforms the survival data capture histories to m-array format.
 MCMC settings in this script, with functions for each of the 4 models.
-Each model has a run model function that defines constants, initial values, data, parameters to monitor, and code to build/run the models in NIMBLE.
+Each model has a "run model" function that defines constants, initial values, data, parameters to monitor, and code to build/run the models in NIMBLE.
 
-### run_scenarios.R
-Script that pulls from the scripts above to run models in parallel and output results.
+###### run_scenarios.R
 
-## 4 - process results
-### 01_load_and_process_data.R
+Main workhorse script that pulls from the scripts above to run models in parallel and output results.
+
+##### 4 - process results
+
+###### 01_load_and_process_data.R
+
 Script that checks the Gelman-Rubin convergence statistic for each output MCMC sample, then thins the chains to reduce file size and places the output into categories based on the population growth rate, $\lambda$.
 
-### 02_compute_geom_means.R
+###### 02_compute_geom_means.R
+
 Computes the geometric means for $\lambda$ in each MCMC output.
 
-# Data
-Contains true parameter scenarios used to simulate population trajectories and observation data, all viable parameter combinations in files '-.lam.combos.RDS' and the selected parameter combinations used to simulate data are in files '-.lam.params.RDS' *(where "-" is low, med, or high)*.
-Folders labeled '-Trajectories' contain the simulated population trajectories for each of the scenarios found in '-.lam.params.RDS'.
-Each simulated trajectory is labeled with the scenario and the simulation number for that scenario.
-For example, in the 'highTrajectory' folder: 'highpopTraj-1-2.RDS' corresponds to the first parameter scenario in 'high.lam.param.RDS' and is the second population trajectory simulated using those parameter values.
+#### 2 - Figures
 
-File 'true.vals.csv' is the true parameter values used in each scenario. 'scenario_ID.csv' tracks the combinations of detection probabilities and IPM models.
+### Results

### 2021-10-12 scripts/2 - Figures/PaperFigures.Rmd
more reorg

--- 2021-09-03 figures/PaperFigures.Rmd
@@
 library(Hmisc)
+library(tidybayes)
+library(wesanderson)
 
@@
 
-<!-- Euring figures -->
-```{r phiad}
-pal1 <- diverging_hcl(n = 7, palette = "Blue-Yellow 3")
-phiad <- rel.bias.vals %>% filter(variable == 'Adult survival') 
-summary(phiad$value) # look at range for scale
+```{r lambdaTrends}
+pal <- rev(wes_palette("Zissou1", 3, type = "continuous"))
 
-## fix faceted labels
-rel.bias.vals$lambda.scenario <- factor(rel.bias.vals$lambda.scenario, levels = c("Decreasing", "Stable", "Increasing"),
-                                ordered = TRUE, labels = c(expression(paste("Decreasing ", lambda)), 
-                                                           expression(paste("Stable ", lambda)), 
-                                                           expression(paste("Increasing ", lambda))))
-rel.bias.vals$simscenarios <- factor(rel.bias.vals$simscenarios, levels = c("L", "M", "H"),
-                                ordered = TRUE, labels = c(expression(paste("low ", italic(p))), 
-                                                           expression(paste("med ", italic(p))), 
-                                                           expression(paste("high ", italic(p)))))
+row.low <- read_csv(here("results", "row_low.csv"))
+row.med <- read_csv(here("results", "row_med.csv"))
+row.high <- read_csv(here("results", "row_high.csv"))
 
-p1 <- ggplot(rel.bias.vals %>% filter(variable == 'Adult survival'),
-             aes(x = factor(phiad.true), y = variable, fill = value)) +
-  geom_tile(color = 'grey50') +
-  xlab('\nTrue Adult Survival') + ylab("") +
-  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
-  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
-               nest_line = TRUE) +
-    scale_fill_gradient2(name = "Relative\nbias",
-                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
-                      # mid = "white", low = "#9FA2FF", high = "#BAAE00", midpoint = 0,
-                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
-                      breaks = c(-0.15, -0.05, 0, 0.05), n.breaks = 4, labels = c("-0.15", "-0.05", "0", "0.05")) +
-  theme_light() +
-  theme(legend.position = 'right',
-        legend.text = element_text(size = 8, angle = 90),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
-        strip.text.x = element_text(face = "bold", color = "black"),
-        strip.background = element_rect(fill = NA, color = "black"),
-        axis.text.x = element_text(angle = 90, vjust = 1.5),
-        ggh4x.facet.nestline = element_line(size = 0.5),
-        panel.border = element_rect(color = "black", fill = NA),  
-        panel.spacing.x = unit(0.75, "line")) +
-        # panel.spacing = unit(1, "lines")) +
-  ggtitle("Adult Survival")
-p1
+scenario_ID <- read_csv(here("data", "scenario_ID.csv")) %>% 
+  select(-c(n.viable.combinations, priority))
+
+# Reformat for plotting
+toplot1 <- row.low %>%
+  select(contains("geomean"), scenario, sims, simscenarios, Quantile) %>%
+  #group_by(model, detection)
+  pivot_longer(cols = starts_with("geomean"), names_to = "Year") %>%
+  filter(!is.na(value)) %>%
+  mutate(Year = str_remove(Year, "geomean\\.")) %>%
+  mutate(Year = as.numeric(Year)) %>%
+  group_by(scenario, sims, simscenarios, Year, Quantile) %>% # checked through here - TODO
+  # summarise(low = quantile(value, 0.025),
+  #           med = quantile(value, 0.5),
+  #           high = quantile(value, 0.975)) %>%
+  ungroup() %>%
+  left_join(scenario_ID, by = "simscenarios") %>% 
+  mutate(scenario = as.factor(scenario),
+         sims = as.factor(sims), 
+         simscenarios = as.factor(simscenarios))
+
+# use reshape cast to spread the quantiles out into lower middle upper for line plot
+toplot2 <- row.med %>%
+  select(contains("geomean"), scenario, sims, simscenarios, Quantile) %>%
+  #group_by(model, detection)
+  pivot_longer(cols = starts_with("geomean"), names_to = "Year") %>%
+  filter(!is.na(value)) %>%
+  mutate(Year = str_remove(Year, "geomean\\.")) %>%
+  mutate(Year = as.numeric(Year)) %>%
+  group_by(scenario, sims, simscenarios, Year, Quantile) %>% # checked through here - TODO
+  # summarise(low = quantile(value, 0.025),
+  #           med = quantile(value, 0.5),
+  #           high = quantile(value, 0.975)) %>%
+  ungroup() %>%
+  left_join(scenario_ID, by = "simscenarios") %>% 
+  mutate(scenario = as.factor(scenario),
+         sims = as.factor(sims), 
+         simscenarios = as.factor(simscenarios))
+
+toplot3 <- row.high %>%
+  select(contains("geomean"), scenario, sims, simscenarios, Quantile) %>%
+  #group_by(model, detection)
+  pivot_longer(cols = starts_with("geomean"), names_to = "Year") %>%
+  filter(!is.na(value)) %>%
+  mutate(Year = str_remove(Year, "geomean\\.")) %>%
+  mutate(Year = as.numeric(Year)) %>%
+  group_by(scenario, sims, simscenarios, Year, Quantile) %>% # checked through here - TODO
+  # summarise(low = quantile(value, 0.025),
+  #           med = quantile(value, 0.5),
+  #           high = quantile(value, 0.975)) %>%
+  ungroup() %>%
+  left_join(scenario_ID, by = "simscenarios") %>% 
+  mutate(scenario = as.factor(scenario),
+         sims = as.factor(sims), 
+         simscenarios = as.factor(simscenarios))
+
+toplot <- bind_rows(toplot1, toplot2, toplot3) %>% 
+  select(-simscenarios) %>% 
+  transform(dataset = ifelse(is.na(det.MR)&!is.na(det.prod), 'Counts+Prod', 
+                             ifelse(!is.na(det.MR)&is.na(det.prod), 'Counts+MR',
+                                    ifelse(is.na(det.MR)&is.na(det.prod), 'Counts', 'Full')))) %>% 
+  group_by(Quantile, Year, det.abund, lambda, dataset, det.MR, det.prod) %>% 
+  # took mean over demographic scenario (n = 25) and replicate (n = 50)
+  # and mark recapture detection and fecundity
+  # AEB - is it ok to take mean of quantiles? review here ######
+  summarize(value = mean(value), .groups = "keep") %>% 
+  ungroup() %>% 
+  mutate(Quantile = str_remove(Quantile, "\\%")) %>% 
+  mutate(Quantile = paste("X", Quantile, sep = "")) %>% 
+  reshape2::dcast(dataset + Year +  det.MR + det.prod + det.abund   + lambda ~ Quantile, value.var = "value") %>% 
+  mutate(Year = Year + 1) %>% 
+  filter(Year %in% c(15)) %>% 
+  mutate(Year = factor(Year)) %>% 
+  mutate(det.abund = factor(det.abund, levels = c("L", "M", "H"))) %>% 
+  mutate(det.prod = factor(det.prod, levels = c("L", "M", "H"))) %>% 
+  mutate(det.MR = factor(det.MR, levels = c("L", "M", "H"))) %>% 
+  transform(lambda = factor(lambda, levels = c("L", "M", "H"), 
+                            labels = c("Decreasing", "Stable", "Increasing"))) %>%
+  transform(dataset = factor(dataset, levels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'),
+                             labels = c('Full', 'Counts+Prod', 'Counts+MR', 'Counts'))) %>% 
+  mutate(intercept = case_when(
+    lambda == "Decreasing" ~ 0.95, 
+    lambda == "Stable" ~ 1,
+    lambda == "Increasing" ~ 1.05))
+
+pdf(here("figures", "lambdaTrends.pdf"), width = 8, height = 8)
+rainbow2 <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")
+ggplot(toplot) +
+  geom_point(aes(x = Year, y = X50, col = det.abund, group = det.abund), position = position_dodge(width = 0.5)) +
+  geom_linerange(aes(x = Year, ymin = X2.5, ymax = X97.5, col = det.abund, group = det.abund), position = position_dodge(width = 0.5)) +
+  geom_hline(aes(yintercept = intercept), linetype = 'dotted') +
+  geom_hline(aes(yintercept = 1.0), linetype = 'solid') +
+  xlab('Years') + 
+  ylab('Lambda') + 
+  #ylim(0.87, 1.14) +
+  facet_grid(dataset ~ lambda, scales = 'free') +
+  theme_bw() +
+  theme(legend.position = 'top',
+        plot.subtitle = element_text(size = 10, hjust = 0.5, vjust = 1)) +
+  scale_color_manual(values = rainbow2[-c(1,4)], name = 'Abundance detection level') 
+dev.off()
 ```
-
-```{r phijuv}
-pal2 <- diverging_hcl(n = 7, palette = "Green-Orange")
-phi1 <- rel.bias.vals %>% filter(variable == 'Juv survival')
-summary(phi1$value)
-
-p2 <- ggplot(rel.bias.vals %>% filter(variable == 'Juv survival'), 
-             aes(x = factor(phi1.true), y = variable, fill = value)) +
-  geom_tile(color = 'grey50') +
-  xlab('\nTrue Juvenile Survival') + ylab("") +
-  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
-  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
-               nest_line = TRUE) +
-  scale_fill_gradient2(name = "Relative\nbias",
-                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
-                      # mid = "white", low = "#11C638", high = "#EF9708", midpoint = 0,
-                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
-                      breaks = c(-0.2, 0, 0.2, 0.5, 0.8), n.breaks = 5, 
-                      labels = c("-0.2", "0", "0.2", "0.5", "0.8")) +
-  theme_light() +
- theme(legend.position = 'right',
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
-        strip.text.x = element_text(face = "bold", color = "black"),
-        strip.background = element_rect(fill = NA, color = "black"),
-        axis.text.x = element_text(angle = 90, vjust = 1.5),
-        ggh4x.facet.nestline = element_line(size = 0.5),
-        panel.border = element_rect(color = "black", fill = NA),  
-        panel.spacing.x = unit(0.75, "line")) +
-        # panel.spacing = unit(1, "lines")) +
-  ggtitle("Juvenile Survival")
-p2
-```
-
-```{r fec}
-pal3 <- diverging_hcl(n = 7, palette = "Cyan-Magenta")
-fec <- rel.bias.vals %>% filter(variable == 'Fecundity')
-summary(fec$value)
-
-p3 <- ggplot(rel.bias.vals %>% filter(variable == 'Fecundity'), 
-             aes(x = factor(fec.true), y = variable, fill = value)) +
-  geom_tile(color = 'grey50') +
-  xlab('\nTrue Fecundity') + ylab("") +
-  # facet_grid(~lambda.scenario+simscenarios, drop = T, scales = 'free_x') +
-  facet_nested(~lambda.scenario+simscenarios, drop = T, scales = 'free_x', labeller = label_parsed,
-               nest_line = TRUE) +
-  scale_fill_gradient2(name = "Relative\nbias",
-                      # mid = "#f7f7f7", low = "#a1d76a", high = "#e9a3c9", midpoint = 0,
-                      # mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0,
-                      mid = "white", low = "#0FCFC0", high = "#F79CD4", midpoint = 0, 
-                      breaks = c(-0.4, -0.2, 0, 0.2), n.breaks = 4, 
-                      labels = c("-0.4", "-0.2", "0", "0.2")) +
-  theme_light() +
-  theme(legend.position = 'right',
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        #strip.background = element_rect(fill = "#EEEEEE", color = "#FFFFFF"),
-        strip.text.x = element_text(face = "bold", color = "black"),
-        strip.background = element_rect(fill = NA, color = "black"),
-        axis.text.x = element_text(angle = 90, vjust = 1.5),
-        ggh4x.facet.nestline = element_line(size = 0.5),
-        panel.border = element_rect(color = "black", fill = NA),  
-        panel.spacing.x = unit(0.75, "line")) +
-        # panel.spacing = unit(1, "lines")) +
-  ggtitle("Fecundity")
-p3
-```
-
-```{r combo}
-## export out
-pdf(here("figures", "params_bias.pdf"), width = 12, height = 8)
-plot_grid(p3, p2, p1, nrow = 3)
-dev.off()
-
-## try with overview plot
-all <- ov + (p3/p2/p1)
-ggsave(filename = here("figures", "all_bias.pdf"), plot = all, width = 22, height = 12)
-#use # for spaces with missing plots
-# layout_string <- 'A#BB
-#                   ##CC
-#                   ##DD'
-# junk <- ov + (p3/p2/p1) + plot_layout(design=layout_string,
-#                               widths=c(3,1,5,6),
-#                               heights=c(5,3))
-```

### 2021-10-12 scripts/figures/PlotTheme.R

@@ -1,55 +1,0 @@
-# CUSTOM PLOT THEME
-
-theme_ipmsim <- function(){ 
-  font <- "Helvetica"   #assign font family up front
-  
-  theme_minimal() %+replace%    #replace elements we want to change
-    
-    theme(
-      
-      #grid elements
-      panel.grid.major = element_blank(),    #strip major gridlines
-      panel.grid.minor = element_blank(),    #strip minor gridlines
-      
-      #since theme_minimal() already strips axis lines, 
-      #we don't need to do that again
-      
-      #text elements
-      plot.title = element_text(             #title
-        family = font, color = "black",      #set font family
-        size = 20,                #set font size
-        face = 'bold',            #bold typeface
-        hjust = 0,                #left align
-        vjust = 2),               #raise slightly
-      
-      plot.subtitle = element_text(          #subtitle
-        family = font, color = "black",           #font family
-        size = 14),               #font size
-      
-      axis.title = element_text(             #axis titles
-        family = font, color = "black",           #font family
-        size = 12),               #font size
-      
-      axis.ticks.x.bottom = element_line(colour = "grey20"), 
-      axis.ticks.y.left = element_line(colour = "grey20"), 
-      
-      axis.text = element_text(              #axis text
-        family = font, color = "black",           #axis family
-        size = 12),                #font size
-      
-      legend.title = element_text(             #axis titles
-        family = font, color = "black",           #font family
-        size = 12),               #font size
-      
-      legend.text = element_text(             #axis titles
-        family = font, color = "black",           #font family
-        size = 12),              #font size
-      
-      strip.text = element_text(             #axis titles
-        family = font, color = "black",           #font family
-        size = 12)               #font size
-      
-      #since the legend often requires manual tweaking 
-      #based on plot content, don't define it here
-    )
-}

### 2021-10-12 scripts/figures/lambdaTrendsPlot.R

@@ -1,168 +1,0 @@
-# AEB NOTE 
-# this needs to be modified
-
-# 9 panel plot
-# rows - low, med, high lambda
-# columns - low, med, high detection
-
-#or 3? column is model type?
-
-library(tidybayes)
-library(ggplot2)
-library(wesanderson)
-library(here)
-library(tidyverse)
-library(reshape2)
-
-pal <- rev(wes_palette("Zissou1", 3, type = "continuous"))
-
-row.low <- read_csv(here("results", "row_low.csv"))
-row.med <- read_csv(here("results", "row_med.csv"))
-row.high <- read_csv(here("results", "row_high.csv"))
-
-scenario_ID <- read_csv(here("data", "scenario_ID.csv")) %>% 
-  select(-c(n.viable.combinations, priority))
-
-# Reformat for plotting
-toplot1 <- row.low %>%
-  select(contains("geomean"), scenario, sims, simscenarios, Quantile) %>%
-  #group_by(model, detection)
-  pivot_longer(cols = starts_with("geomean"), names_to = "Year") %>%
-  filter(!is.na(value)) %>%
-  mutate(Year = str_remove(Year, "geomean\\.")) %>%
-  mutate(Year = as.numeric(Year)) %>%
-  group_by(scenario, sims, simscenarios, Year, Quantile) %>% # checked through here - TODO
-  # summarise(low = quantile(value, 0.025),
-  #           med = quantile(value, 0.5),
-  #           high = quantile(value, 0.975)) %>%
-  ungroup() %>%
-  left_join(scenario_ID, by = "simscenarios") %>% 
-  mutate(scenario = as.factor(scenario),
-         sims = as.factor(sims), 
-         simscenarios = as.factor(simscenarios))
-
-# use reshape cast to spread the quantiles out into lower middle upper for line plot
-
-toplot2 <- row.med %>%
-  select(contains("geomean"), scenario, sims, simscenarios, Quantile) %>%
-  #group_by(model, detection)
-  pivot_longer(cols = starts_with("geomean"), names_to = "Year") %>%
-  filter(!is.na(value)) %>%
-  mutate(Year = str_remove(Year, "geomean\\.")) %>%
-  mutate(Year = as.numeric(Year)) %>%
-  group_by(scenario, sims, simscenarios, Year, Quantile) %>% # checked through here - TODO
[…]

### 2021-10-12 README.Rmd
even more repo updates

@@
 
+Contains code used to produce figures for manuscript.
+
 ### Results
+
+Processed results files referenced in figure code `lowout.csv`, `medout.csv`, and `highout.csv` are quite large and are available upon request.
+`Processed csvs` contains posterior medians for all converged model runs.
+`row_-` files contains estimates of population trend, $\lambda$, where "-" is "low", "med", or "high" and corresponds to a decreasing, stable, or increasing population trend, respectively.

### 2021-10-15 figures/manuscript_figs.R
code for figures in latest manuscript only

@@ -1,0 +1,499 @@
+## IPM Sim Figures
+## Adapted from PaperFigures.RMD
+## A DuVall
+## 15 Oct 2021
+
+library(tidyr)
+library(dplyr)
+library(cowplot)
+library(ggplot2)
+library(coda)
+library(captioner)
+library(knitr)
+library(reshape2)
+library(here)
+library(RColorBrewer)
+library(colorspace)
+library(ggh4x)
+library(Hmisc)
+library(patchwork)
+library(tidybayes)
+
+rainbow2 <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#4aaaa5", "#a3d39c", "#f6b61c", "chocolate2", "red3")
+
+scenarios <- read.csv(file = here::here('data', 'scenario_ID.csv'), header = T, stringsAsFactors = F)
+
+low.params <- readRDS(file = here::here('data', 'low.lam.params.rds')) %>%
+  transform(scenario = 1:25)
+
+med.params <- readRDS(file = here::here('data', 'med.lam.params.rds')) %>%
+  transform(scenario = 1:25)
+
+high.params <- readRDS(file = here::here('data', 'high.lam.params.rds')) %>%
+  transform(scenario = 1:25)
+
+all.meds <- read.csv(file = here::here('figures', 'Processed csvs', 'all.meds.csv'), 
+                     header = T, stringsAsFactors = F)
+
+######################################################
+##################### Bias ###########################
+######################################################
+
+rel.bias <- all.meds %>%
+  transform(p.surv.true = ifelse(det.abund == 'L', 0.3, 
+                                 ifelse(det.abund == 'M', 0.5, 
+                                        ifelse(det.abund == 'H', 0.8, NA)))) %>%
+  transform(mean.p.true = ifelse(det.MR == 'L', 0.3, 
+                                 ifelse(det.MR == 'M', 0.5, 
+                                        ifelse(det.MR == 'H', 0.8, NA)))) %>%
+  #calculate relative bias, mean across scenarios
+  transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
+            phiad.bias = (phiad.obs-phiad.true)/phiad.true,
+            fec.bias = (fec.obs-fec.true)/fec.true,
+            p.surv.bias = (p.surv-p.surv.true)/p.surv.true,
+            mean.p.bias = (mean.p-mean.p.true)/mean.p.true) %>%
[…]

### 2024-10-02 figures/manuscript_figs.R
Formatting figures 3 and 4 for manuscript

@@ -87,7 +87,8 @@ all.meds.sc <- read.csv(file = here::here('figures', 'Processed csvs', 'all.meds
                         header = T, stringsAsFactors = F)
 
 ## just bias; same as above, keep 'scenario'
-rel.bias.sc <- all_meds_sc %>%
+#rel.bias.sc <- all_meds_sc %>%
+rel.bias.sc <- all.meds.sc %>%
   transform(phi1.bias = (phi1.obs-phi1.true)/phi1.true,
             phiad.bias = (phiad.obs-phiad.true)/phiad.true,
             fec.bias = (fec.obs-fec.true)/fec.true) %>%
@@ -337,42 +338,80 @@ dplyr::summarize(value = mean(value), .groups = "keep") %>%
 
 #### Figure 3: RMSE and bias ecological paramters x count survey detection ####
 
+dataset.labs <- c("Full IPM", "Abundance & Prod.", "Abundance & Surv.", "Abundance Only")
+names(dataset.labs) <- c("Full IPM", "Abundance & Productivity", "Abundance & Survival", "Abundance Only")
+lambda.labs <- c("Decrease", "Stable", "Increase")
+names(lambda.labs) <- c("Decreasing", "Stable", "Increasing")
+
 ## bias dot plot
 a1 <- ggplot(rel.bias.few  %>% filter(variable %nin% obs.pars), 
              aes(x = det.abund, y = bias, col = factor(variable), group = factor(variable),
                  shape = factor(variable))) +
-  geom_point() + geom_line() +
+  geom_point() + 
+  geom_line() +
   geom_hline(aes(yintercept = 0), linetype = 'dotted') +
-  xlab('Count survey detection') + ylab('Relative bias') +
-  facet_grid(dataset~lambda.scenario, scales = 'free_x', labeller = label_wrap_gen()) +
-  ylim(c(-1.75, 1.75)) +
+  #xlab('Count survey detection') + ylab('Relative bias') +
+  #facet_grid(dataset~lambda.scenario, scales = 'free_x', labeller = label_wrap_gen()) +
+  #ylim(c(-1.75, 1.75)) +
+  scale_x_discrete(labels = c("L", "M", "H")) +
+  xlab('Count survey detection') + 
+  ylab('Relative bias') +
+  facet_grid(dataset~lambda.scenario, scales = 'free_x', 
[…]

### 2024-10-02 README.Rmd
Update abstract; add title and authors

--- 2021-08-17 README.Rmd
@@
 
-Understanding the processes that control population dynamics in wild populations is a critical component of conservation and management but is frequently limited by missing demographic information, uncertainty, and temporally or spatially misaligned datasets (Schaub & Abadi 2011, Zipkin & Saunders 2017).
-Integrated population models (IPMs; Besbeas et al.
-2002, Brooks et al.
-2004) are an increasingly popular tool in ecology that can help overcome these limitations by combining disparate datasets in a single, unified analysis (Schaub & Abadi 2011).
-This approach can provide several advantages, such as improved precision (Schaub et al.
-2007, Tavecchia 2009, Abadi et al.
-2010) and the estimation of parameters that would otherwise be unidentifiable (Besbeas 2005, Schaub et al.
-2007, Veran & Lebreton 2008), which is often the case for key demographic rates such as recruitment or immigration.
-These advances have practical benefits for resource managers that must make decisions based on limited available information.
-As IPMs become more widely used, however, it is critical that they are robustly evaluated.
-In particular, the limitations of the IPM framework are still underexplored, particularly in situations where available data are sparse, of poor quality, or when data to inform specific parameters are entirely absent.
+# Effects of species life history and data availability on integrate population model performance
 
-Through a simulation analysis, we will identify monitoring programs and demographic scenarios where integrated population models may not lead to improved precision or return unbiased estimates of parameters that would otherwise be unidentifiable when certain datasets are excluded.
-We apply this simulation study to to a passerine bird life history.
-We assume all model assumptions are met and compare parameter estimation for a traditional IPM with data informing abundance (count data), survival (mark-recapture data), and reproductive output (nest monitoring data) to situations when either of the latter two datasets are omitted.
-In doing so, we will refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool or might not be necessary or warrant the collection of additional data.
-Improving our understanding of the strengths and weaknesses of the IPM framework can provide helpful insights for the design and implementation of field survey programs so as to most efficiently make use of limited resources to inform management efforts aimed at monitoring or recovering species that are deemed ecosystem indicators or depleted and in need of conservation intervention.
+Abby E. Bratt, Caroline D. Cappello, Amelia J. DuVall, Hannah A. Sipe, Amanda J.
+Warlick, Beth Gardner, Sarah J. Converse
 
-# Scripts, current version
+Code provided for peer review and may change
 
-## 0 - preparing scenarios
+## Abstract
 
-### compute_time_calc.R
+Integrated population models (IPMs) are an increasingly popular tool for population modeling in ecology.
+By combining datasets in a unified analysis, IPMs link demographic rates with population dynamics and can thereby reduce bias, improve precision, and estimate parameters that would otherwise be unidentifiable using traditional approaches.
+However, the performance of IPMs in different situations remains underexplored and the conditions under which IPMs produce more precise estimates compared with those produced in singular analyses are poorly understood.
+
+We assessed the performance of IPMs across a range of life history parameters, population trajectories, and data availability scenarios using simulated count, mark-resight, and productivity data typical of a passerine monitoring program.
+We examined relative bias and root mean squared error for abundance, trend in abundance, survival, and fecundity across models when all datasets were included versus when one or two datasets were omitted or when the included datasets were of poorer quality (i.e., lower detection probabilities).
+
+Parameter estimates generally showed higher bias when the datasets that more directly informed those parameters were omitted or when the true data-generating parameters were representative of more extreme life history traits (e.g., high survival and low reproduction that characterizes large mammals or the lower survival and higher reproductive output that characterizes insects or amphibians).
+Higher detection probabilities alone did little to reduce bias in demographic rates.
+However, higher detection probabilities in count surveys were more effective at reducing uncertainty in abundance trends than using more of the available datasets.
+
+Our results suggest that the unique context of a given monitoring program and study species will largely determine whether collecting multiple data streams versus collecting a single high-quality data stream (e.g., increasing detection probabilities or sample sizes) will most improve parameter estimates.
+Practitioners should carefully consider the best ways of answering the specific ecological or management questions of interest (e.g., estimating demographic rates versus detecting trends) given the species’ life history and the costs of data collection.
+This work provides a foundation from which to further explore the performance of IPMs and to inform the design of monitoring programs while considering the costs and benefits of data collection.
+
+## Scripts, current version
+
+### 0 - preparing scenarios
+
+#### compute_time_calc.R
 Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time for each model.
 
-### generate_scenarios.R
+#### generate_scenarios.R
 
@@
 
-## 1 - simulating data
+### 1 - simulating data
 
-### IPM_sim_2.0function.R
+#### IPM_sim_2.0function.R
 
@@
 
-### simulateTrajectories.R
+#### simulateTrajectories.R
 
@@
 
-## 2 - models
+### 2 - models
 
-### IPMinitvalues.R
+#### IPMinitvalues.R
 Functions for putting data in the form required by the survival model.
 
-### IPM_marray.R
+#### IPM_marray.R
 Script contains 4 IPM NIMBLE models.
@@
 
-## 3 - run models
+### 3 - run models
 
-### run_scenarios_helperFns.R
+#### run_scenarios_helperFns.R
 
@@
 
-### run_scenarios.R
+#### run_scenarios.R
 
@@
 
-## 4 - process results
+### 4 - process results
 
-### 01_load_and_process_data.R
+#### 01_load_and_process_data.R
 
@@
 
-### 02_compute_geom_means.R
+#### 02_compute_geom_means.R
 
@@
 
-# Data
+## Data
 

### 2024-10-02 README.Rmd
Minor formatting edits

@@
 
-# Effects of species life history and data availability on integrate population model performance
+## Effects of species life history and data availability on integrate population model performance
 
@@
 
-Code provided for peer review and may change
+Code provided for peer review
 
@@
 
-#### compute_time_calc.R
+##### compute_time_calc.R
 Code for approximating the total run time, given number of simulated datasets, number of parameter scenarios, computer cores, and estimated run time for each model.
 
-#### generate_scenarios.R
+##### generate_scenarios.R
 
@@
 
-#### IPM_sim_2.0function.R
+##### IPM_sim_2.0function.R
 
@@
 
-#### simulateTrajectories.R
+##### simulateTrajectories.R
 
@@
 
-#### IPMinitvalues.R
+##### IPMinitvalues.R
 Functions for putting data in the form required by the survival model.
 
-#### IPM_marray.R
+##### IPM_marray.R
 Script contains 4 IPM NIMBLE models.
@@
 
-#### run_scenarios_helperFns.R
+##### run_scenarios_helperFns.R
 
@@
 
-#### run_scenarios.R
+##### run_scenarios.R
 
@@
 
-#### 01_load_and_process_data.R
+##### 01_load_and_process_data.R
 
@@
 
-#### 02_compute_geom_means.R
+##### 02_compute_geom_means.R
 
